\documentclass[11pt,a4paper]{article}
\newcommand{\docTitle}{Developer Guide}
\newcommand{\soloDevAuthor}{}   % the Developer Guide is Vítor Geraldes alone (he authored the code)
\input{preamble.tex}

\begin{document}

\manualfrontmatter{Developer Guide}{
Choupo is a trademark of TalentGround Lda.

Anthropic, Claude, and Claude Code are trademarks or registered trademarks of
Anthropic, PBC.

Linux is a registered trademark of Linus Torvalds.

Python is a trademark or registered trademark of the Python Software
Foundation.
}

% =====================================================================
%   Table of contents
% =====================================================================
\pagenumbering{roman}
\setcounter{tocdepth}{2}
\markboth{Contents}{Contents}
\tableofcontents
\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}

% =====================================================================
\section*{Preface --- the contract this code makes with you}
\addcontentsline{toc}{section}{Preface --- the contract this code makes with you}

If you have written CFD glue for a CFD solver, or stared too long at a
black-box solver, you already know why this project exists.
Every Newton step is auditable. Every dispatch is explicit. Every
dependency is in the C\texttt{++} standard library. The price of that
discipline is that contributors have to play by a few rules. They are
short, and they are non-negotiable.

This guide is opinionated on purpose. Read \S\ref{sec:phil} carefully
--- those three principles override everything else --- and treat
\S\ref{sec:donts} as the project's only blunt list of don'ts.

\paragraph{Independence and provenance.}  Choupo is an independent
project.  Its file-based case organization (\code{controlDict}-style
dictionaries, \code{constant/} and time directories such as \code{0/})
was inspired in part by the conventions popularised by OpenFOAM.  Choupo
is not affiliated with, endorsed by, or sponsored by OpenCFD Ltd or the
OpenFOAM Foundation.  The modules closest in spirit are independent
implementations, not adaptations: the dictionary parser is Choupo's own
component-aware tokenizer; model selection uses explicit factory
registrations (deliberately \emph{not} runtime-selection macros,
\S\ref{sec:factory}); the \code{Rosenbrock23} integrator implements the
published Shampine \& Reichelt (SIAM J.\ Sci.\ Comput.\ 18, 1997)
formulas; and the time-directory writer implements the layout convention
from scratch.  No OpenFOAM source code was adapted or reused.


% =====================================================================
\section{Philosophy --- the contract we promise}\label{sec:phil}

\begin{whatyoulllearn}
The three invariants that every contribution is judged against, in
plain language.
\end{whatyoulllearn}

\begin{enumerate}[leftmargin=*]
  \item \textbf{Pedagogical clarity beats elegance.}
        The simulator must be readable by a chemical-engineering
        student in their third year. No template metaprogramming, no
        auto-magic, no clever short-circuits. Every Newton iteration
        and every $K$-value must be inspectable in both the source and
        the run-time output.

  \item \textbf{Source transparency --- no hidden registration.}
        Every factory base class has an \emph{explicit}
        \code{registerBuiltins()} called from \file{main.cpp}. No
        \code{REGISTER\_TYPE(MyClass)} macros, no static initialisers,
        no linker-dependent constructors. The student grep-searches
        \code{registerBuiltins} and \emph{sees} every built-in type.

  \item \textbf{Dependency purity --- no copyleft or restrictive libraries linked into the binary.}
        \cpp{}~17, the C\texttt{++} standard library, and GNU make.
        Nothing else. No Boost, no Eigen, no Sundials, no CMake.
        Hand-rolled Newton, Gauss elimination, RK4, Wegstein, Michelsen.
\end{enumerate}

\begin{keyidea}
An ``industrial-grade'' contribution that violates any of these will
be rejected even if it converges better and runs faster.
\end{keyidea}

% =====================================================================
\section{Repository tour}

\begin{whatyoulllearn}
What lives where. Useful as a map; you'll come back to it.
\end{whatyoulllearn}

\begin{lstlisting}
Choupo/
|-- Makefile                       top-level: release/debug switch
|-- make/
|   |-- compiler.mk                CXX, CXXFLAGS, PLATFORM detection
|   |-- rules.mk                   pattern rules + dependency tracking
|   `-- wasm.mk                    Emscripten WASM build (gui/)
|-- build/<PLATFORM>[Debug]/       all.o/.d/binary (gitignored)
|-- choupoSolve               symlink -> last built binary
|-- etc/bashrc                     sourceable shell env
|-- bin/                           runCase, runApplication, listCases, cleanCase
|-- data/
|   `-- standards/                  the FROZEN, committee-managed reference tree
|       |-- components/             <name>.dat --- component catalogue (arity-1)
|       |-- materials/              <name>.dat --- material database
|       |-- membranes/              <name>.dat --- NF/RO membrane catalogue
|       |-- utilities/              <name>.dat --- plant-utility catalogue
|       |-- binaryPairs/            NRTL / Wilson / UNIQUAC pair files (arity-2)
|       |-- henrysLaw/              Henry's-law pairs (arity-2)
|       |-- solution/               <solute>-<solvent>.dat --- aqueous TIER (arity-2)
|       |-- electrolyte/            ions.dat + Pitzer/eNRTL catalogues (arity-2)
|       |-- eos/<model>/            model-keyed binaryInteractions (k_ij)
|       `-- unifac/                 group-interaction tables
|-- docs/                          THIS FILE + userGuide.{md,tex,pdf}
|-- src/
|   |-- core/                       Dictionary, SimulationResult, Constants, Types
|   |-- streams/                    ProcessStream
|   |-- thermo/                     Component, ThermoPackage, Database +
|   |                                phase/, vaporPressure/, activityCoefficient/,
|   |                                equationOfState/, heatCapacity/, membrane/
|   |-- materials/                  Material, MaterialRegistry
|   |-- solver/                     NewtonRaphson, NewtonND, Wegstein, StabilityTest
|   |-- unitOperations/             UnitOperation base + per-category folders
|   |-- outerDriver/                OuterDriver base + SweepDriver + FitBinaryPair
|   |-- postProcessing/             PostProcessor base + Sizing/Costing passes
|   |-- propertyOps/                PropertyOperation base (choupoProps)
|   |-- control/                    Controller base + PID/Schedule (choupoCtrl)
|   |-- reporting/                  BalanceMath, EnergyBalanceReport (the ledger)
|   |-- io/                         SolutionWriter --- OpenFOAM-style instants
|   `-- applications/               choupoSolve / choupoBatch / choupoCtrl /
|                                     choupoProps -- one main.cpp per binary
|-- tests/                         regression scripts (bin/runTests)
|-- tutorials/                     end-to-end cases (~190, under
|                                     steady/ batch/ ctrl/ props/ plant/)
|-- gui/                           web GUI (see CLAUDE.md Sec. 12)
|-- LICENSE                        GPL-3.0-or-later
`-- README.md
\end{lstlisting}

\begin{keyidea}
\textbf{Data lives under \file{data/standards/}, write-guarded.}  The engine
\emph{refuses} to write under \file{data/standards/} --- adding a number there
is a deliberate curation act, not a side effect of a run.  WHERE a number
belongs (an intrinsic field in \file{components/<name>.dat}, a pair file in a
catalogue, or an operation's \file{constant/}) is decided by the
\textbf{data-governance doctrine} (\file{docs/ai/data-doctrine.md}): the
canonical reference for every curation question, summarised in
\S\ref{sec:data-cascade}.
\end{keyidea}

\begin{keyidea}
There is \emph{one binary per problem class}, four in all --- they share
the same library under \file{src/\{core,thermo,solver,\dots\}} and differ
only in their \file{applications/<name>/main.cpp}:
\begin{itemize}[leftmargin=*,nosep]
  \item \file{choupoSolve} --- steady-state simulation, $F(\mathbf{x})=0$
        (root-finding);
  \item \file{choupoBatch} --- batch / recipe-driven, per-vessel ODE
        integration + time-triggered events;
  \item \file{choupoCtrl}  --- dynamic continuous + control,
        $d\mathbf{Y}/dt = f(\mathbf{Y},\mathbf{u},t)$ with PID/schedule
        controllers;
  \item \file{choupoProps} --- property evaluation + parameter fitting
        (a distinct problem class: neither root-finding nor integration).
\end{itemize}
We split by \emph{problem class}, never by numerical \emph{strategy} (no
\code{*Foam}-style proliferation): within each binary the strategies
(Newton, Wegstein, RK4, Nelder--Mead, \dots) are run-time selections via
dicts, not compile-time ones.
\end{keyidea}

% =====================================================================
\section{Build system}

\begin{lstlisting}
make all              # release, -O2, build/linux64Gcc/
make MODE=debug all   # -O0 -g3, build/linux64GccDebug/
make print            # show resolved variables
make clean            # rm objs + binary for current MODE
make distclean        # rm entire build/
\end{lstlisting}

Sources are discovered with \code{find src -name '*.cpp'} --- drop a new
\file{.cpp} anywhere under \file{src/} and it picks up automatically.
Header dependencies are auto-tracked via \code{-MMD -MP} (\file{.d} files
alongside each \file{.o}).

Platform string is derived from \code{uname -s -m}. Currently only
\code{linux64Gcc} is exercised; \code{darwin64Clang} and
\code{linux64Clang} should ``just work'' with the same Makefile (modulo
\code{<filesystem>} linking on older toolchains --- see
\S\ref{sec:debug}).

The WASM build (\code{make wasm}) is a sibling target --- see
\file{make/wasm.mk} and \file{CLAUDE.md} \S13 for the Emscripten quirks.

% =====================================================================
\section{Build architecture: shared library, optional modules, case code}
\label{sec:build-arch}

Since the build is \emph{dynamic} and \emph{modular}.  Three forces
drove it: incremental compilation (change one binary, don't rebuild the
engine), a clean extension story (users add unit ops without touching
\file{src/}), and --- load-bearing for the modular architecture ---
making it hard to produce a standalone binary that carries the whole
engine inside.  The model: an engine built as a shared
library, and small binaries that \emph{reference} it.

\begin{keyidea}
A binary never embeds the engine; it links \file{libchoupo.so} at
build time and finds it at run time via \code{rpath}.  A binary copied to
a machine without the library simply does not start.  This is the
same model --- one engine, many small binaries that share it, so
rebuilding a binary never rebuilds the engine.
\end{keyidea}

\subsection{The shared engine library (\file{libchoupo.so})}

\code{make lib} archives every engine object (everything under
\file{src/} except the per-application entry points) into a single shared
object:

\begin{lstlisting}
make lib              # -> build/<plat>/libchoupo.so   ($(CXX) -shared)
\end{lstlisting}

The engine is compiled \code{-fPIC} (position-independent) on Linux.  Each
binary then links \code{-lchoupo} with \code{rpath} set to
\code{\$ORIGIN} --- the binary's own directory --- which is where the
\file{.so} is built; the project-root symlinks resolve to the real binary
location, so \code{\$ORIGIN} still points at the \file{.so}.  The result
is tiny binaries:

\begin{center}
\begin{tabular}{@{}l r l@{}}
\toprule
Artefact & Size & Contains the engine? \\
\midrule
\file{libchoupo.so}             & 1.7 MB & it \emph{is} the engine \\
\file{choupoSolve} (binary)     & 86 KB  & no --- references it \\
\file{choupoCase} (a user's)    & $\sim$100 KB & no --- references it \\
\bottomrule
\end{tabular}
\end{center}

\begin{pitfall}
The \code{win64MinGW} cross-build keeps \emph{static} linking on purpose:
the generated \file{.exe} is self-contained.  The \code{LIB\_DEP} /
\code{LIB\_LINK} Makefile variables carry the platform branch (objects
directly for \code{win64MinGW}, \code{-lchoupo} on Linux).
\end{pitfall}

\subsection{Optional modules --- a clean modular split}

The engine can carry OPTIONAL modules as separate \file{.so} files, so a
slim core (the teaching engine) builds without them.  The membrane module is
Vítor's NF/RO research code --- conceptually distinct from the core and kept
in its own library so a core-only build is possible.  Everything is
GPL-3.0-or-later; the split is modular, not a licence boundary:

\begin{center}
\begin{tabular}{@{}l l l@{}}
\toprule
Library & Contents & Licence \\
\midrule
\file{libchoupo.so}          & core (thermo, solver, unit ops, reports) & GPL-3.0-or-later \\
\file{libchoupoMembrane.so}  & NF/RO membrane (\code{spiralWoundModule}) & GPL-3.0-or-later \\
\bottomrule
\end{tabular}
\end{center}

The mechanism is the same link-time stub-swap used for user code
(\S\ref{sec:case-code}), applied to a module:

\begin{enumerate}[leftmargin=1.6em]
  \item The optional unit op lives under its own folder
    (\file{src/unitOperations/membrane/}); its objects are filtered
    \emph{out} of \code{LIB\_OBJS} and compiled into
    \file{libchoupoMembrane.so} instead.
  \item It is \textbf{not} registered in
    \code{UnitOperation::registerBuiltins()} (the core must not depend on
    it).  Instead a \file{registerMembrane.cpp} in the module defines a
    \code{registerModules()} hook that registers the type via the usual
    explicit factory.
  \item \code{choupoSolve/main.cpp} calls \code{registerModules()} right
    after \code{registerBuiltins()}.  The core binary links an empty
    \file{registerModulesStub.cpp} \emph{only} when the module is absent;
    the Makefile (\code{MEMBRANE\_PRESENT := \$(wildcard
    src/unitOperations/membrane)}) links exactly the \file{.so}
    \emph{or} the stub, never both.
\end{enumerate}

\begin{lstlisting}
# slim core-only build: remove the module source, rebuild
rm -rf src/unitOperations/membrane && make all
# -> only libchoupo.so is produced; the membrane type is unknown:
#    "unknown type 'spiralWoundModule'"
\end{lstlisting}

\textbf{To add another optional module}, do exactly what the membrane
does: a folder under \file{src/}, kept out of \code{registerBuiltins()},
with its own \code{registerModules()} compiled into its own \file{.so}.
(When a \emph{second} module appears, generalise the single
\code{registerModules()} hook into a small list --- still explicit, still
no auto-registration.)

\subsection{Case-local code (\file{case/code/})}
\label{sec:case-code}

A user --- a student, a customer --- may write their own \cpp{} under a
case's \file{code/} folder and have it compiled into a per-case binary.
Build-time only: \textbf{no runtime \code{dlopen}, no macro
self-registration}, faithful to the glass-box credo.

\begin{lstlisting}
case/
  code/
    MyOp.H  MyOp.cpp        # the user's unit op (derives UnitOperation)
    registerUserTypes.cpp   # defines registerUserTypes() -> registerType("myOp",...)
    options                 # OPTIONAL: extra -I / -l / flags (see below)
  system/  constant/
\end{lstlisting}

\paragraph{What \file{code/} can register.}  The primary, tutorialised
use is a new \emph{unit operation}.  But \code{registerUserTypes()} runs
\emph{after} every \code{registerBuiltins()}, so it can register
\textbf{any type with an explicit factory} --- the same hook, the same
pattern:

\begin{center}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
What & via & Status \\
\midrule
unit operation & \code{UnitOperation::registerType} & demonstrated (userOp01/02) \\
thermo model ($\gamma$, EoS, $P^{\mathrm{sat}}$, $c_p$) &
  \code{ActivityModel::registerModel}, \dots & possible, no tutorial yet \\
outer driver (sweep / optim / DesignSpec) & \code{OuterDriver::registerType} & possible \\
report / sizing / costing & \code{Report::registerType}, \dots & possible \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{What the user writes.}  Three files in \file{code/}: the unit
op (\file{MyOp.\{H,cpp\}}, a normal class deriving \code{UnitOperation}
with a \code{solve()} override), and \file{registerUserTypes.cpp}, which
puts the class in the factory:

\begin{lstlisting}
// case/code/registerUserTypes.cpp
void registerUserTypes() {
    Choupo::UnitOperation::registerType(
        "myOp", [] { return std::make_unique<Choupo::MyOp>(); });
}
\end{lstlisting}

\paragraph{How it is compiled.}  \code{bin/buildCode} runs one ordinary
\code{g++} (it prints the exact command --- nothing hidden):

\begin{lstlisting}
g++ -std=c++17 -O2 -I src \             # (1) the engine's headers
    build/<plat>/.../choupoSolve/main.o \  # (2) the engine main() (calls
                                            #     registerUserTypes)
    case/code/registerUserTypes.cpp \   # (3) the user's registration
    case/code/MyOp.cpp \                # (4) the user's unit op
    -Lbuild/<plat> -lchoupo [-lchoupoMembrane] \  # (5) engine.so (+ optional module)
    -Wl,-rpath,build/<plat> \           # (6) where to find the.so at run time
    -o case/code/.build/choupoCase     # (7) the per-case binary (~100 KB)
\end{lstlisting}

\noindent The user's \file{.cpp} are compiled fresh each time (their
\file{.o} are temporaries in \file{/tmp}, dropped after linking --- only
the binary remains).  The binary links the engine \file{.so}
DYNAMICALLY (5): it references the engine, never embeds it.

\paragraph{The trick --- who supplies \code{registerUserTypes()}.}  The
engine's \file{main.o} \emph{calls} \code{registerUserTypes()} (right
after \code{registerBuiltins()}) but does \emph{not} define it --- an
unresolved symbol.  The linker resolves it from exactly one place, and
that place is the build-time choice:

\begin{center}
\begin{tabularx}{\linewidth}{@{}l X l@{}}
\toprule
Binary & \code{registerUserTypes()} from & Result \\
\midrule
stock \code{choupoSolve}  & \file{registerUserTypesStub.cpp} (empty) & registers nothing \\
the case's \code{choupoCase} & \file{case/code/registerUserTypes.cpp} & registers \code{myOp} \\
\bottomrule
\end{tabularx}
\end{center}

\noindent In the \code{buildCode} command above the stub is \emph{not}
linked --- only the user's file is --- so the linker resolves the symbol
with the user's definition.  Plain symbol resolution, no weak symbols, no
\code{dlopen}, no macros.  (Optional modules use the same pattern with
\code{registerModules()}, \S4.2.)

\paragraph{How it is used --- at run time.}
\begin{lstlisting}
choupoCase runs:
  main()
    -> UnitOperation::registerBuiltins()  // cstr, flash, mixer,...
    -> registerModules()                  // optional modules (membrane.so)
    -> registerUserTypes()   // THE USER'S: registry()["myOp"] = constructor
  Flowsheet reads  type myOp;  in flowsheetDict
    -> UnitOperation::New("myOp")         // look up in the std::map
    -> the user's constructor -> object -> the user's solve()
\end{lstlisting}

\noindent The factory is just a \code{std::map<string, function>};
\code{registerType} adds an entry, \code{New} looks one up.  \code{myOp}
sits beside the built-in types --- no special case.

\paragraph{What \code{solve()} receives --- streams and thermo.}  The
op's whole world arrives through two arguments:
\begin{lstlisting}
int solve(const DictPtr& dict, const ThermoPackage& thermo, int verbosity);
\end{lstlisting}

\emph{Streams (the \code{dict}).}  Choupo is sequential-modular: an op
does NOT rummage a global stream registry.  The \code{Flowsheet}
\emph{injects} the unit's inlet stream into the \code{dict} as sub-dicts,
plus the unit's own \code{operation} block:
\begin{lstlisting}
auto feed = dict->subDict("feed");        // F (kmol/s SI), T, P of the inlet
auto comp = dict->subDict("composition"); // mole fractions z_i
auto oper = dict->subDict("operation");   // YOUR parameters (from flowsheetDict)
\end{lstlisting}
You read your inputs, compute, and publish outputs into \code{produced\_};
the Flowsheet binds them to the next units by the \code{outputs (... )}
order.  (For several inlets, the unit declares \code{inputs (a b )} and
reads each.)

\emph{Thermodynamics (the \code{thermo}).}  The \code{ThermoPackage\&}
gives full access to whatever models the case configured --- you do not
need to know \emph{which} ($\gamma$ = NRTL? EoS = SRK?); you call the
method and the selected model answers:
\begin{lstlisting}
thermo.n();  thermo.indexOf("benzene");  thermo.comp(i);  // count, index, Component
thermo.comp(i).vp().Psat_Pa(T);                            // vapour pressure (Antoine/...)
thermo.Kvec(T, P, x, y);                                   // K-values (the case's gamma-phi)
thermo.activity().gamma(T, x);                             // activity coefficients gamma_i
thermo.Hliquid(T, x);  thermo.Hvapour(T, y);              // sensible-datum enthalpies
thermo.H_real(T, P, y);                                    // real gas (H_ig + EoS departure)
thermo.H_stream_formation(T, P, vf, z);                    // elements-datum (carries Hf)
\end{lstlisting}
So a user op can flash, balance energy, or solve an equilibrium with the
case's thermo --- exactly what the built-in CSTR / flash / column do; it
is a first-class citizen, not an add-on.

\begin{pitfall}
\textbf{Transport properties (viscosity, conductivity, diffusivity) do
not exist yet} --- they are on the roadmap as a \code{TransportModel}
with a factory.  When they land they will be reachable through the same
\code{thermo} handle; today a user op has thermodynamics only.
\end{pitfall}

\begin{center}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
Command & What it does \\
\midrule
\code{buildCode <case>} & \emph{compiles only} (shows the exact g++
command, reports compiler errors); the development-loop command \\
\code{runCase <case>}   & \emph{runs} the case --- and if it has a
\file{code/} folder, compiles it first (visibly) via \code{buildCode} \\
\bottomrule
\end{tabularx}
\end{center}

Both step up automatically if invoked from inside the \file{code/}
folder.  External dependencies (a property library, GSL, Eigen, \dots)
go in an optional \file{code/options} file:

\begin{lstlisting}
# case/code/options
INCLUDE = -I/opt/coolprop/include
LIBS    = -L/opt/coolprop/lib -lCoolProp
FLAGS   = -O3
\end{lstlisting}

\begin{pitfall}
User code lives in \file{case/code/}, \textbf{never} in \file{src/}.
Three reasons: isolation (the engine tree stays auditable), authorship /
licence (\file{src/} is GPL-3.0-or-later; the user owns their
\file{code/}), and pedagogy (a clear engine-vs-mine boundary).  Those
external libraries are the \emph{user's} dependencies on the
\emph{user's} code --- the engine itself stays dependency-free.
\end{pitfall}

% =====================================================================
\section{Core types}

\begin{whatyoulllearn}
Four types live in \file{src/core/} and \file{src/streams/} and propagate
through every layer of the simulator: get familiar with them before
touching anything else.
\end{whatyoulllearn}

\subsection{\code{Dictionary}}

hierarchical plain text. Every case dict is parsed into a
\code{Dictionary}. Key operations:

\begin{lstlisting}[language=C++]
auto d = Dictionary::fromFile("system/flowsheetDict");

scalar  T = d->lookupScalar("T");                       // raw
scalar  P = d->lookupScalar("P", Dims::pressure);       // dim-checked
auto    sub = d->subDict("operation");
auto    units = d->lookupDictList("units");
auto    comps = d->lookupWordList("components");

// Path-based mutation --- used by outer drivers
d->setScalarAtPath("units[0].operation.refluxRatio", 3.2);
\end{lstlisting}

The tokenizer treats \code{[} and \code{]} as word chars (so paths
work). Optional \code{FoamFile} headers are skipped silently for
compatibility with files from other tools.

\paragraph{Three input forms for scalars.} The parser
accepts named-unit (\code{P 1 bar;}), the bracket
(\code{A\_w [-1 2 1 0 0] 1.5e-12;}), and raw SI (\code{R 0.5;}).
Named and bracket forms tag the entry with its physical dimensions
\code{[M L T Theta N]} so the dim-checked
\code{lookupScalar(key, expectedDims)} overload can throw a
clear error on mismatch.  Raw entries pass unchecked --- the
caller is asserting the value is canonical SI.

Available named dimensions live under \code{Dims::} in
\file{src/core/Dimensions.H} (\code{pressure}, \code{molarFlow},
\code{massFlow}, \code{temperature}, \code{length}, \code{area},
\code{volume}, \code{concentration}, \code{molarHeatCap},
\code{viscosity}, \code{diffusivity}, \code{permeabilityWater},
\code{massTransferCoeff},...).  Adopt them at the call sites where
the unit op reads its operating parameters.

\subsection{\code{ProcessStream}}

What travels between unit operations (units: SI canonical):

\begin{lstlisting}[language=C++]
struct ProcessStream {
    std::string         name;
    scalar              F;       // total FLUID molar flow [kmol/s]
    scalar              T;       // K
    scalar              P;       // Pa
    std::vector<scalar> z;       // FLUID mole fractions, size = NC
    scalar              vf;      // vapour fraction (VLE outputs)
    std::vector<scalar> s;       // solid molar flow per component [kmol/s]
    ParticleSizeDist    psd;     // size distribution of the solid phase
    std::string         category;// "" = process; non-empty = utility leg
};
\end{lstlisting}

\begin{keyidea}
\code{F}, \code{z} and \code{vf} describe the \emph{fluid} phases only; the
\textbf{solid is carried alongside in \code{s[]}} (kmol/s per component, empty
or all-zero = no solids), with its size distribution in \code{psd}.  A
crystalliser writes its crystals into \code{s[]}; the balance machinery
(\file{src/reporting/BalanceMath.H}) reads \code{s[]} so mass \emph{and} energy
close over the solid too (see \S\ref{sec:energy-datum}).  The \code{category}
tag (populated by \code{utility <name>;}) lets the energy ledger separate a
utility leg --- heat crossing the boundary --- from a process stream.
\end{keyidea}

Mass-basis flow is a derived view, not a stored field --- if you
need kg/s, call \code{F\_mass(stream, thermo)} from
\file{src/streams/StreamMass.H}, which computes
$F\cdot\sum_i z_i\,\mathrm{MW}_i$ on demand.  One source of
truth for the flow (molar); mass cannot drift out of sync.

When you produce streams in your unit-op, fill \emph{all} fields.
At print sites you typically scale by 3600 for kmol/h and by
$10^{-5}$ for bar, or hand the value plus its dimensions to
\code{DisplayUnits::instance().format(v, dims)} which honours the
case's \code{controlDict.units} preferences.

\subsection{\code{SimulationResult}}

What a single pass produces --- input to post-processors and outer
drivers:

\begin{lstlisting}[language=C++]
struct SimulationResult {
    std::map<std::string, ProcessStream>     streams;
    std::map<std::string, std::map<std::string, scalar>> kpis;
    std::map<std::string, EquipmentSizing>   sizings;      // SizingPass
    std::map<std::string, CostBreakdown>     costs;        // CostingPass
    bool                                     converged;
};
\end{lstlisting}

The \code{kpis} map is keyed \code{unitName} $\to$ \code{\{ kpiName $\to$
value \}}. Outer drivers read it; post-processors append to
\code{sizings} / \code{costs}.  (The \code{sizings} field was named
\code{dimensions} until; renamed because \code{Dimensions} is
now the physical-dimensions tag.)

\subsection{\code{ThermoPackage}}

Holds:

\begin{itemize}[leftmargin=*,nosep]
  \item components (vector of \code{Component}),
  \item one or more \code{Phase} instances (vapor / liquid / solid),
  \item convenience getters: \code{nComponents()}, \code{component(i)},
        \code{phase(name)}, \code{Psat(i, T)}, \code{gamma(phase, x, T)},
        etc.
\end{itemize}

Every unit operation receives a \code{const ThermoPackage\&} in
\code{solve()}.

\paragraph{Two enthalpy datums --- and why only one survives.}  The package
exposes enthalpy on two references:

\begin{itemize}[leftmargin=*,nosep]
  \item \code{Hliquid(T,x)} / \code{Hvapour(T,y)} --- the \textbf{sensible}
        datum, anchored at an arbitrary \(T_{\mathrm{ref}}\).  Fine for a
        single-stream heat-up, useless across a reactor: it silently drops the
        heat of reaction.
  \item \code{H\_stream\_formation(T,P,vf,z)} --- the \textbf{elements} datum
        (formation enthalpies at 298.15\,K, the same reference as the ideal-gas
        \(H_{ig}\)).  The reaction enthalpy then \emph{emerges} as the
        difference of formation enthalpies, so a balance closes \emph{through}
        a reactor, a crystalliser, or a heat of solution.
\end{itemize}

\begin{keyidea}
\textbf{One datum, no fallback (\S\ref{sec:energy-datum}).}  The energy-balance
ledger runs on the elements datum \emph{only}.  The former silent
``fall back to the sensible datum'' arm was \textbf{deleted}: a component
missing its \code{standardThermochemistry} now \emph{throws}, naming the component, so a
curation gap is loud, never a quietly wrong closure.
\end{keyidea}

% =====================================================================
\section{The factory pattern --- explicit, no macros}\label{sec:factory}

\begin{whatyoulllearn}
The single registration mechanism used everywhere in the code, and
why we refuse the macro-based alternative.
\end{whatyoulllearn}

Every extensible base class follows this template:

\begin{lstlisting}[language=C++]
class UnitOperation {
public:
    using Factory = std::function<std::unique_ptr<UnitOperation>()>;

    static void registerType(const std::string& name, Factory f);
    static std::unique_ptr<UnitOperation> New(const std::string& type);
    static std::vector<std::string> availableTypes();
    static void registerBuiltins();          // <-- called in main()

    // ---- virtual interface ----
    virtual const std::string& type() const = 0;
    virtual int  solve(const DictPtr&, const ThermoPackage&, int verbosity) = 0;
    virtual std::vector<ProcessStream> producedStreams() const { return {}; }
    virtual std::map<std::string, scalar> kpis() const { return {}; }

protected:
    // helpers...
};
\end{lstlisting}

\code{registerBuiltins()} (in the corresponding \file{.cpp}) is the
single place that \emph{lists every type}:

\begin{lstlisting}[language=C++]
void UnitOperation::registerBuiltins()
{
    auto reg = [](const std::string& n, Factory f) { registerType(n, f); };

    reg("isothermalFlash", []{ return std::make_unique<IsothermalFlash>(); });
    reg("flash",           []{ return std::make_unique<IsothermalFlash>(); });
    reg("adiabaticFlash",  []{ return std::make_unique<AdiabaticFlash>();  });
    reg("bubbleT",         []{ return std::make_unique<BubblePoint>();     });
    reg("dewT",            []{ return std::make_unique<DewPoint>();        });
    reg("cstr",            []{ return std::make_unique<CSTR>();            });
    reg("pfr",             []{ return std::make_unique<PFR>();             });
    reg("heatExchanger",   []{ return std::make_unique<HeatExchanger>();   });
    reg("HX",              []{ return std::make_unique<HeatExchanger>();   });
    reg("mixer",           []{ return std::make_unique<Mixer>();           });
    reg("splitter",        []{ return std::make_unique<Splitter>();        });
    reg("distillationColumn", []{ return std::make_unique<DistillationColumn>(); });
    reg("column",             []{ return std::make_unique<DistillationColumn>(); });
}
\end{lstlisting}

\code{main.cpp} calls each \code{registerBuiltins()} once at startup, in
order. That is the \emph{entire} registration story. No magic, no
surprises.

\begin{pitfall}
Do \emph{not} introduce \code{REGISTER\_TYPE(...)} macros,
anonymous-namespace static instances, or any pattern that runs
registration as a side effect of TU loading. We have rejected this
approach explicitly because the linker can drop unused TUs and the
failure mode is silent.
\end{pitfall}

% =====================================================================
\section{Adding a new unit operation}\label{sec:add-unitop-quick}

\begin{whatyoulllearn}
The five moving parts in outline --- skeleton, implementation, registration,
tutorial, regression.  The full \emph{tim-tim-por-tim-tim} walkthrough, with a
real worked example and the mandatory WASM rebuild, is in
\S\ref{sec:recipe-unitop}; this is the at-a-glance version.
\end{whatyoulllearn}

\subsection{Step 1 --- Skeleton}

Create \file{src/unitOperations/<category>/MyOp.H}:

\begin{lstlisting}[language=C++]
#pragma once
#include "core/Dictionary.H"
#include "thermo/ThermoPackage.H"
#include "unitOperations/UnitOperation.H"

namespace Choupo {

class MyOp : public UnitOperation {
public:
    MyOp() = default;
    const std::string& type() const override { return type_; }

    int solve(const DictPtr&         dict,
              const ThermoPackage&   thermo,
              int                    verbosity) override;

    std::vector<ProcessStream> producedStreams() const override { return produced_; }
    std::map<std::string, scalar> kpis() const override { return kpis_; }

private:
    inline static const std::string type_ = "myOp";
    std::vector<ProcessStream>      produced_;
    std::map<std::string, scalar>   kpis_;
};

} // namespace Choupo
\end{lstlisting}

\subsection{Step 2 --- Implementation}

\file{src/unitOperations/<category>/MyOp.cpp} --- implement
\code{solve()}. Boilerplate to follow:

\begin{lstlisting}[language=C++]
int MyOp::solve(const DictPtr& dict, const ThermoPackage& thermo, int verbosity)
{
    // 1. The Flowsheet INJECTS the inlet + your params as sub-dicts of `dict`
    //    (sequential-modular: no global stream registry to rummage).
    auto feed = dict->subDict("feed");          // F (kmol/s SI), Tfeed, Pfeed
    auto comp = dict->subDict("composition");   // mole fractions z_i
    auto op   = dict->subDict("operation");     // YOUR parameters
    const scalar F     = feed->lookupScalar("F",     Dims::molarFlow);
    const scalar Tfeed = feed->lookupScalar("Tfeed", Dims::temperature);
    const scalar T     = op->lookupScalar("T",       Dims::temperature);

    // 2. Echo inputs (verbosity >= 3 --- the pedagogical default)
    if (verbosity >= 3)
        std::cout << "  [MyOp] T = " << T << " K, F_in = "
                  << (F * 3600.0) << " kmol/h\n";   // print in kmol/h, store SI

    // 3. Solve the model --- use solver/ classes for Newton/Wegstein.
    //    newton1D(f, dfdx, x0, NROptions) -> NRResult { root, iterations, converged }
    solver::NROptions nro;  nro.lower = 200.0;  nro.upper = 700.0;
    auto r = solver::newton1D([&](scalar x){ /* g(x) */ return 0.0; },
                              [&](scalar x){ /* g'(x) */ return 1.0; },
                              /*x0=*/Tfeed, nro);

    // 4. Build produced streams into produced_ (order = your contract;
    //    the Flowsheet binds them to names via `outputs (... )`)
    produced_.clear();
    ProcessStream out;
    out.name = "myOp_out";
    out.F = F;  out.T = T;  out.P = feed->lookupScalar("Pfeed", Dims::pressure);
    out.z.resize(thermo.n());                    // fill ALL fields
    produced_.push_back(out);

    // 5. Populate KPIs (they feed sizing / costing / sensitivity)
    kpis_.clear();
    kpis_["T_out"] = T;
    kpis_["F"]     = F;

    return 0;     // 0 == converged
}
\end{lstlisting}

\subsection{Step 3 --- Register}

Edit \file{src/unitOperations/UnitOperation.cpp}:

\begin{lstlisting}[language=C++]
#include "<category>/MyOp.H"
//...
void UnitOperation::registerBuiltins() {
    //...
    reg("myOp", []{ return std::make_unique<MyOp>(); });
}
\end{lstlisting}

\subsection{Step 4 --- Tutorial}

Create \file{tutorials/myOp01\_quick\_smoke/} with
\file{system/controlDict}, \file{system/flowsheetDict}, and
\file{constant/thermoPhysPropDict}. Keep the case trivial so it doubles as a
regression test.

\subsection{Step 5 --- Build, WASM, regression}

\begin{lstlisting}
make all                 # recursive find picks the new .cpp up automatically
make wasm-gui            # MANDATORY: the GUI's WASM is a SEPARATE build (see below)
bin/runTests             # NaN/inf guard on every case + golden-master KPIs
\end{lstlisting}

\begin{pitfall}
\textbf{Rebuild the WASM or the browser will not see your type.}  The WASM
solver is a separate compile from the native binary.  Symptom of a stale
\file{.wasm}: the native run works, but the GUI errors
\code{UnitOperation::New: unknown type 'myOp'}.  Fix with \code{make wasm-gui}
(force a clean one with \code{make wasm-clean \&\& make wasm} if it says
``Nothing to be done'').  Never validate with a bare
\code{choupoSolve > /dev/null \&\& PASS} loop --- a solver can return NaN with
exit code 0; \code{bin/runTests} is the only honest gate (\S\ref{sec:regression}).
\end{pitfall}

All previously-passing tutorials \emph{must} still pass.

% =====================================================================
\section{Adding a new thermodynamic model}\label{sec:add-thermo-model}

Same explicit-factory spirit as everywhere else, with \code{ActivityModel} /
\code{EquationOfState} / \code{VaporPressureModel} / \code{HeatCapacityModel} /
\code{Phase} as the base.  The one difference from \code{UnitOperation} is
\textbf{the factory takes construction arguments} (a thermo model needs the
configuring sub-dict, and usually the component list), so its \code{Factory}
signature is not the nullary unit-op one --- match the base you are extending:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
Base & \code{Factory} / register call \\
\midrule
\code{VaporPressureModel} & \code{(const DictPtr\&)};
   \code{registerModel(name, f)} \\
\code{HeatCapacityModel}   & \code{(const DictPtr\&)};
   \code{registerModel(name, f)} \\
\code{ActivityModel}       & \code{(const DictPtr\&, const std::vector<std::string>\&)};
   \code{registerModel(name, f)} \\
\code{EquationOfState}     & \code{(const DictPtr\&, const std::vector<Component>\&)};
   \code{registerModel(name, f)} \\
\bottomrule
\end{tabularx}
\end{center}

\noindent The detailed, copy-pasteable recipe --- with the validation-first
AAD step and the EoS data-doctrine path --- is in
\S\ref{sec:recipe-property}.  The short version, adding UNIQUAC as an activity
model:

\begin{enumerate}[leftmargin=*]
  \item \file{src/thermo/activityCoefficient/UNIQUAC.\{H,cpp\}}
        inheriting \code{ActivityModel}. Override the pure virtuals
        \code{gamma(scalar T\_K, const sVector\& x)}, \code{n()},
        \code{modelName()}; a constructor \code{UNIQUAC(const DictPtr\&, const
        std::vector<std::string>\&)} reads the pair parameters.
  \item In \file{src/thermo/activityCoefficient/ActivityModel.cpp}, add the
        \code{\#include} and \emph{one} line in \code{registerBuiltins()}:
\begin{lstlisting}[language=C++]
#include "UNIQUAC.H"
//...
void ActivityModel::registerBuiltins() {
    registerModel("ideal",  [](const DictPtr&, const std::vector<std::string>& n)
        -> std::unique_ptr<ActivityModel>
        { return std::make_unique<IdealSolution>(n.size()); });
    registerModel("NRTL",   [](const DictPtr& d, const std::vector<std::string>& n)
        -> std::unique_ptr<ActivityModel>
        { return std::make_unique<NRTL>(d, n); });
    registerModel("Wilson", [](const DictPtr& d, const std::vector<std::string>& n)
        -> std::unique_ptr<ActivityModel>
        { return std::make_unique<Wilson>(d, n); });
    registerModel("UNIQUAC",[](const DictPtr& d, const std::vector<std::string>& n)   // new
        -> std::unique_ptr<ActivityModel>
        { return std::make_unique<UNIQUAC>(d, n); });
}
\end{lstlisting}
  \item Use it in a tutorial's \file{thermoPhysPropDict}
        (inside \code{equilibrium.liquid}):
\begin{lstlisting}[language=dict]
activityModel { model UNIQUAC; binaryParameters {... } }
\end{lstlisting}
\end{enumerate}

% =====================================================================
\section{Property operations (\code{choupoProps})}\label{sec:propops}

\code{choupoProps} runs \emph{property operations} rather than a flowsheet.
Each inherits \code{PropertyOperation} (same explicit factory as everywhere
else; \code{registerBuiltins()} is called from
\file{src/propertyOps/PropertyOperation.cpp}). An op reads a \code{state}
(\(T\), \(P\), composition), a \code{properties} list and an \code{output}
block, and writes a CSV the GUI auto-plots. Built-ins: \code{propertyPoint},
\code{propertyScan1D}, \code{propertyScan2D}, \code{propertyScanTernary},
\code{fitParameters}, \code{vleConsistency}, \code{estimateComponent}. Property
keywords live in \file{PropertyEvaluator.cpp}: per-component
\code{Psat\_<c>}, \code{gamma\_<c>}, \code{y\_eq\_<c>}, \code{Cp\_liquid\_<c>}
(each guarded --- e.g.\ \code{Psat} on a component with no vapour-pressure
model throws a clear error instead of dereferencing null), and mixture scalars
\code{Z}, \code{v\_molar}, \code{H\_real}, \code{S\_real}, transport, etc.
\code{PropertyScan1D} wraps each evaluation in \code{try/catch}, writing
\code{nan} on failure so one undefined component never aborts the whole scan.

\paragraph{\code{propertyScanTernary} --- ternary phase diagrams by reuse.}
Walks the composition simplex (\(x_1+x_2+x_3=1\)) at fixed \((T,P)\) and, at
every node, calls the SAME solver the \code{isothermalFlash} unit op calls ---
so the diagram cannot disagree with a single flash. Three \code{mode}s:
\begin{itemize}[leftmargin=*]
  \item \code{bubbleT} --- a boiling-temperature \emph{surface}: per node
        \code{BubblePoint::compute}; needs only the three vapour pressures
        (ideal binaries are fine), no liquid-liquid data.
  \item \code{lle} (default) --- a \emph{solubility} map: the liquid--liquid
        flash (\code{PhaseSet::LL}) gives a clean binodal + tie-lines, no
        spurious subcooled vapour (the extraction/decanter view).
  \item \code{vlle} --- the full vapour\,+\,two-liquid classification
        (\code{PhaseSet::VLLE}).
\end{itemize}
Each simplex node is independent, so the scan is embarrassingly parallel: the
optional \code{shard \{ k; n; \}} block computes only rows with \(i \bmod n = k\)
(round-robin, balanced). The GUI fans \(N=\)\,cores\(-2\) shards across \(N\)
WASM workers and merges the partial CSVs (the union of shards reproduces the
full grid exactly). No physics is reimplemented --- the op is a triangular loop
+ a classifier + a CSV writer over \code{solveCore}/\code{BubblePoint}.

\paragraph{Adding a property op.} Mirror the recipe of
\S\ref{sec:factory}: implement \file{src/propertyOps/MyOp.\{H,cpp\}} inheriting
\code{PropertyOperation} (override \code{type()}, \code{run(dict,thermo,verbosity)},
optionally \code{diagnostics()}), then one \code{reg("myOp", ...)} line in
\code{PropertyOperation::registerBuiltins()}. Rebuild the native binary AND the
WASM (\code{make wasm-gui}) or the browser will not see the new op.

% =====================================================================
\section{The thermophysical property architecture}\label{sec:property-architecture}

Property data in Choupo is \emph{declarative systems with OpenFOAM-like
explicit files}.  A case does not hand the solver a bag of numbers; it
\emph{declares} its thermophysical system --- \file{constant/thermoPhysPropDict}
(\code{recordType thermophysicalPropertySystem; schemaVersion 2;}), one
record naming the components, the equilibrium formulation, the model in
each phase slot, and the parameter files the run needs.  The one-sentence
contract:

\begin{quote}\itshape
\code{thermoPhysPropDict} declares, \code{ThermoPackageBuilder} assembles,
\code{ThermoPackage} computes; curation produces \code{.dat}, runtime does not
estimate.
\end{quote}

\subsection{The flow}

\begin{lstlisting}
thermoPhysPropDict        the case DECLARES this (declarative)
     |
ThermoPackageBuilder      runtime ASSEMBLY: loads + assembles, never estimates
     |                    (src/thermo/ThermoPackageBuilder.cpp)
ThermoPackage             the mounted runtime COMPUTE object (unchanged)
     |
unit operations           consume the ThermoPackage, as before
\end{lstlisting}

\emph{Terminology (load-bearing).}  The runtime assembly step is the
\code{ThermoPackageBuilder} --- it \textbf{loads}.  It is \emph{never} a
``resolver'': \emph{resolver} is reserved for \textbf{curation-time} estimation
(the Props Guide).  Builder ${}={}$ runtime, loads; resolver ${}={}$ curation,
may estimate; the two never mix.

\subsection{The two axes}

A substance never \emph{is} molecular or electrolyte --- the same NaCl plays
different roles, and the role is chosen by the package a case selects, never
stored on the substance.  Two orthogonal choices define the role:

\begin{description}
\item[Representation] (the \emph{package} activates it): lumped $\to$ complete
  dissociation $\to$ partial / ion-pair $\to$ multi-ion.  The complete,
  $K$-free split is the \code{dissociatesTo} ion stoichiometry on the
  component --- a formula-like \emph{identity}; the moment a split carries an
  equilibrium constant (an ion pair, a hydrate) it is \emph{chemistry} and
  lives under \code{chemistry/}.
\item[Reference] (the \emph{method} declares it): which datum the species
  anchor to --- solid, pure-liquid Raoult, aqueous infinite-dilution, or
  fused-salt.  \code{species/} records are medium-agnostic (one Na$^{+}$ file);
  the thermo data inside are medium-tagged blocks (\code{aqueousThermo}).
\end{description}

\subsection{The seven data homes}

\begin{lstlisting}
data/standards/
  components/<name>.dat        identity (name, MW, nonvolatile)
                               + dissociatesTo { Na 1; Cl 1; }  [salts]
  species/<name>.dat           one modelSpecies file per species:
                               formula + charge + medium-tagged thermo
                               (charge 0 for O2/N2(aq); never fed to a flowsheet)
  phases/solid/<phase>.dat     crystalline phase (rho_p, k_v)
  chemistry/                   REAL equilibria (stoichiometry + K + dH):
    salts/                       calorimetric dissolution anchors
    mineralSolubility/           PHREEQC mineral logK/dH (speciation engine)
    aqueousSpeciation/  gasLiquid/  ionExchange/
  parameters/
    Pitzer/{pairs,lambda,psi}/...             Pitzer binaries + ternaries
    eNRTL/<c>-<a>.dat                         eNRTL tau pairs
    NRTL/ UNIQUAC/ Wilson/ Henry/ SRK/ ...    molecular pair catalogues
  (the system is declared INLINE in constant/thermoPhysPropDict -- THE
   CENTRE; there is no shared package catalogue, and the old per-model
   methods/<name>.dat ceremony records were retired)
\end{lstlisting}

\code{components/} stays \emph{flat} (the loader resolves \code{<name>.dat} by
exact name, O(1)).  There is deliberately \textbf{no} second component record
per salt (no \code{apparent/} twin --- one component, one file), no
user-facing ``true species'' vocabulary (an ion is not \emph{more real} than
the salt; it is the speciation a model postulates --- so: \emph{model
species}), and no \code{propertySets/} kind (it had zero readers and was
removed rather than kept as ceremony).  A minimal electrolyte system
(active chemistry beside it in \file{constant/chemistryDict}):

\begin{lstlisting}
recordType    thermophysicalPropertySystem;
schemaVersion 2;
components ( water NaCl );
equilibrium
{
    formulation electrolyteGammaPhi;
    aqueous
    {
        solvent            water;
        apparentComponents ( NaCl );
        activityModel { model Pitzer; }
        compositionBasis molality;
    }
    vapour { fugacityModel idealGas; }
}
\end{lstlisting}

The solid equilibrium activates through the chemistry selection (the
\(K_{sp}\) anchor comes from the component's \code{solidPhases}); the Pitzer
pair is resolved by cation--anion name from the parameter catalogue.

\subsection{The builder contract}

\code{ThermoPackageBuilder::build} dispatches on the declared
\code{equilibrium.formulation}.  For an electrolyte formulation it (1)~reads the
component list; (2)~identifies the ONE salt as the component whose standards
record carries \code{dissociatesTo}, and takes its identity (MW, role) from
\code{components/<salt>.dat}; (3)~classifies the ions by charge sign from
each ion's \code{species/<name>.dat}; (4)~loads the crystal (\(\rho_p, k_v\)) from
\code{phases/solid/} and the saturation anchor from \code{chemistry/salts/},
both by the phase name the package's \code{chemistry.salts} declares;
(5)~assembles the Pitzer/eNRTL kernel through the same
\file{SaltFromCatalogue} helpers the legacy path used --- byte-identical by
construction; (6)~mounts the \code{ThermoPackage}.  \textbf{It loads and
assembles; it does not estimate} --- a missing record fails clear:

\begin{lstlisting}
FATAL: electrolyteGammaPhi selected the Pitzer activity model,
       but required parameter pair Na-Cl was not found.
\end{lstlisting}

For a molecular formulation (\code{gammaPhi} with NRTL) the same entry
point reads the binary pair from \code{parameters/NRTL/} and mounts the
classical NRTL package --- one stack, electrolyte and molecular alike.  The
reference conventions are the \code{standardState} keys on the phase/group
blocks (\code{pureLiquid}, \code{infiniteDilution}); the old per-model
\code{methods/<name>.dat} ceremony records were retired once the dispatch
became fully native.

\subsection{Status}

The legacy salt-level electrolyte front-end is fully retired: the
\code{electrolyte\{\}} component block is gone from every salt \code{.dat}
(ion identity now derives from \code{dissociatesTo}; dissolution data live in
\code{chemistry/salts/}), \code{ElectrolyteActivity::configure()} and the
\code{model pitzer;}/\code{model eNRTL;} factory registrations are deleted,
and the electrolyte tutorials declare \code{electrolyteGammaPhi} systems.
A molecular case is the same record degenerate --- a \code{gammaPhi} block,
no dissociation, no chemistry; lightweight is not legacy.  Two
electrolyte front-ends remain deliberately separate: \emph{multi-ion
speciation} (\code{pitzerHMW}/\code{davies} via a \code{propsDict}
\code{speciate} op) and the flat membrane/DSPM-DE data path.  The
mixed-solvent, segment-based eNRTL of Chen \& Song (2004) ships as
\code{ENRTLMixedSolvent} with a primary-data validation tutorial
(\file{tutorials/props/electrolyte/enrtl\_mixed\_nacl\_ethanol\_esteso}).

% =====================================================================
\section{Adding components, materials, reactions}\label{sec:add-data}

These are \emph{data-driven} --- no \cpp{} changes needed.  WHERE each number
belongs is not a matter of taste: it is decided by the data-governance doctrine
(\file{docs/ai/data-doctrine.md}, summarised in \S\ref{sec:data-cascade}).
Read that before adding or moving a value.

\subsection{Components}

Choupo's component data cascades through tiers, lowest precedence to highest:
\code{local < standard < case < sector < unit} (\code{local} is the private,
gitignored \file{data/local/} tier; every use is announced
\code{[local] UNVERIFIED}).

\begin{itemize}[leftmargin=*]
  \item \file{data/standards/components/<name>.dat} --- the
        \textbf{curated central catalogue} shipped with the simulator.  Carries
        only \textbf{arity-1 intrinsic} data: identity, \code{critical},
        \code{gasIdeal}, \code{liquidPure}, \code{solid}, \code{transport}, the
        \code{standardThermochemistry} datum, and any \code{eosParameters\{\}}
        model-keyed block.  Treat it as read-only --- the engine
        \emph{refuses} to write here; a change is a reviewed curation act.  A
        \textbf{pair} number (NRTL \(\tau\), Henry, \(k_{ij}\), a heat of
        solution) is \textbf{never} copied into the \code{.dat}; it lives in a
        catalogue, referenced by name.
  \item \file{<case>/constant/components/<name>.dat} (and the same at
        \file{<sector>/} and \file{<unit>/}) --- \textbf{per-scope overlays}
        for sample-specific refinements (a particular powder's sorption
        isotherm, a laboratory measurement).
\end{itemize}

\begin{keyidea}
\textbf{The overlay merges block-by-block, NOT field-by-field.}  An overlay
carrying \code{solid \{ rho\_p 1610; \}} replaces the \emph{entire}
\code{solid\{\}} block --- the standard's \code{k\_v} is \emph{lost}, not
deep-merged.  A reference-state block is the atomic unit of physical meaning;
mixing a sample's \code{rho\_p} with the catalogue's \code{k\_v} would be a
silent hidden hybrid.  \textbf{The curator must copy the whole block they
refine.}  (This corrects the pre-doctrine ``field-by-field'' behaviour.)
\end{keyidea}

The file format is identical in both tiers:

\begin{lstlisting}[language=dict]
component <name>;
mw     78.11;            // g/mol
Tc    562.05;            // K
Pc     48.95;            // bar
omega  0.2103;
Tb    353.24;            // K
Hvap_Tb  30720.0;        // J/mol at Tb
Vliq   89.4e-6;          // m^3/mol (Wilson)

vaporPressure
{
    model    Antoine;
    A        4.01814;
    B        1203.835;
    C       -53.226;
}

idealGasHeatCapacity
{
    model       polynomial;
    coeffs      ( -31.368  0.4746  -3.107e-4  8.524e-8 );
    Tmin        200.0;
    Tmax        1500.0;
}

liquidHeatCapacity
{
    model       polynomial;
    coeffs      ( 162.94  -0.3493  9.51e-4  0.0 );
    Tmin        278.0;
    Tmax        500.0;
}

standardThermochemistry
{
    dHf_298     82930.0;        // J/mol  (NIST ideal-gas value)
    s_298         269.20;       // J/(mol*K)  (third-law absolute)
    phase       gas;            // phase in which dHf_298 / s_298 are tabulated
}
\end{lstlisting}

Source the values from DIPPR / Reid-Prausnitz-Poling and \emph{cite the
source in a top-of-file comment} so they stay auditable.

\paragraph{The \code{phase} field (mandatory in \code{standardThermochemistry}).}
The reference state for formation enthalpy is documented in the Theory
Guide (chapter on the elements datum).  The \code{phase} keyword tells
the solver in \emph{which phase} \code{dHf\_298} is tabulated, so
\code{Component::h\_formation(T, targetPhase)} can build the right
integration path through Hvap / Hfus jumps and the matching $c_p$ leg.
Three values are allowed:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}llX@{}}
\toprule
\code{phase} & required data & typical species \\
\midrule
\code{gas}    & \code{idealGasHeatCapacity}, plus \code{HvapTb}+Tc+Tb if the
                liquid leg of any stream is needed
              & \emph{the convention --- all 45 standard volatiles + radicals;
                 follows NIST / JANAF} \\
\code{liquid} & \code{liquidHeatCapacity}; \code{HvapTb}+Tc+Tb if any
                stream goes through the gas leg
              & reserved for compounds whose Hf is published in the
                condensed datum (rare) \\
\code{solid}  & \code{liquidHeatCapacity} (used as the dissolved-solute
                approximation $c_p^{\mathrm{solid}}\!\approx\!c_p^{\mathrm{liq}}$)
              & crystalline non-volatiles whose Hf cannot be referenced to
                gas because the compound never vaporises (sucrose) \\
\bottomrule
\end{tabularx}
\end{center}

The default is \code{gas} for backward compatibility with the earliest
component files, but \textbf{new components should always declare it
explicitly} so the datum is unambiguous at the source of truth.  A
\code{solid} or \code{liquid} declaration without the matching $c_p$
model throws a clear error from \code{Component::h\_formation} the
first time a stream containing the component is evaluated.

\begin{pitfall}
\textbf{A missing \code{standardThermochemistry} is now fatal, not papered over.}  The
energy ledger runs on the elements datum \emph{only} (\S\ref{sec:energy-datum});
the old ``fall back to the sensible datum with a note'' path was deleted.  A
component that appears in a stream the balance touches but carries no formation
datum throws, naming the component --- a loud curation gap, never a quietly
wrong closure.  Curate the eleven historical gaps (or the new component's
datum) before relying on the balance.
\end{pitfall}

Add the new component name to the case's \code{thermoPhysPropDict}
\code{components
(...)} list (or to a per-case \file{constant/components/<name>.dat}
overlay if the data is sample-specific) and the new entry is picked up
automatically --- no recompile needed.

\subsection{Materials}

\file{data/materials/<name>.dat}:

\begin{lstlisting}[language=dict]
material <name>;
rho          7900.0;     // kg/m^3
F_M          1.0;        // Guthrie material factor
sigma_y      138.0e6;    // Pa
maxT         700.0;      // K
maxP         50.0;       // bar
\end{lstlisting}

\code{MaterialRegistry::loadFrom()} slurps the directory on simulator
start. No registration step needed.

\subsection{Utilities}

\file{data/standards/utilities/<name>.dat} catalogues a plant utility
(steam header, cooling-water loop, hot-oil loop, refrigeration brine).
The same registry pattern as materials and membranes: the
\code{UtilityCatalogue::loadFrom(dataRoot)} call in each main slurps
the directory at startup; the stream parser then resolves
\code{utility <name>;} keys against it.

\begin{lstlisting}[language=dict]
utility     steamLP;
tier        heating;            // heating | cooling
mechanism   condensation;       // condensation | sensible | evaporation

components  ( water );
state       saturatedVapour;
P           2.5 bar;
T_in        400.3 K;
T_out       400.3 K;             // condensate same T as steam

dutyPerKg   2186000.0;           // J/kg delivered  (Hvap_water at T_sat)
cost        12.0;                // EUR/GJ delivered
costYear    2022;
description "Low-pressure saturated steam header";
\end{lstlisting}

To add a new utility, drop a new \code{.dat} into the directory --- no
\cpp{} change.  When the utility loop carries a non-water fluid (oils,
salts, brines), the \code{components} list points at a corresponding
entry under \file{data/standards/components/}; the catalogue ships
\code{dowthermA}, \code{hitecSalt} and \code{propyleneGlycol30}
pseudo-components for that purpose.  See the user guide for the
case-author syntax \code{utility <name>;} and the \code{utilities}
report that aggregates kg/h, MW, and EUR/h per category.

\subsection{Reactions}

In a case's \file{constant/reactions}:

\begin{lstlisting}[language=dict]
reactions
{
    ethanol_to_water_test
    {
        stoichiometry
        {
            ethanol      -1.0;
            water        +1.0;
        }
        kinetics
        {
            type        powerLaw;
            A           1.0e6;        // s^-1
            E           80000.0;      // J/mol
            orders      { ethanol 1.0; }
        }
    }
}
\end{lstlisting}

Referenced from a CSTR / PFR by name
(\code{reaction ethanol\_to\_water\_test;}).

\begin{keyidea}
\textbf{The reactions library cascades UP, like the thermophysical
system.}  When a
branch of a fractal plant (\S\ref{sec:fractal}) is run \emph{standalone}, its
named-reaction library resolves by walking up the parent-folder chain (capped
at six levels) exactly as the thermo package does, and \textbf{announces its
origin} --- \code{reactions library: loaded --- LOCAL} vs \code{inherited from
<parent>/...}.  Before this symmetry was restored, a sub-sector run alone died
with ``reaction not in constant/reactions'' because only the local folder was
consulted: the two cascades were asymmetric.  No-silent-crutch: the cascade is
loud.
\end{keyidea}

% =====================================================================
\section{Adding an outer driver / post-processor}

\subsection{Outer driver}

\file{src/outerDriver/MyDriver.\{H,cpp\}} inherits \code{OuterDriver}.
Override:

\begin{lstlisting}[language=C++]
class MyDriver : public OuterDriver {
public:
    void readFromDict(const DictPtr&) override;
    int  run() override;       // calls simulator_(flowsheetDict_) N times
};
\end{lstlisting}

The simulator functor and dict templates are injected by \file{main.cpp}:

\begin{lstlisting}[language=C++]
auto driver = OuterDriver::New(outerDict);
driver->setSimulator     (simulate);          // SimulationResult(DictPtr)
driver->setFlowsheetDict (flowsheetDict);     // mutable flowsheet template
driver->setThermoDict    (thermoDict);        // mutable thermoPhysPropDict
driver->run();
\end{lstlisting}

The simulator closure captures the thermo dict by \code{shared\_ptr},
so any in-place \code{setScalarAtPath()} the driver makes on
\code{thermoDict\_} is picked up on the next \code{simulator\_()} call.

Inside \code{run()}, you typically:

\begin{enumerate}[leftmargin=*,nosep]
  \item Mutate \code{flowsheetDict\_->setScalarAtPath(...)} to change a
        sweep variable, OR \code{thermoDict\_->setScalarAtPath(...)} to
        change a model parameter (e.g.\ NRTL \code{a\_ij}).
  \item Call \code{simulator\_(flowsheetDict\_)} $\to$ get a
        \code{SimulationResult}.
  \item Extract KPIs / stream fields, accumulate residuals / responses.
  \item Decide next step (LM / Nelder-Mead / GA / \dots).
  \item Eventually report and return $0$ (converged) or $1$ (failed).
\end{enumerate}

Two concrete examples to read first:
\begin{itemize}[leftmargin=*,nosep]
  \item \file{src/outerDriver/SweepDriver.\{H,cpp\}} ($\sim 150$ LOC)
        --- 1-D sensitivity sweep.
  \item \file{src/outerDriver/FitBinaryPair.\{H,cpp\}} ($\sim 300$ LOC)
        --- Marquardt LM with FD Jacobian, mutates both flowsheet
        (composition) and thermo (pair params).
\end{itemize}

Register your driver in \code{OuterDriver::registerBuiltins()}.

\subsection{Post-processor}

\file{src/postProcessing/MyPass.\{H,cpp\}} inherits \code{PostProcessor}.
Single method:

\begin{lstlisting}[language=C++]
int run(SimulationResult&) override;
\end{lstlisting}

Append to \code{result.sizings} / \code{result.costs}, write CSVs,
print to stdout. Register in \code{PostProcessor::registerBuiltins()}.

\code{PostProcessor::buildChain(postDict)} walks \code{postDict}
top-level keys and instantiates one pass per recognised key. Add yours
to the chain-builder in \file{PostProcessor.cpp}.

% =====================================================================
\section{Adding a sizing model}

\code{EquipmentSize} is a sub-factory used by \code{SizingPass}. Add
\file{src/postProcessing/sizing/MyEquip.\{H,cpp\}} inheriting
\code{EquipmentSize}:

\begin{lstlisting}[language=C++]
class MyEquip : public EquipmentSize {
public:
    EquipmentSizing size(const std::string& unitName,
                         const SimulationResult& result,
                         const Material&  material,
                         const DictPtr&   designRules) override;
};
\end{lstlisting}

Use \code{result.streams.at(...)} and
\code{result.kpis.at(unitName).at(...)} to get the process data, plus
\code{designRules} for the design specs (pressure design, $F_M$, joint
efficiency, \dots). Fill \code{EquipmentSizing}:

\begin{lstlisting}[language=C++]
EquipmentSizing d;
d.unitName       = unitName;
d.equipmentType  = "myEquip";
d.material       = material.name;
d.values["V_R"]    =...;       // canonical size
d.values["t_wall"] =...;       // m
d.values["weight"] =...;       // kg
return d;
\end{lstlisting}

Register in \code{EquipmentSize::registerBuiltins()}.

The \code{Guthrie} costing model reads \code{d.equipmentType} to
dispatch to the right correlation --- so if your new equipment needs
new cost coefficients, edit
\file{src/postProcessing/costing/Guthrie.cpp} accordingly.

% =====================================================================
\section{Coding conventions}

\subsection{Style}

\begin{itemize}[leftmargin=*]
  \item \cpp{}~17. \code{std::variant}, \code{std::optional},
        \code{std::unique\_ptr}, \code{std::filesystem} --- yes.
        No raw \code{new} / \code{delete}. No \code{goto}.
  \item \textbf{Brace style:} Allman / open-brace on a new line for
        function definitions; Egyptian (\code{if (x) \{}) for control flow.
  \item \textbf{Indent:} 4 spaces. No tabs anywhere.
  \item \textbf{Naming:} \code{snake\_case} for locals, \code{camelCase}
        for methods, \code{PascalCase} for types, \code{UPPER\_SNAKE}
        only for macros (we use exactly one: include guards via
        \code{\#pragma once}).
  \item Compile cleanly with \code{-Wall -Wextra -Wpedantic} and fix
        anything that fires.
\end{itemize}

\subsection{Includes}

\begin{itemize}[leftmargin=*,nosep]
  \item Headers compile in isolation --- every \file{.H} includes what
        it needs.
  \item Forward-declare aggressively in headers; pull full definitions
        only in \file{.cpp}.
  \item The build does not use precompiled headers.
\end{itemize}

\subsection{Errors}

\begin{itemize}[leftmargin=*,nosep]
  \item User-facing: throw \code{std::runtime\_error} with a message a
        student can understand. Don't \code{assert} on user-data
        invariants.
  \item Programmer-facing (impossible state): \code{assert(...)} is fine.
  \item Never silently swallow errors. If a sub-step fails, surface it.
\end{itemize}

\subsection{Comments}

\begin{itemize}[leftmargin=*,nosep]
  \item Default to writing \emph{no} comments. Identifiers should be
        obvious.
  \item Equations and references \emph{deserve} a comment. Cite the
        source (Reid-Prausnitz-Poling page, DIPPR ID, Henley-Seader
        chapter).
  \item Top-of-file banner is preserved on every source --- do not
        delete.
\end{itemize}

\subsection{Header banner}

Every \file{.H} and \file{.cpp} starts with:

\begin{lstlisting}
/*---------------------------------------------------------------------------*\
|  Choupo                                                                     |
|                                                                             |
|  Open-source chemical process simulator                                     |
|                                                                             |
|  Copyright (C) 2026 Vítor Geraldes                                          |
|  SPDX-License-Identifier: GPL-3.0-or-later |
-------------------------------------------------------------------------------
Class
    Choupo::<ClassName>

Description
    <one-paragraph description>
\*---------------------------------------------------------------------------*/
\end{lstlisting}

Don't omit this on new files --- it doubles as licence notice.

% =====================================================================
\section{Regression testing}\label{sec:regression}

Regression is run with \code{bin/runTests}.  Run it before every
change set; all previously-passing tutorials must still pass.

\begin{lstlisting}
bin/runTests                 # every tutorial (4 binaries + buildCode cases)
bin/runTests <case>          # one case
bin/runTests --record <case> # refresh that case's golden-master `expected`
\end{lstlisting}

\code{runTests} does two checks, both stronger than a plain exit-code
loop:

\begin{enumerate}[leftmargin=1.6em]
  \item \textbf{NaN / inf guard --- every case.}  The structured result
    block is scanned for NaN or inf values; a hit FAILS the case.  This
    is the check a bare \code{binary > /dev/null \&\& PASS} loop
    \emph{cannot} do: a solver can return NaN with exit code 0, and the
    old loop reported it green.  That exact trap once hid a broken
    \file{membrane01} (NaN KPIs, exit 0) after a pressure-unit refactor.
  \item \textbf{Reference KPIs --- cases with an \file{expected} file.}
    Each line pins a KPI or stream value to its golden-master value with
    a relative tolerance; any deviation FAILS, printing
    \code{got X expected Y}.
\end{enumerate}

\begin{keyidea}
The \file{expected} file lives in the case directory.  Format
(\code{\#} comments allowed):
\begin{lstlisting}
# kind   name      key         expected        reltol
kpi      flash01   V_over_F    0.30398357314   1e-6
stream   vapor     F           0.00844398814   1e-6
\end{lstlisting}
\code{kind} is \code{kpi} (inside the named unit's KPIs) or
\code{stream} (inside the named stream).  Pin the numbers worth
guarding; \code{--record} writes them from the current run once you trust
it.  A case with no \file{expected} still gets the NaN guard.
\end{keyidea}

At, four canonical cases ship an \file{expected} (flash01, cstr01,
column01, membrane01); the rest run under the smoke + NaN guard.  Add an
\file{expected} whenever a case's numbers are load-bearing.

% =====================================================================
\section{Debugging}\label{sec:debug}

\subsection{Debug build}

\begin{lstlisting}
make MODE=debug all
gdb build/linux64GccDebug/choupoSolve
(gdb) r tutorials/flash01_benzene_toluene
\end{lstlisting}

The debug build leaves a separate symlink \file{choupoSolve.debug}
so the release binary stays usable in parallel.

\subsection{Verbosity}

\file{controlDict \{ verbosity 4; \}} enables debug-level prints ---
every Newton iteration, every $K$-value evaluation, every Wegstein
step. For LL flashes that misconverge, watch for stagnation:
$g(x)$ should shrink monotonically near the root.

\subsection{Sanitizers}

Edit \file{make/compiler.mk} for the debug branch:

\begin{lstlisting}[language=make]
ifeq ($(MODE),debug)
    CXXFLAGS += -fsanitize=address,undefined -fno-omit-frame-pointer
    LDFLAGS  += -fsanitize=address,undefined
endif
\end{lstlisting}

Then rebuild. Useful when a unit-op produces NaN out of nowhere or
when adding a new factory and a dispatch crashes.

\subsection{The most common failure: NaN propagation}

Symptoms: \code{verbosity 3} shows clean inputs, then
\code{V/F = nan} mid-flash. Almost always a $\log P_i^{\mathrm{sat}}$
with $P_i \le 0$ (Antoine outside its valid range) or an empty
composition (\code{F = 0}). Add asserts where you suspect it.

% =====================================================================
\section{Architecture deep-dive --- the three layers}

\begin{whatyoulllearn}
The contract between the outer driver, the simulator core, and the
post-processor. Read this twice before adding cross-layer logic.
\end{whatyoulllearn}

\begin{lstlisting}
+-----------------------------------------------------------------+
| Layer 1: OuterDriver  (src/outerDriver/)                        |
|   Sensitivity sweep, optimisation, parameter estimation, MC.    |
|   Owns: simulator_ functor, flowsheetDict_ (mutable).           |
|   Mutates flowsheetDict_ between calls via setScalarAtPath().   |
+-----------------------------------------------------------------+
                              |  calls
                              v
+-----------------------------------------------------------------+
| Layer 2: Simulator core                                         |
|   runSimulation(flowsheetDict,...)                             |
|     -> ThermoPackage::readFromDict(thermoDict, db)              |
|     -> Flowsheet::solve(flowsheetDict, thermo, verbosity)       |
|     -> SimulationResult { streams, kpis, converged }            |
|   Pure function: same inputs -> same outputs.                   |
+-----------------------------------------------------------------+
                              |  result
                              v
+-----------------------------------------------------------------+
| Layer 3: PostProcessor (src/postProcessing/)                    |
|   Sizing  -> equipment dimensions (V_R, A, t_wall, weight,...) |
|   Costing -> Guthrie/Turton with CEPCI, USD->EUR                |
|   (future) Reporting -> PDF / HTML / CSV bundles                |
|   Augments SimulationResult; never re-enters the simulator.     |
+-----------------------------------------------------------------+
\end{lstlisting}

\paragraph{The contract:}

\begin{itemize}[leftmargin=*]
  \item Layer 1 \emph{never touches the simulator's internal state} ---
        it only mutates the dict it owns and calls the functor.
  \item Layer 2 \emph{is pure} ---
        \code{runSimulation(d) == runSimulation(d)} for the same dict.
        This is what lets Layer 1 do finite-difference Jacobians and
        Levenberg-Marquardt cleanly.
  \item Layer 3 \emph{never re-runs the simulator} --- it only reads
        the result. If you need to re-run (e.g. for utility costing), do
        it via an outer driver.
\end{itemize}

\begin{keyidea}
Rule of thumb: if a quantity can be computed from a converged stream
table plus KPIs, it belongs in a post-processor. If it requires
re-solving the flowsheet, it belongs in an outer driver.
\end{keyidea}

% =====================================================================
\section{The fractal flowsheet --- nesting and \code{flattenNode}}
\label{sec:fractal}

\begin{whatyoulllearn}
How a plant of sectors of units collapses to ONE flat solver problem, why the
hierarchy is namespace-only, and how a branch runs standalone.
\end{whatyoulllearn}

A Choupo case is a \textbf{fractal}: a composite \emph{sector} nests leaf units
(or further sectors), to any depth.  But the solver never sees the tree.
\code{flattenNode} walks the declared children and collapses the hierarchy into
\textbf{one flat solver problem} with dotted \code{plant.sector.unit} names ---
the hierarchy is \emph{authoring and namespace} only, never a recursive
solver-within-a-solver.  One flat Newton-on-tears recycle solves the whole
plant.

\begin{keyidea}
\textbf{The hierarchy is discovered from the \emph{dicts}, not the filesystem.}
\code{flattenNode} iterates the \code{children(... )} declared in
\code{flowsheetDict}; a child loads as a sector only when its own
\code{flowsheetDict} exists.  This is why the numeric instant directories
(\S\ref{sec:instants}) sitting beside the sector folders are never mistaken for
sectors --- they are not declared children and have no \code{flowsheetDict}.
\end{keyidea}

\paragraph{Lean fractal folders + the dual-reader.}  A node's topology may live
either way, both first-class:
\begin{itemize}[leftmargin=*,nosep]
  \item the classic \file{<node>/system/flowsheetDict}, or
  \item the lean \file{<node>/flowsheetDict} directly at the node root.
\end{itemize}
A \textbf{dual-reader} keeps both loading: the child loader reads
\file{<child>/flowsheetDict} first and \emph{falls back} to
\file{<child>/system/flowsheetDict}; the top-level \code{choupoSolve} does the
same at the case root.  Existing fractal cases (esterification, twoSectorDemo)
load byte-identically through the fallback --- \textbf{no forced mass-migration}.

\paragraph{A branch runs alone.}  Because both \code{thermoPhysPropDict}
and the
reactions library cascade UP the parent chain (\S\ref{sec:add-data}), pointing
\code{choupoSolve} at a sub-sector folder runs that branch as a complete case,
inheriting the parent's thermo and reactions and announcing what it inherited.

\paragraph{The model-boundary rule.}  A per-unit \code{thermo \{\}} override
REPLACES the global models for that unit (components stay global, via
\code{thermoFor}).  A stream crossing such a boundary is re-interpreted there.
The governing law: \textbf{\(H\) is the conserved truth, \(T\) is the
model-dependent readout}.  A model boundary is not a physical device, so there
is no real \(\Delta T\) to absorb; H-continuity is a convention, not a law.  The
default is \textbf{hold-\(T\), let-\(H\)-jump} (the discontinuity is visible in
the printed enthalpy); a silent ``hold \(H\), flex \(T\)'' is rejected.  The
honest opt-in is a \textbf{model-boundary audit} --- print \(\Delta H =
H_{\mathrm{down}}-H_{\mathrm{up}}\) at fixed \((T,P,z)\), fold it into the
first-law ledger, and hard-refuse across any phase / \code{vf} / speciation
flip.  Full rationale: \file{docs/ai/energy.md}.

% =====================================================================
\section{Solution instants --- OpenFOAM-style per-iteration snapshots}
\label{sec:instants}

\begin{whatyoulllearn}
Where the converged (and intermediate) stream tables are written, and why the
layout is faithful to OpenFOAM's time directories.
\end{whatyoulllearn}

A run writes \textbf{durable per-iteration snapshots} of every stream, modelled
on OpenFOAM's time directories.  \code{src/io/SolutionWriter} owns the writes;
the durability core is atomic (\code{.tmp} $\to$ \code{rename} + \code{fsync},
gated on \code{flushEach}), a restart re-seeds the recycle tear from the latest
instant, and a litter sweep purges stale numeric dirs.

\paragraph{The layout --- instants at the CASE ROOT.}  An instant is a bare
iteration-\emph{number} directory at the case root, beside \file{constant/},
\file{system/} and the fractal sector folders.  ``The time speaks for itself'' ---
there is no \file{solution/} wrapper.

\begin{lstlisting}
case/
  constant/  system/  reports/        # the usual case payload
  FERMENTATION/  CONCENTRATION/  ...   # fractal SECTOR folders (alphabetic)
  0/                                   # instant 0 (the initial seed)
    streams                              # ROOT flat stream table (the spine)
    FERMENTATION/ streams                # per-sector view
    CONCENTRATION/ streams
    byUnit/                              # per-unit breakout
  1/  2/  3/                           # later recycle iterations
  latest -> 3                          # convenience symlink to the highest
  solution.log                         # the instant index
\end{lstlisting}

\begin{keyidea}
\textbf{A stream IS a surface field (\(\phi\)); the OpenFOAM analogy is exact.}
The payload file is named \code{streams}; \code{latest} is the
highest-numbered instant (latestTime style --- restart discovers it by the max
integer dir name).  A load-bearing \emph{all-digit filter} on the numeric-dir
scan guarantees a sector folder (alphabetic) is never swept or mistaken for an
instant, and the fractal loader iterating declared children (not a filesystem
walk) guarantees the reverse.
\end{keyidea}

The pilot is \file{tutorials/plant/ChemicalPlantTutorial}; \code{writeInterval}
controls how many instants land (set it to \code{1} on a 3-iteration case or
you only see \code{0} and the final --- and miss the whole tear-residual
march).

% =====================================================================
\section{The data cascade --- arity, scope, kind}\label{sec:data-cascade}

\begin{whatyoulllearn}
The mechanical decision procedure for WHERE every thermodynamic number lives.
This is the engine-side summary; the canonical, ratified reference is
\file{docs/ai/data-doctrine.md} --- read it before any curation act.
\end{whatyoulllearn}

\begin{keyidea}
\textbf{A datum's home is decided by ARITY, then SCOPE, then KIND --- in that
order.}  Read the quantity's \emph{definition}, not its frequency of use.
\end{keyidea}

\paragraph{(1) Arity --- how many species does the definition name?}
\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{Arity-1} (the molecule alone on the table, no partner named ---
        \emph{not even implicitly}) $\to$ INTRINSIC,
        \file{data/standards/components/<name>.dat}: \(T_c, P_c, \omega\), MW,
        \code{standardThermochemistry}, \code{solid}, \code{liquidPure},
        \code{transport}.
  \item \textbf{Arity-2} (the definition must name a partner --- a solvent, a
        counter-ion, an interaction species, \emph{even water}) $\to$ PAIR/SET
        data in a \textbf{catalogue}, keyed by the pair, cited per value,
        referenced by name.  NRTL \(\tau\), Henry \(H\), Pitzer \(\beta\), EoS
        \(k_{ij}\), a heat of solution.  \textbf{Never copied into the
        \code{.dat}.}
\end{itemize}

\paragraph{(2) Scope --- at what level is the number TRUE?}  A datum lives at
the \emph{highest level where it is true} (the canonical value is a curation act
in \file{data/standards/}).  A lower node overlays only when it makes the value
\emph{more} true --- a sample-measured refinement.  Precedence, low to high:
\code{local < standard < case < sector < unit} (\code{local} = the private,
announced \file{data/local/} tier).

\paragraph{(3) Kind --- the molecule, or the molecule-in-a-machine?}  A
\emph{rate} (kinetics) or a \emph{geometry / distribution} (PSD) is NOT
component data at all --- it lives in the operation's \code{constant/}
(\code{constant/crystallisation}, \code{constant/dryingKinetics},
\code{constant/reactions}).  PSD is a \emph{stream} attribute.

\paragraph{The aqueous TIER and the default-solvent resolver.}  Water gets
\emph{no} exemption from the arity test: an ``in-water'' property is pair data.
Its one earned privilege is a single canonical, by-name aqueous reference tier
--- \file{data/standards/electrolyte/ions.dat} for ions, and
\file{data/standards/solution/<solute>-<solvent>.dat} for molecular solutes
(the heat of solution).  The package declares a default solvent once
(\code{solvent water;}); when an in-solvent property is needed and no solvent is
named at the call site, the resolver uses the default \textbf{and announces it
every run}:

\begin{lstlisting}
[thermo] dHsoln(sucrose): solvent not named -> DEFAULT = water;
         solution/sucrose-water.dat  [Putnam & Kilday 1986]
\end{lstlisting}

\noindent Off-default it \emph{fails with a remedy} (``provide
solution/sucrose-ethanol.dat or set \code{solvent water;} explicitly'') --- it
never silently substitutes the water number.  This is implemented in
\code{ThermoPackage} (\code{solventName\_}, the \code{SolutionRegistry} walk).

\paragraph{The overlay merges block-by-block.}  As in \S\ref{sec:add-data}: a
reference-state block (\code{solid\{\}}, \code{gasIdeal\{\}},
\code{liquidPure\{\}}, each \code{eosParameters.<model>\{\}}) is the atomic unit
of physical meaning.  An overlay replaces the whole block --- the curator copies
the whole block they refine; a lone-scalar overlay is the forbidden hidden
hybrid.

\paragraph{Model extensibility --- \code{eosParameters\{ <model>\{\dots\} \}}.}
The same definition test, asked of the \emph{model}:
\begin{enumerate}[leftmargin=*,nosep]
  \item A model needing only the corresponding-states triad \(\{T_c, P_c,
        \omega\}\) (every cubic: SRK, PR, \dots) \textbf{reads the triad off the
        Component and derives its own \(a_c, b, m, \alpha(T)\) in source} ---
        adding it touches \emph{zero} data files (this is why adding PR after
        SRK touched no \code{.dat}).  Do not pre-bake \(a_c\) into the
        \code{.dat}.
  \item A model-specific PURE parameter that cannot be generated (PC-SAFT's
        \(m, \sigma, \varepsilon/k\)) goes in a \code{model}-keyed sub-block on
        the component, \code{eosParameters\{ PCSAFT\{\dots\} \}}, each factory
        reading \emph{its own} sub-block and ignoring the rest.
  \item PAIR parameters (\(k_{ij}\)) go in a model-keyed catalogue,
        \file{data/standards/eos/<model>/binaryInteractions/<i>-<j>.dat}, with
        the always-permitted inline override in the package.
\end{enumerate}
A model with no required \code{eosParameters.<model>} block \textbf{fails with a
remedy}, never a silent corresponding-states fallback.

% =====================================================================
\section{The elements datum --- one enthalpy surface}\label{sec:energy-datum}

\begin{whatyoulllearn}
Why the energy balance runs on ONE datum, how the heat of reaction / solution /
crystallisation emerge from it, and what throws when a datum is missing.
\end{whatyoulllearn}

A heat balance that has to close \emph{through} a reactor, a crystalliser, or a
dissolution cannot use a sensible (arbitrary-\(T_{\mathrm{ref}}\)) datum --- it
would silently drop the heat of reaction.  Choupo runs the energy ledger on the
\textbf{elements datum}: formation enthalpies at 298.15\,K, the same reference
as the ideal-gas \(H_{ig}\), so the reaction / phase-change / dissolution
enthalpy \emph{emerges} as the difference of formation enthalpies.

\begin{keyidea}
\textbf{One datum, no silent fallback (Vítor's law: elements-only,
throw-on-gap).}  The second datum and every silent path were deleted ---
\code{streamH\_sensible}, \code{solidH\_sensible}, the \code{tryDatum(elements
= false)} fallback, the \code{EnergyRef::Sensible} arm.
\code{streamH\_elements} now \emph{propagates a real missing-Hf gap}, naming the
component, instead of swallowing it.  A curation gap is loud, never a quietly
wrong closure.
\end{keyidea}

\paragraph{The solid term.}  A stream's enthalpy is computed over both phases:
the fluid (\code{s.F}, \code{s.z}) \textbf{and} the solid carried in
\code{s.s[]} (kmol/s per component), on the same elements datum ---
\(\sum_i s.s[i]\cdot h^\circ_i(\mathrm{solid}, T)\), the same solid formation
leg \code{Component::h\_formation(T,"solid")} uses.  Before this, the crystals a
crystalliser had just made were silently dropped from the stream enthalpy and
the heat of crystallisation read zero; counting \code{s.s[]} (plus the aqueous
solution tier for the dissolved solute) closed the crystalliser balance.

\paragraph{Utility legs vs process streams --- no double-count.}  The ledger
partitions each unit's ins/outs by \code{category}
(\file{src/reporting/BalanceMath.H}, \code{EnergyBalanceReport.cpp}):
utility-tagged legs (steam $\to$ condensate) feed a \code{qBoundary}
accumulator --- the real heat crossing the boundary --- while process streams
feed \(h_{\mathrm{in}} / h_{\mathrm{out}}\).  An evaporator's \code{duty\_kW} is
delivered \emph{through its own steam stream}, so it is INTERNAL and must NOT
also be summed as an external item; a process-to-process exchanger's duty is
likewise internal.  Closure \(= 100\cdot\Delta H / \text{supplied}\) over
genuine boundary heat only.  This killed a fake \(2131\,\%\) non-closure that
was a pure double-count, not lost latent heat.

\paragraph{One enthalpy surface everywhere.}  The published stream \code{.H}
(Flowsheet) and the isothermal-flash duty (\code{IsothermalFlash.cpp}) now run
on the \emph{same} canonical elements-datum form the report reads
(\code{H\_stream\_formation}, \(h_{ig} - \Delta h_{\mathrm{vap}}(T)\)), not a
Watson-at-298 + \(\int c_p^{\mathrm{liq}}\) variant.  The two differed by the
liquid latent model and surfaced as a \(\sim24\,\%\) gap between a flash's
reported \(Q\) and its own stream \(\Delta H\); unifying the surface moves only
the reported \emph{duty} (an energy KPI), never the material flows.

% =====================================================================
\section{Extending Choupo --- step by step}\label{sec:recipes}

\begin{whatyoulllearn}
Two complete, copy-pasteable walkthroughs ---
\emph{tim-tim-por-tim-tim} --- each ending in a green
\code{bin/runTests}.  Recipe~A: a new unit operation, end to end.  Recipe~B: a
new property-prediction method, with the validation-first step that earns your
trust before you simulate with it.
\end{whatyoulllearn}

% --------------------------------------------------------------------
\subsection{Recipe A --- a new unit operation, end to end}
\label{sec:recipe-unitop}

\textbf{The worked example: \code{isothermalSplitter}} --- a unit that splits
one feed into two outlets by a fraction \(\theta\) to the top, both outlets at
the feed's \(T, P, z\) (composition unchanged; the spec is the split fraction).
Trivially simple physics on purpose --- so every \emph{mechanical} step is
visible.

\paragraph{Step 1 --- The header.}
\file{src/unitOperations/mixer/IsothermalSplitter.H} (it belongs beside the
existing \code{Splitter}):

\begin{lstlisting}[language=C++]
#pragma once
#include "core/Dictionary.H"
#include "thermo/ThermoPackage.H"
#include "unitOperations/UnitOperation.H"

namespace Choupo {

class IsothermalSplitter : public UnitOperation {
public:
    IsothermalSplitter() = default;
    const std::string& type() const override { return type_; }

    int solve(const DictPtr& dict,
              const ThermoPackage& thermo,
              int verbosity) override;

    std::vector<ProcessStream>      producedStreams() const override { return produced_; }
    std::map<std::string, scalar>   kpis()            const override { return kpis_; }

private:
    inline static const std::string type_ = "isothermalSplitter";
    std::vector<ProcessStream>      produced_;
    std::map<std::string, scalar>   kpis_;
};

} // namespace Choupo
\end{lstlisting}

\noindent Note the exact \code{solve} signature ---
\code{int solve(const DictPtr\&, const ThermoPackage\&, int)} --- and that
\code{type()} returns a reference to a member string.  Prepend the licence
banner (\S on header banner); it doubles as the licence notice.

\paragraph{Step 2 --- The implementation.}
\file{src/unitOperations/mixer/IsothermalSplitter.cpp}.  The Flowsheet injects
the inlet and your parameters as sub-dicts of \code{dict}; you publish outlets
into \code{produced\_} in the order your \code{outputs (... )} declares:

\begin{lstlisting}[language=C++]
#include "mixer/IsothermalSplitter.H"
#include "core/Dimensions.H"
#include <iostream>

namespace Choupo {

int IsothermalSplitter::solve(const DictPtr& dict,
                              const ThermoPackage& thermo,
                              int verbosity)
{
    auto feed = dict->subDict("feed");          // F (kmol/s SI), Tfeed, Pfeed
    auto comp = dict->subDict("composition");   // mole fractions z_i
    auto op   = dict->subDict("operation");     // YOUR parameters

    const scalar F     = feed->lookupScalar("F",     Dims::molarFlow);
    const scalar Tfeed = feed->lookupScalar("Tfeed", Dims::temperature);
    const scalar Pfeed = feed->lookupScalar("Pfeed", Dims::pressure);
    const scalar theta = op->lookupScalar("splitFraction");    // [-], 0..1

    if (theta < 0.0 || theta > 1.0)
        throw std::runtime_error(
            "isothermalSplitter: splitFraction must be in [0,1]");

    std::vector<scalar> z(thermo.n());
    for (std::size_t i = 0; i < thermo.n(); ++i)
        z[i] = comp->lookupScalar(thermo.comp(i).name());

    if (verbosity >= 3)
        std::cout << "  [isothermalSplitter] F = " << (F * 3600.0)
                  << " kmol/h, split " << theta << " to top\n";

    produced_.clear();                          // ORDER == outputs(...) order
    ProcessStream top;  top.name = "top";
    top.F = theta * F;  top.T = Tfeed;  top.P = Pfeed;  top.z = z;
    produced_.push_back(top);

    ProcessStream bot;  bot.name = "bottom";
    bot.F = (1.0 - theta) * F;  bot.T = Tfeed;  bot.P = Pfeed;  bot.z = z;
    produced_.push_back(bot);

    kpis_.clear();                              // KPIs feed sizing/costing/sweep
    kpis_["F_top"]         = top.F;
    kpis_["F_bottom"]      = bot.F;
    kpis_["splitFraction"] = theta;

    return 0;                                   // 0 == converged
}

} // namespace Choupo
\end{lstlisting}

\begin{keyidea}
\textbf{Fill EVERY field of a produced stream} (\code{F, T, P, z}, and
\code{vf} / \code{s} if relevant).  A half-filled stream propagates NaN
downstream and the NaN guard fails the case --- by design.
\end{keyidea}

\paragraph{Step 3 --- Register it (the one line).}  Edit
\file{src/unitOperations/UnitOperation.cpp}: add the include beside the other
mixer includes, and one line in \code{registerBuiltins()}:

\begin{lstlisting}[language=C++]
#include "mixer/IsothermalSplitter.H"       // (1) beside #include "mixer/Splitter.H"
//...
void UnitOperation::registerBuiltins()
{
    auto reg = [](const std::string& name, auto&& maker) {
        registerType(name, std::forward<decltype(maker)>(maker)); };
    //...
    reg("splitter",           []{ return std::make_unique<Splitter>(); });
    reg("isothermalSplitter", []{ return std::make_unique<IsothermalSplitter>(); }); // (2)
    //...
}
\end{lstlisting}

\noindent That is the \emph{entire} registration story --- no macro, no static
initialiser.  A student greps \code{registerBuiltins} and sees your type listed.

\paragraph{Step 4 --- Build (the recursive pickup).}

\begin{lstlisting}
make all
\end{lstlisting}

\noindent The Makefile discovers sources with \code{find src -name '*.cpp'}, so
the new \file{.cpp} is compiled with no Makefile edit; \code{-MMD -MP} tracks
the new header dependency automatically.

\paragraph{Step 5 --- Rebuild the WASM (MANDATORY).}

\begin{lstlisting}
make wasm-gui          # = choupoSolve + choupoProps, the two the GUI uses
\end{lstlisting}

\begin{pitfall}
\textbf{Skip this and the GUI throws \code{UnitOperation::New: unknown type
'isothermalSplitter'}} even though the native binary runs the type fine.  The
WASM solver is a separate Emscripten compile; a new built-in only reaches the
browser after a WASM rebuild.  If \code{make wasm} reports ``Nothing to be
done'', force it: \code{make wasm-clean \&\& make wasm}.  Never run two
\code{make wasm} concurrently --- they clobber \file{gui/public/wasm/}.
\end{pitfall}

\paragraph{Step 6 --- A minimal tutorial case.}  Keep it trivial so it doubles
as a regression test.  \file{tutorials/steady/mixer/splitter01\_isothermal/}:

\begin{lstlisting}
splitter01_isothermal/
  splitter01.cho                 # empty marker (the GUI's openable entity)
  system/
    controlDict                  # application choupoSolve; verbosity 3;
    flowsheetDict                # the topology below
  constant/
    thermoPhysPropDict           # v2 record: components + gammaPhi ideal
\end{lstlisting}

\noindent The \file{system/flowsheetDict} (one feed boundary, one unit, two
product boundaries):

\begin{lstlisting}[language=dict]
boundary
{
    inlets  ( feed );
    feed { F 1.0 kmol/s; Tfeed 350 K; Pfeed 1 bar;
            composition { benzene 0.5; toluene 0.5; } }
}

units
(
    {
        name      S1;
        type      isothermalSplitter;
        in        feed;
        outputs   ( top bottom );        // ORDER matches produced_ order
        operation { splitFraction 0.3; }
    }
);
\end{lstlisting}

\paragraph{Step 7 --- Add KPIs to the regression (\code{expected}).}  Your
\code{kpis()} already feed sizing, costing and the sweep/optimisation drivers;
pin the load-bearing ones to golden masters.  Drop a
\file{tutorials/steady/mixer/splitter01\_isothermal/expected}:

\begin{lstlisting}
# kind   name   key             expected   reltol
kpi      S1     F_top           0.3        1e-6
kpi      S1     F_bottom        0.7        1e-6
stream   top    F               0.3        1e-6
\end{lstlisting}

\noindent then record/verify and run the full gate:

\begin{lstlisting}
bin/runTests --record tutorials/steady/mixer/splitter01_isothermal   # write from a trusted run
bin/runTests                                                          # the whole corpus must stay green
\end{lstlisting}

\begin{keyidea}
\textbf{The gate is \code{bin/runTests}, not a bare exit-code loop} ---
NaN/inf guard on \emph{every} case plus the \code{expected} KPI/stream
comparison (\S\ref{sec:regression}).  A solver can return NaN with exit code 0;
only the structured-result scan catches it.
\end{keyidea}

% --------------------------------------------------------------------
\subsection{Recipe B --- a new property-prediction method}
\label{sec:recipe-property}

A property model plugs into one of four bases.  \textbf{Pick the base by the
quantity you predict}, then follow the same \code{fromDict} + \code{registerModel}
pattern --- and validate against measured data \emph{before} you trust it.

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
You predict\dots & Base & \code{registerModel} factory signature \\
\midrule
\(P^{\mathrm{sat}}(T)\)        & \code{VaporPressureModel} & \code{(const DictPtr\&)} \\
\(c_p(T)\)                     & \code{HeatCapacityModel}  & \code{(const DictPtr\&)} \\
\(\gamma_i(T,x)\)              & \code{ActivityModel}      & \code{(const DictPtr\&, const std::vector<std::string>\&)} \\
\(\varphi_i,\,Z,\,H^{\mathrm{res}}\) & \code{EquationOfState} & \code{(const DictPtr\&, const std::vector<Component>\&)} \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{B.1 --- A vapour-pressure model (the simplest base).}  Say you add a
two-parameter \code{Clapeyron} \(P^{\mathrm{sat}}\).

\begin{enumerate}[leftmargin=*]
  \item \file{src/thermo/vaporPressure/Clapeyron.\{H,cpp\}}, inheriting
        \code{VaporPressureModel}, override the pure virtual
        \code{scalar Psat\_Pa(scalar T\_K) const}; a constructor
        \code{Clapeyron(const DictPtr\&)} reads the parameters (mirror
        \file{Antoine.cpp}, whose ctor is \code{Antoine::Antoine(const
        DictPtr\&)}).
  \item Register --- include + one line in
        \code{VaporPressureModel::registerBuiltins()}:
\begin{lstlisting}[language=C++]
#include "Clapeyron.H"
//...
void VaporPressureModel::registerBuiltins()
{
    registerModel("Antoine",
        [](const DictPtr& d) -> std::unique_ptr<VaporPressureModel>
        { return std::make_unique<Antoine>(d); });
    registerModel("AmbroseWalton",
        [](const DictPtr& d) -> std::unique_ptr<VaporPressureModel>
        { return std::make_unique<AmbroseWalton>(d); });
    registerModel("Clapeyron",                                          // new
        [](const DictPtr& d) -> std::unique_ptr<VaporPressureModel>
        { return std::make_unique<Clapeyron>(d); });
}
\end{lstlisting}
  \item Select it in a component's \code{vaporPressure\{ model Clapeyron; ... \}}
        block.
\end{enumerate}

\paragraph{B.2 --- Validate FIRST (the AAD step).}  A property model is
\textbf{not trustworthy until it has been checked against measured data}.
Choupo's \code{choupoProps} binary is the validation weapon: run the
\emph{measured} dataset through the model and read the average absolute
deviation (AAD) before you let the model into a flowsheet.

\begin{lstlisting}
choupoProps <validationCase>          # e.g. a propertyScan1D over measured T
# the run prints, per point:  model vs measured, dev %;  headline:
#   AAD = 1.8 %   AAD tolerance = 3.0 %   -> PASS
\end{lstlisting}

\begin{keyidea}
\textbf{The number that earns trust is the AAD against measured data, not that
the model compiles.}  The bench prints the per-point deviation and a headline
AAD against a tolerance (the same machinery \code{HeatTransferBench} and the
Pitzer activity validation use); a model that misses its anchor beyond
tolerance fails loudly.  This is the property half of ``see, then decide'' ---
you simulate with a model \emph{because} you have seen its AAD, never because a
dropdown recommended it.
\end{keyidea}

\paragraph{B.3 --- An equation of state (and the data-doctrine path).}  An EoS
is the one base where WHERE the parameters live is a doctrine decision.  The
canonical reference is \file{docs/ai/data-doctrine.md}~\S4; in short:

\begin{enumerate}[leftmargin=*]
  \item \textbf{Corresponding-states EoS (reads only the triad).}  If the model
        needs only \(\{T_c, P_c, \omega\}\), \textbf{read them off the
        \code{Component} and derive the rest in source} --- touch \emph{zero}
        data files.  This is exactly what \code{SRK} and \code{PR} do: the
        \code{SRK} constructor reads \code{c.Tc()}, \code{c.Pc()},
        \code{c.omega()} and computes \(a_c = \Omega_a R^2 T_c^2 / P_c\),
        \(b = \Omega_b R T_c / P_c\), \(m = m(\omega)\).  Do \emph{not}
        pre-bake \(a_c\) into the \code{.dat} (it would duplicate a value the
        model owns and let the two drift).
  \item \textbf{Model-specific PURE parameters (cannot be generated).}  PC-SAFT's
        \(\{m, \sigma, \varepsilon/k\}\) live in a \code{model}-keyed sub-block
        on the component --- additive, never disturbing the triad:
\begin{lstlisting}[language=dict]
// inside the component .dat
eosParameters
{
    PCSAFT  { m 2.4653;  sigma 3.6478e-10;  epsilon_k 287.35;
              provenance { origin regressed; method "Gross & Sadowski 2001"; } }
    // SRK/PR appear in NO key here -- they read Tc, Pc, omega
}
\end{lstlisting}
        Each factory reads \emph{its own} \code{eosParameters.<model>} sub-block
        by name and ignores the rest, so one molecule carries SRK + PR +
        PC-SAFT simultaneously, each with its own \code{provenance\{\}}.
  \item \textbf{Pair parameters \(k_{ij}\).}  A model-keyed catalogue,
        \file{data/standards/eos/<model>/binaryInteractions/<i>-<j>.dat}, with
        the always-permitted inline override.  \code{SRK::fromDict} already
        reads an inline \code{binaryInteractions} list of
        \code{\{ i; j; kij; \}} dicts and resolves names positionally against
        the component list.
\end{enumerate}

\paragraph{The explicit factory, and the open follow-up.}  Register the EoS the
same explicit way, dispatching to a \code{fromDict} static (the
\code{Component}-list overload):

\begin{lstlisting}[language=C++]
registerModel("PCSAFT",
    [](const DictPtr& dict, const std::vector<Component>& comps)
        -> std::unique_ptr<EquationOfState>
    { return PCSAFT::fromDict(dict, comps); });
\end{lstlisting}

\begin{pitfall}
\textbf{Glass-box requirement:} an EoS with no required
\code{eosParameters.<model>} block on a component must \textbf{fail with a
remedy} (``PCSAFT needs m, sigma, epsilon\_k for X; fit them with choupoProps or
pick SRK/PR which run off Tc, Pc, omega''), \emph{never} a silent
corresponding-states fallback.  \textbf{Open follow-up} (noted in the doctrine):
a uniform \code{modelParams} accessor on \code{Component} that hands a factory
its \code{eosParameters.<model>} sub-dict by model name is not yet generalised
--- today each EoS that needs a pure block reads it itself.  Until that lands,
copy the \code{liquidViscosity\{\}} raw-sub-dict reader pattern.
\end{pitfall}

% =====================================================================
\section{Roadmap}

Ordered by impact for the MCFT course and V\'\i tor's research:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}llX@{}}
\toprule
\textbf{Target} & \textbf{Idea}                                                          & \textbf{Difficulty} \\
\midrule (shipped) & Levenberg-Marquardt parameter-estimation driver (\code{fitBinaryPair})  & --- \\ & Nelder-Mead optimisation driver (minimise cost / TAC)                   & easy \\ & LL flash via 2nd-order Michelsen TPD with multiple restarts             & hard \\ & Solids: inert phase + filter / cyclone / dryer unit-ops                 & medium \\ & Crystalliser + PSD basics                                               & hard \\ & NRTL distillation robustness (Naphtali-Sandholm)                        & medium-hard \\
v1.0            & Equation-oriented sparse Newton for the full flowsheet                  & very hard \\
$> $ v1         & Membrane NF/RO unit op (V\'\i tor's research bridge)                    & research-grade \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{Tech-debt items also welcome:}
\begin{itemize}[leftmargin=*,nosep]
  \item Reference logs for every tutorial $\to$ in-tree regression diff.
  \item Reaction-kinetics library (Arrhenius, Langmuir-Hinshelwood,
        Michaelis-Menten) as a \code{Kinetics} base class.
\end{itemize}

% =====================================================================
\section{Things to NOT do}\label{sec:donts}

\begin{whatyoulllearn}
A short list, each item born from a real incident. Memorise it.
\end{whatyoulllearn}

\begin{itemize}[leftmargin=*]
  \item Do \textbf{not} suggest a Python rewrite or wrapper. The course
        teaches numerical methods in \cpp{}; rewriting in Python would
        defeat the purpose.
  \item Do \textbf{not} add macro auto-registration. Explicit
        \code{registerBuiltins()} is non-negotiable (see \S\ref{sec:factory}).
  \item Do \textbf{not} add a heavy dependency --- even a ``small'' one.
        Pedagogical visibility outweighs convenience.
  \item Do \textbf{not} split a binary by numerical strategy
        (\code{*Foam}-style proliferation). Strategy is run-time. Splitting
        by \emph{problem class} --- the four binaries Solve / Batch / Ctrl /
        Props --- is the deliberate exception, not a precedent for more.
  \item Do \textbf{not} bypass the dict format. Never use YAML, JSON, or
        TOML inside the case directory. The dict is the canonical wire
        format.
  \item Do \textbf{not} relitigate decisions in \file{CLAUDE.md} \S10
        without first reading what's already been discussed and why.
  \item Do \textbf{not} merge code that breaks the regression, even if
        the failure is ``obviously unrelated''.
\end{itemize}

% =====================================================================
\section{Quick reference}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Task}                           & \textbf{Command / file} \\
\midrule
Build release                            & \code{make all} \\
Build debug                              & \code{make MODE=debug all} \\
Clean                                    & \code{make clean} / \code{make distclean} \\
Source env                               & \code{source etc/bashrc} \\
Run a case                               & \code{runCase tutorials/<name>} \\
Run from inside case dir                 & \code{runCase} \\
List tutorials                           & \code{listCases} \\
List built-in unit-op types (in code)    & \code{UnitOperation::availableTypes()} \\
Tweak a dict field programmatically      & \code{dict->setScalarAtPath(path, v)} \\
Add a unit op (full walkthrough)         & \S\ref{sec:recipe-unitop} (quick: \S\ref{sec:add-unitop-quick}) \\
Add a property model (full walkthrough)  & \S\ref{sec:recipe-property} (quick: \S\ref{sec:add-thermo-model}) \\
Add a component / material / reaction    & \S\ref{sec:add-data} \\
Add an outer driver / post-processor     & --- (see the outer-driver section) \\
Where does a number live?                & \S\ref{sec:data-cascade} + \file{docs/ai/data-doctrine.md} \\
Energy datum / balance closure           & \S\ref{sec:energy-datum} \\
Fractal flowsheet + instants             & \S\ref{sec:fractal}, \S\ref{sec:instants} \\
Project conventions / AI rules           & \file{CLAUDE.md} \\
User-facing how-to                       & \file{docs/userGuide.pdf} \\
\bottomrule
\end{tabularx}
\end{center}

\vspace{2cm}
\noindent\rule{\linewidth}{0.3pt}
\noindent\small\textit{Colophon.}\quad This manual was typeset in Latin
Modern with the Choupo shared LaTeX preamble. Sources live in
\file{docs/}; rebuild with \code{cd docs \&\& make}.

\end{document}
