\documentclass[11pt,a4paper]{article}
\newcommand{\docTitle}{Explorer Guide}
\input{preamble.tex}

\title{}
\author{}
\date{}

\begin{document}

\manualfrontmatter{Explorer Guide}{
Choupo is a trademark of TalentGround Lda.

Anthropic, Claude, and Claude Code are trademarks or registered trademarks of
Anthropic, PBC.
}

% =====================================================================
\section{What the Explorer is}

The Explorer is Choupo's \emph{interactive thermodynamics bench}: pick
components, pick a view, and SEE what the selected property models do ---
instantly, in the browser, with no case authoring at all.  It answers the
question that comes \emph{before} any simulation: \textbf{``what does this
thermodynamics look like?''}

It sits deliberately between two other surfaces:
\begin{itemize}[leftmargin=*,nosep]
  \item The \textbf{props bench} (\code{choupoProps} cases) asks the same
        questions \emph{reproducibly}: a case file, golden-locked numbers,
        validation blocks.  The Explorer is its sketchpad --- what you find
        here, you pin down there.
  \item The \textbf{simulators} consume the models the Explorer displays.
        A surprise in a flowsheet (an azeotrope, a scaling mineral, a
        runaway $\gamma$) is usually visible in the Explorer in thirty
        seconds.
\end{itemize}

Everything the Explorer draws is computed by the SAME C++ engine
(compiled to WebAssembly) with the SAME curated catalogue
(\file{data/standards/}) the solvers use.  There is no ``display model'':
if the Explorer shows an azeotrope at $x = 0.89$, the distillation column
will hit it at $x = 0.89$.

\section{The workflow}

\begin{enumerate}[leftmargin=*,nosep]
  \item \textbf{Pick components} in the left panel.  The list is the live
        catalogue; each card shows the data the component actually carries
        (criticals, $P^{\mathrm{sat}}$, $c_p$, formation --- and what it
        lacks).  Case-local components (from an open case's
        \file{constant/}) appear alongside, marked.
  \item \textbf{Pick a view} (the tabs enable themselves when their
        component-count and data requirements are met; a disabled tab tells
        you WHY it is disabled).
  \item \textbf{Pick the method} where more than one applies ---
        activity model, equation of state, $P^{\mathrm{sat}}$ correlation.
        Overlaying two methods on the same axes is the fastest modelling
        lesson Choupo offers: the gap between the curves IS the modelling
        uncertainty.
  \item \textbf{Export}: every view is a Plotly figure (zoom, hover, save
        PNG) backed by numbers you can carry into a props case for the
        golden-locked version.
\end{enumerate}

\section{The views, one by one}

\subsection{Property vs $T$/$P$ (scan)}
Any pure-component property --- $P^{\mathrm{sat}}$, liquid density,
$c_p$, viscosity, enthalpy --- swept over temperature or pressure, for one
or many components.  Use it to: check a correlation's range before trusting
it in a case; compare components; spot the nonsense that extrapolation
produces outside a fit's window (the guide's recurring lesson: a
correlation is DATA, and data has a domain).

\subsection{Pure phase diagram ($P$--$T$)}
The full phase map of one substance: sublimation, melting and vapour-pressure
lines meeting at the triple point, ending at the critical point.  Expect
water's negatively-sloped melting line (the anomaly), and the supercritical
region where the liquid/vapour distinction dissolves.

\subsection{Binary boiling envelope ($T$--$x$--$y$)}
The bubble and dew curves of a binary at fixed $P$.  This is where
azeotropes announce themselves: the curves touch, and no distillation can
cross the touching point.  Switch the activity model (ideal $\to$ NRTL
$\to$ UNIFAC) and watch the envelope bend --- ideal solutions cannot
azeotrope; real ones do.

\subsection{$\gamma(x)$ --- activity coefficients}
The raw non-ideality of a binary: $\gamma_1, \gamma_2$ against composition,
with the infinite-dilution values $\gamma^\infty$ at the edges.
$\gamma > 1$: the pair dislikes mixing (positive deviation, minimum-boiling
azeotropes); $\gamma < 1$: association (maximum-boiling).  The
Gibbs--Duhem constraint ties the two curves together --- one cannot bend
without the other answering.

\subsection{McCabe--Thiele (distillation)}
The classic $x$--$y$ construction: equilibrium curve, operating lines,
$q$-line, stages stepped off graphically.  Change reflux and watch the
stage count respond; push the specs toward an azeotrope and watch the pinch
strangle the staircase.  The fastest intuition-builder for distillation
that exists --- and it runs on the same $K$-values the rigorous column
uses.

\subsection{Binary flash ($x$--$y$ + lever rule)}
One flash, drawn: feed, vapour and liquid on the equilibrium diagram with
the lever rule made visible.  Use it to understand what a single
equilibrium stage can and cannot do.

\subsection{Binary and ternary LLE}
Liquid--liquid splitting from the mixing Gibbs energy: $g_{\mathrm{mix}}(x)$
with the common-tangent construction (binary), and the ternary solubility
envelope with tie-lines (UNIFAC-predicted).  Expect the tangent points ---
not the minima --- to define the coexisting phases; that distinction is the
whole thermodynamics of demixing.

\subsection{Ternary boiling surface}
$T_{\mathrm{bubble}}$ over the composition triangle: valleys are
distillation boundaries, and azeotropes sit at the stationary points.

\subsection{Psychrometric chart}
Humid air: humidity vs temperature with saturation, wet-bulb and enthalpy
structure --- the working chart for dryers and HVAC, computed from the same
gas-phase thermodynamics as everything else.

\subsection{Scaling (SI vs recovery)}
The membrane/brine screening view: concentrate a water composition over
recovery and watch each mineral's saturation index climb toward
$\mathrm{SI} = 0$ (the precipitation line).  The Davies-vs-Pitzer method
switch is the lesson: at high ionic strength the two activity models give
DIFFERENT scaling verdicts, and choosing one is an engineering decision.

\subsection{Equilibrium map (Gibbs)}
The chemical-reaction view: pick two or more gas-phase species (with plain
formulas --- the atom matrix is derived from them, glass-box) and a feed,
and get iso-lines of a product's equilibrium mole fraction over the
$T\times P$ plane.  The whole point is the tension it makes visible: for
ammonia ($\mathrm{N_2+3H_2\rightleftharpoons 2NH_3}$) the equilibrium
optimum sits cold and high-pressure, yet the labelled industrial window
sits hot --- where the catalyst can work.  Answers the beginner's real
confusion first (``where do I write the reaction?'' --- you don't; Gibbs
redistributes the ATOMS you fed, shown as a live element balance).
\emph{Click any cell} for its full equilibrium composition and the
ready-to-run \code{gibbsReactor} case dict --- the map is a case generator.
The \emph{advanced} toggle exposes the temperature-approach $\Delta T$
(the empirical closeness-to-equilibrium knob commercial tools hide); it snaps rather
than animates, and ghosts the $\Delta T = 0$ contours underneath so you
read the before/after, not a motion.  Unconverged cells are marked, never
interpolated.

\subsection{Steam tables (IF97)}
Saturation and superheat properties of water from IAPWS-IF97 ---
$h_f, h_g, h_{fg}$, entropies, volumes --- the backbone of every steam and
power-cycle case, browsable.

\section{Provenance and honesty rules}

The Explorer inherits the project's credo:
\begin{itemize}[leftmargin=*,nosep]
  \item \textbf{Every number has a source.}  Component cards show where each
        value came from; estimates carry their UNVALIDATED badge; nothing is
        silently filled in.
  \item \textbf{Disabled means explained.}  A view that cannot run tells you
        which requirement failed (component count, missing VLE data, missing
        UNIFAC groups) --- never a dead button.
  \item \textbf{The Explorer never writes.}  It reads the catalogue and the
        open case; promoting data, fitting parameters and locking goldens
        belong to the props bench and the curation tools.  Sketch here,
        commit there.
\end{itemize}

\section{From Explorer to case: the hand-off}

A typical session: you notice in the $T$--$x$--$y$ view that your solvent
pair azeotropes at 78\,\% --- so the simple column in your flowsheet cannot
deliver 99\,\% purity.  The hand-off is: (1) reproduce the envelope as a
props case (\code{propertyScanBinary}) so the finding is golden-locked;
(2) switch the flowsheet to the azeotrope-aware design (pressure-swing or
entrainer); (3) cite the Explorer finding in the case header.  The Explorer
is where you LEARN; cases are where knowledge is kept.

\end{document}
