\documentclass[11pt,a4paper]{article}
\newcommand{\docTitle}{Props Workflow Guide}
\newcommand{\propertyTheoryAuthors}{}
\input{preamble.tex}

\title{}
\author{}
\date{}

\begin{document}

\manualfrontmatter{Properties Guide}{
Choupo is a trademark of TalentGround Lda.

Anthropic, Claude, and Claude Code are trademarks or registered trademarks of
Anthropic, PBC.

NIST WebBook is a product or service of the National Institute of Standards
and Technology.
}

\pagenumbering{roman}
\setcounter{tocdepth}{2}
\markboth{Contents}{Contents}
\tableofcontents
\clearpage
\pagenumbering{arabic}

% =====================================================================
%   Introduction --- the difficulty, the map, the promise
% =====================================================================
\begin{quote}
\itshape ``Thermodynamics is a funny subject. The first time you go through it,
you don't understand it at all. The second time you go through it, you think you
understand it, except for one or two points. The third time you go through it,
you know you don't understand it, but by that time you are so used to it, it
doesn't bother you any more.''
\upshape\par\smallskip\hfill --- Arnold Sommerfeld
\end{quote}

\section{Introduction: the difficulty, the map, the promise}

\emph{What this guide is --- and is not.}  This is the props-\textbf{mode
workflow} guide: how you \emph{estimate} a missing compound, \emph{fit} its
binary pairs, and \emph{consistency-test} the data --- the see-then-decide loop
of the Props/Explore workspace.  It is \emph{not} the property \emph{theory}.
For the first-principles derivations of NRTL, Wilson, UNIQUAC, UNIFAC, the cubic
equations of state (SRK, Peng--Robinson), the electrolyte models (Pitzer,
eNRTL, aqueous speciation), Henry's law, Rackett, Ambrose--Walton, sublimation
and the NASA--CEA Gibbs-of-formation data, see the \textbf{Theory Guide} (the
in-GUI model panels deep-link straight into it).

If thermodynamics felt slippery the first time --- the formalism pristine, yet
somehow never self-sufficient when you sat down with real data --- you are in
good company: Sommerfeld, who reshaped quantum physics, said he never wrote his
thermodynamics book because he never felt he had understood the subject.  The
difficulty is real, but it is not formal incompleteness.  It is that an
\emph{axiomatic-looking} macroscopic theory rests, underneath, on emergent
statistical facts and on parameters that must be \emph{measured}, not derived.
The confusion is the seam between the two showing through.  This guide does not
hide that seam --- it makes it the organising idea.

\subsection{The target: where theory grips}

Picture property prediction as a target (Figure~\ref{fig:bullseye}): pristine at
the core, increasingly empirical toward the rim.

\begin{figure}[h]
\centering
\begin{tikzpicture}[font=\sffamily]
  % Concentric shells: pristine (white) core out to empirical (grey) rim.
  \fill[gray!28,draw=gray!55] (0,0) circle (5cm);
  \fill[gray!20,draw=gray!55] (0,0) circle (4cm);
  \fill[gray!13,draw=gray!55] (0,0) circle (3cm);
  \fill[gray!7, draw=gray!55] (0,0) circle (2cm);
  \fill[white,  draw=gray!55] (0,0) circle (1cm);
  % Ring titles (upper band of each shell).
  \node[font=\bfseries\small,align=center] at (0,0)   {The four\\laws};
  \node[font=\small,align=center]          at (0,1.5) {Formalism};
  \node[font=\bfseries\small,text=accent]  at (0,2.5) {Models};
  \node[font=\small,align=center]          at (0,3.5) {Data};
  \node[font=\small,align=center]          at (0,4.5) {Process \& numerics};
  % Descriptors (lower band) so the map reads without the caption.
  \node[font=\scriptsize\itshape,align=center]         at (0,-1.45) {$H,S,G$, Maxwell,\\equilibrium};
  \node[font=\scriptsize\itshape,align=center,text=accent] at (0,-2.45) {EoS $\cdot$ $\gamma$ $\cdot$ kinetics};
  \node[font=\scriptsize\itshape,align=center]         at (0,-3.45) {fitted parameters,\\finite range};
  \node[font=\scriptsize\itshape,align=center]         at (0,-4.55) {flash, column, recycle};
\end{tikzpicture}
\caption{The layers of property prediction. The two inner shells follow from the
laws alone. The accent shell --- \emph{constitutive models} --- is the first
fault line: the laws say \emph{that} fugacities must match across phases, never
\emph{how} to compute one for a real mixture. From there outward, the theory is
necessary but no longer sufficient.}
\label{fig:bullseye}
\end{figure}

\begin{itemize}[leftmargin=1.4em,itemsep=2pt]
\item \textbf{The four laws (core).}\quad The zeroth, first, second and third
  laws define temperature, internal energy, entropy, and --- via Nernst --- an
  absolute zero of entropy. Parameter-free.
\item \textbf{Formalism.}\quad Pure manipulation yields the potentials $H$, $G$,
  $A$, the Maxwell relations, the equilibrium criteria (equality of fugacity
  across phases, $\Delta G=-RT\ln K$ for reaction) and the phase rule. Still
  derivable from nothing but the laws.
\item \textbf{Constitutive models --- where it first grips.}\quad The laws
  require fugacities to be equal across phases; they do not tell you how to
  \emph{compute} a fugacity for a real mixture. You must supply a model they do
  not give you --- an equation of state, an activity model, a kinetic law. In
  Choupo: \code{idealGas}/\code{SRK}/\code{PR}; \code{ideal}/\code{NRTL}/%
  \code{Wilson}/\code{UNIFAC}.
\item \textbf{Data --- ``go to the laboratory.''}\quad Those models carry
  parameters thermodynamics cannot predict: $T_c$, $P_c$, $\omega$, the Antoine
  constants, $C_p^\circ(T)$, $\Delta H_f^\circ$, $S^\circ$, the NRTL $a_{ij}$.
  Each is measured or estimated, and each has a finite validity range. In
  Choupo, these are the \code{.dat} files.
\item \textbf{Process \& numerics.}\quad Where the predicted numbers are
  consumed --- flash, distillation, recycle --- and grip a second time, now
  numerically (multiple EoS roots, phase stability, initialisation).
\end{itemize}

\begin{keyidea}
Sommerfeld's difficulty is this picture: a theory that \emph{looks} axiomatic to
the rim, but stops being self-sufficient at the accent shell. The honest move is
not to paper over the seam between rigorous derivation and fitted correlation ---
it is to show exactly where it runs.
\end{keyidea}

\subsection{The glass-box promise}

Because the outer shells are empirical, every number Choupo predicts should
travel with two things attached:

\begin{itemize}[leftmargin=1.4em,itemsep=2pt]
\item \textbf{Provenance} --- which method produced it (rigorous derivation,
  fitted correlation, or group-contribution estimate) and the source of each
  parameter (a measured datum or an estimate).
\item \textbf{Validity awareness} --- whether a correlation is \emph{extrapolating
  beyond its fitted range}, with a warning when it is, rather than a confident
  number where none is warranted.
\end{itemize}

\begin{intuition}
Ask Choupo for the vapour pressure of nitrogen at room temperature and it
\emph{refuses}: $\mathrm{N_2}$ is supercritical there (its critical temperature
is $-147\,$\textdegree C), and a saturation pressure does not exist above the
critical point. The $P^{\mathrm{sat}}$ curve simply \emph{breaks} at $T_c$
instead of extrapolating Antoine into a meaningless ${\sim}600\,$bar. That
refusal is the whole promise in miniature: the model declines to invent a number
the physics forbids.
\end{intuition}

\subsection{Where we are heading}

A single property over a wide temperature span is not one problem --- it is
several physical regimes stitched together, and your confidence in any predicted
number rises and collapses as you cross them. Figure~\ref{fig:ribbon} is the
destination of this guide: the vapour pressure of a
\textbf{water\,/\,ethanol\,/\,benzene} mixture from $-200$ to $1000\,$\textdegree C,
drawn not as a curve but as a band of \emph{confidence}.

\begin{figure}[h]
\centering
\begin{tikzpicture}[font=\sffamily]
  % x = (T + 200) * 0.0125  -> -200 maps to 0, +1000 maps to 15.
  \def\barh{0.85}
  \fill[gray!12]   (0,0)     rectangle (2.25,\barh);   % no data (cryogenic)
  \fill[accent!78] (2.25,0)  rectangle (3.75,\barh);   % measured
  \fill[gray!40]   (3.75,0)  rectangle (7.175,\barh);  % fitted -> extrapolating
  \fill[gray!12]   (7.175,0) rectangle (15,\barh);     % supercritical / decomposes
  \draw[gray!55]   (0,0) rectangle (15,\barh);
  % Band labels.
  \node[font=\scriptsize,align=center]          at (1.125,\barh+0.34) {no data\\\itshape(all solid)};
  \node[font=\scriptsize,text=white]            at (3.0,\barh/2)      {measured};
  \node[font=\scriptsize,align=center]          at (5.46,\barh/2)     {fitted $\rightarrow$ extrapolating};
  \node[font=\scriptsize,align=center]          at (11.1,\barh/2)     {supercritical / decomposes --- \itshape does it exist?};
  % Critical-temperature wall (where the property ceases to exist).
  \draw[densely dashed,accent,thick] (7.175,-0.15) -- (7.175,\barh+0.7);
  \node[font=\scriptsize,text=accent,align=center,above] at (7.175,\barh+0.7) {water $T_c$\\\itshape no $P^{\mathrm{sat}}$ beyond};
  % Temperature ticks.
  \foreach \t/\x in {-200/0, 0/2.5, 100/3.75, 1000/15} {
    \draw (\x,0) -- (\x,-0.14);
    \node[font=\scriptsize,below] at (\x,-0.14) {\t};
  }
  \node[font=\scriptsize,below] at (7.5,-0.62) {temperature ($^\circ$C)};
\end{tikzpicture}
\caption{The same property, five honesties. A narrow window where we actually
have data; degradation outward as the correlation extrapolates; a hard
\emph{wall} at the critical temperature where the property ceases to exist (the
nitrogen refusal, generalised); and a cryogenic desert where every species has
frozen solid (ethanol last, at $-114\,$\textdegree C) and there is simply no
data. The theory runs to both edges; reality and data do not.}
\label{fig:ribbon}
\end{figure}

We build each piece of that picture in the chapters that follow --- estimating a
missing component (\S\ref{ch:joback}), curating the mixture models against lab
points (\S\ref{ch:pairs}), and testing whether the data itself is even consistent
(\S\ref{ch:consistency-props}). We read that span regime by regime at the end of
this introduction (\S\ref{sec:span}), and return to solve it numerically, end to
end, at the close.

\subsection{Anatomy of a component: where the numbers come from}

Open any \code{.dat} and you see a flat list of numbers. The natural question ---
the one this whole guide turns on --- is: \emph{which of these are independent
inputs, and which derived properties does the engine actually build from them?}
Let us read \file{acetone.dat} field by field and follow each number to the
property it feeds and to the line of code that consumes it.

\begin{center}\footnotesize
\begin{tabularx}{\linewidth}{@{}l>{\raggedright\arraybackslash}X@{}}
\toprule
\textbf{Field} & \textbf{What it is, and what it builds} \\
\midrule
\multicolumn{2}{@{}l}{\itshape Identity} \\
\code{MW} & molar mass; converts molar $\leftrightarrow$ mass everywhere. \\
\addlinespace[2pt]
\multicolumn{2}{@{}l}{\itshape Caloric anchors --- \code{standardThermochemistry}} \\
\code{dHf\_298} $\Delta H_f^\circ$ & the \emph{position} of the enthalpy scale: bond energy $C_p$ knows nothing about (combustion calorimetry). \\
\code{s\_298} $S^\circ$ & the \emph{absolute} (third-law) entropy at 298\,K --- a true zero-referenced value, not a difference. \\
\addlinespace[2pt]
\multicolumn{2}{@{}l}{\itshape Caloric shape --- \code{idealGasHeatCapacity}} \\
\code{coefficients} $C_p^\circ(T)$ & the master correlation: integrated to give the \emph{temperature shape} of $H$, $S$, $G$ (below). \\
\addlinespace[2pt]
\multicolumn{2}{@{}l}{\itshape Corresponding-states trio} \\
\code{Tc}, \code{Pc} & critical point: set the \emph{size} of the cubic-EoS attraction term $a_c$, and the reduced temperature $T_r=T/T_c$. \\
\code{omega} $\omega$ & acentric factor: sets the \emph{temperature shape} of that attraction, $\alpha(T_r;\omega)$ --- how non-spherical / polar the molecule is. \\
\addlinespace[2pt]
\multicolumn{2}{@{}l}{\itshape Phase change} \\
\code{Tb}, \code{HvapTb} & boiling point + latent heat at $T_b$; with \code{Tc} drive Watson's $\Delta H_{\mathrm{vap}}(T)$. \\
\code{Vliq} & liquid molar volume (Poynting, liquid density). \\
\addlinespace[2pt]
\multicolumn{2}{@{}l}{\itshape Correlations \& recipes} \\
\code{vaporPressure} & Antoine $\log_{10}P^{\mathrm{sat}}=A-B/(T+C)$, with its \code{Trange}; the saturation curve. \\
\code{groups} & the molecular decomposition a group-contribution estimator / UNIFAC reads (the curation recipe, \S\ref{ch:joback}). \\
\code{diffusionVolume} & Fuller gas diffusivity $\Rightarrow$ Lewis number (wet-bulb). \\
\bottomrule
\end{tabularx}
\end{center}

\subsubsection*{The caloric engine: \texorpdfstring{$H$, $S$, $G$}{H, S, G} from \texorpdfstring{$C_p$}{Cp}}

This is the mechanism behind ``other properties that derive from here'' --- and
it is worth seeing \emph{how one arrives at $S(T)$}, since that is the step that
trips everyone up. Start from the laws, at constant pressure. Heat capacity
\emph{is} the temperature slope of enthalpy, $C_p=(\partial H/\partial T)_P$, so
\[
dH = C_p\,dT.
\]
For entropy, the second law gives $dS=\delta q_{\mathrm{rev}}/T$; at constant
pressure the reversible heat is $\delta q_{\mathrm{rev}}=C_p\,dT$, hence
\[
dS = \frac{C_p}{T}\,dT.
\]
That single step --- dividing by $T$ --- is the whole origin of the $\int C_p/T$.
Integrate each from a reference temperature and add the 298\,K anchor; Choupo
stores $C_p^\circ(T)$ and the two anchors (\emph{not} tables of $H$, $S$) and
evaluates:
\begin{align}
H^\circ(T) &= \underbrace{\Delta H_f^\circ}_{\text{position}} + \int_{298.15}^{T} C_p^\circ(T')\,dT'
   &&\text{\footnotesize(\code{Component::h\_pure\_ig}, \file{Component.cpp})}\\
S^\circ(T) &= \underbrace{S^\circ_{298}}_{\text{absolute}} + \int_{298.15}^{T} \frac{C_p^\circ(T')}{T'}\,dT'
   &&\text{\footnotesize(\code{Component::s\_pure\_ig})}\\
G^\circ(T) &= H^\circ(T) - T\,S^\circ(T)
\end{align}
For a mixture the engine adds the entropy of mixing $-R\sum y_i\ln y_i$ and the
pressure term $-R\ln(P/P_{\mathrm{ref}})$ (\code{ThermoPackage::S\_ig}). That is
the \emph{entire} ideal-gas caloric model --- no hidden tables.

These integrals are not done numerically. For a polynomial $C_p$ the primitive
is elementary --- it was solved \emph{once, on paper} --- so each call evaluates
a closed form,
\[
\int_{T_{ref}}^{T}\!\frac{C_p^\circ}{T'}\,dT'
   = a_0\ln\frac{T}{T_{ref}} + \sum_{k\geq1}\frac{a_k}{k}\bigl(T^k-T_{ref}^k\bigr),
\]
exact to machine precision in a handful of operations (\code{PolynomialCp::S}).
The standard correlations all share this property (NASA, Aly--Lee, Shomate have
closed primitives too); Choupo never runs a quadrature loop for $C_p$.

\begin{example}[Acetone $H$ and $S$ at 400\,K --- by hand, then by Choupo]
With $C_p^\circ(T)=6.301+0.2606\,T-1.253\!\times\!10^{-4}T^2+2.038\!\times\!10^{-8}T^3$
(units J\,mol$^{-1}$K$^{-1}$), integrating from 298.15 to 400\,K:
\[
\int C_p^\circ\,dT' = 8431~\text{J/mol},\qquad
\int \frac{C_p^\circ}{T'}\,dT' = 24.19~\text{J/(mol·K)}.
\]
So $H^\circ(400)=-217100+8431=\mathbf{-208669}$~J/mol and
$S^\circ(400)=295.30+24.19=\mathbf{319.49}$~J/(mol·K). Running
\code{choupoProps} on a one-point scan returns exactly
$H_{\mathrm{ig}}=-208669$~J/mol, $S_{\mathrm{ig}}=319.49$~J/(mol·K). The number
on the page is the number in the code --- and at 298.15\,K both integrals vanish,
so $H=\Delta H_f^\circ$ and $S=S^\circ_{298}$, the anchors laid bare.
\end{example}

\begin{keyidea}
$C_p$ gives the \emph{shape}; the formation values give the \emph{position}.
That is why the \code{.dat} lists $C_p^\circ(T)$ \emph{and} $\Delta H_f^\circ$,
$S^\circ$ separately --- not redundancy. And note the asymmetry: entropy has an
absolute zero for free (third law: $S\!\to\!0$ for a perfect crystal at
$T\!=\!0$), so $S^\circ_{298}$ is a genuine absolute; enthalpy has \emph{no}
natural zero, so $\Delta H_f^\circ$ must be \emph{measured} and supplied.
\end{keyidea}

\subsubsection*{What the critical properties and acentric factor actually do}

$T_c$, $P_c$, $\omega$ do nothing in the caloric equations above --- they enter
only when you ask for \emph{real-fluid} behaviour through a cubic equation of
state. There the attraction term is $a(T)=a_c\,\alpha(T_r)$, with $a_c$ fixed by
$T_c,P_c$ and the temperature shape fixed by $\omega$ (Soave/SRK):
\[
\alpha(T) = \bigl[\,1 + m(1-\sqrt{T_r})\,\bigr]^2,\quad
m = 0.48508 + 1.55171\,\omega - 0.15613\,\omega^2,\quad T_r=\tfrac{T}{T_c}
\]
(\code{SRK.cpp}; Peng--Robinson uses the analogous $\kappa(\omega)$). This
$a(T)$ is what produces the departure functions $H-H^\circ$, $S-S^\circ$ and the
compressibility $Z$ --- the real-gas correction layered on top of the ideal-gas
caloric result. The same $T_c,T_b$ also drive Watson's latent heat,
$\Delta H_{\mathrm{vap}}(T)=\Delta H_{\mathrm{vap}}(T_b)\,
[(T_c-T)/(T_c-T_b)]^{0.38}$ (\code{Component.cpp}).

\begin{intuition}
So why does \file{acetone.dat} carry $\omega$ at all if you are running the
default \code{idealGas}? Because the data is \emph{model-agnostic} and the model
decides which fields wake up. Under \code{idealGas} there is no departure, and
$\omega,T_c,P_c$ are simply never read. Select \code{SRK} for a high-pressure
compressor and the very same numbers come alive. The \code{.dat} is the
laboratory record; the model is the question you ask of it.
\end{intuition}

\subsection{From \texorpdfstring{$G$}{G} to equilibrium: what \texorpdfstring{$\Delta G^\circ$}{dG0} is (and is not)}

The single most common confusion in this whole subject is $\Delta G^\circ$
versus $\Delta G$. It is worth settling once, because $G$ is what the reactors
minimise and it is built from exactly the $H$ and $S$ above.

\textbf{First, the $^\circ$.}\quad It means \emph{reference state} --- each
species pure, ideal gas, at $1$\,bar --- but \emph{at the reaction temperature}
$T$. It does \emph{not} mean ``at 298\,K.'' So $\Delta G^\circ=\Delta G^\circ(T)$
varies with temperature; $\Delta G^\circ(298)$ is just one point on a curve.

\textbf{Then, the two are not the same thing.}
\begin{itemize}[leftmargin=1.4em,itemsep=2pt]
\item $\Delta G^\circ$ is the Gibbs change if reactants $\rightarrow$ products
  with \emph{everything in its standard state}. It is a \emph{fixed number} --- a
  property of the reaction at $T$.
\item $\Delta G$ is the \emph{actual} Gibbs change at the real composition (the
  real partial pressures). It is the \emph{driving force}.
\end{itemize}
The reaction quotient $Q$ bridges them, and equilibrium is where the driving
force --- not the reference number --- vanishes:
\[
\Delta G = \Delta G^\circ + RT\ln Q,
\qquad
\Delta G = 0 \text{ at equilibrium} \;\Rightarrow\;
\boxed{\;\Delta G^\circ = -RT\ln K\;}
\]

\begin{keyidea}
At equilibrium it is $\Delta G$ that is zero, \emph{never} $\Delta G^\circ$.
$\Delta G^\circ$ stays fixed and merely sets \emph{where} the equilibrium sits
(the value of $K$); the system gets there by shifting $Q$ until $\Delta G=0$. And
$\Delta G^\circ>0$ does not mean ``no reaction'' --- it means $K<1$, i.e.\ partial
conversion. Only the sign and size of $K$ change.
\end{keyidea}

\textbf{Where the number comes from} --- straight back to the caloric engine:
\[
\Delta G^\circ(T) = \Delta H^\circ(T) - T\,\Delta S^\circ(T)
   = \sum_i \nu_i\, g_i^\circ(T),
\qquad g_i^\circ(T) = h^{\mathrm{ig}}_i(T) - T\,s^{\mathrm{ig}}_i(T).
\]
Note the subtlety that confuses everyone: $h^{\mathrm{ig}}$ carries the
\emph{formation} enthalpy ($\Delta H_f^\circ$, an elements datum) while
$s^{\mathrm{ig}}$ is the \emph{absolute} (third-law) entropy. Mixing the two
references looks wrong --- but it is exactly right, because in a \emph{balanced}
reaction the element contributions cancel through $\sum_i\nu_i$. That is precisely
why the \code{.dat} carries $\Delta H_f^\circ$ \emph{and} $S^\circ$ separately:
each plays a different role, and they combine correctly only across a balanced
stoichiometry (\code{Component::g\_pure\_ig}; consumed in
\code{PFR.cpp}/\code{CSTR.cpp} as $K=\exp(-\Delta G^\circ/RT)$ and in the Gibbs
reactor's direct $G$-minimisation).

\smallskip
\noindent\textbf{How $K$ moves with $T$ (Gibbs--Helmholtz).}\quad
$\Delta G^\circ(T)$ is not a frozen datum --- it has its own temperature law, and
that law is a workhorse relation worth knowing by name. Gibbs--Helmholtz states
$\left[\partial(G/T)/\partial T\right]_P = -H/T^2$; applied to
$\Delta G^\circ=-RT\ln K$ it collapses to the van~'t~Hoff equation,
\[
\boxed{\;\frac{d\ln K}{dT} = \frac{\Delta H^\circ}{RT^2}\;}
\]
which is just Le~Chatelier made quantitative: an exothermic reaction
($\Delta H^\circ<0$) has $K$ \emph{falling} as $T$ rises. In Choupo it doubles as
a glass-box self-check --- the $K(T)$ the reactors build from $g^\circ(T)$ at each
temperature must show exactly this slope, or the enthalpy and the equilibrium
constant are disagreeing with each other.

\begin{intuition}
These relations are not ornaments. Gibbs--Helmholtz is \emph{how the equilibrium
moves with temperature} (above); its sibling, the \textbf{Gibbs--Duhem} equation
$\sum_i x_i\,\mathrm{d}\ln\gamma_i = 0$ (constant $T,P$), says the activity
coefficients of a mixture are \emph{not independent}. That single constraint is
exactly why measured VLE data can be \emph{tested} for thermodynamic consistency
--- and why an activity model is built from one $G^{\mathrm{E}}$ function rather
than fitting each $\gamma_i$ freely. It is the engine of
\S\ref{ch:consistency-props}.
\end{intuition}

\subsection{A worked span: water\,/\,ethanol\,/\,benzene, \texorpdfstring{$-200$}{-200} to \texorpdfstring{$1000\,$\textdegree C}{1000 C}}
\label{sec:span}

Put the three together and sweep the temperature. This one mixture crosses
\emph{every} regime of property prediction, and is the honest way to see both how
properties are obtained \emph{and} where the models give out --- which is exactly
the colour band of Figure~\ref{fig:ribbon}, now read regime by regime. (Critical
temperatures: ethanol $241\,$\textdegree C, benzene $289\,$\textdegree C, water
$374\,$\textdegree C; melting points: ethanol $-114$, water $0$, benzene
$5.5\,$\textdegree C.)

\begin{center}\footnotesize
\begin{tabularx}{\linewidth}{@{}p{2.45cm}>{\raggedright\arraybackslash}X>{\raggedright\arraybackslash}X@{}}
\toprule
\textbf{Regime} & \textbf{How a property is obtained} & \textbf{Where the model gives out} \\
\midrule
$-200$ to $\sim0\,$\textdegree C \newline {\scriptsize solids (SLE)}
 & melting points $T_m$, $\Delta H_{\mathrm{fus}}$, solid $C_p$; freezing-point depression; sublimation $P^{\mathrm{sat}}$
 & \textbf{low-T edge.} Below every Antoine \code{Trange}: $P^{\mathrm{sat}}$ would extrapolate blindly. Solid-phase data is sparse --- a cryogenic \emph{data desert}. \\
\addlinespace
$\sim0$ to $\sim65\,$\textdegree C \newline {\scriptsize two liquids (LLE)}
 & activity model (\code{NRTL}/\code{UNIFAC}) for $\gamma$; water/benzene nearly immiscible, ethanol the cosolvent
 & VLE-fitted parameters do \emph{not} serve LLE; \code{Wilson} cannot represent a liquid split at all; without $T$-dependent parameters the miscibility gap appears/vanishes wrongly. \\
\addlinespace
$\sim65$ to $\sim100\,$\textdegree C \newline {\scriptsize boiling (VLLE)}
 & Antoine $P^{\mathrm{sat}}$ $+$ $\gamma$--$\phi$ $+$ three-phase flash (Gibbs minimisation); ethanol/water + ternary azeotrope ($\sim64.9\,$\textdegree C)
 & azeotrope sensitivity --- a small $\gamma$ error moves the column design; bubble-point (Wang--Henke) distillation is unstable \emph{through} an azeotrope (use \code{method simultaneous}). \\
\addlinespace
$\sim100$ to $\sim290\,$\textdegree C \newline {\scriptsize vapour $\to$ near-critical}
 & cubic EoS (\code{SRK}/\code{PR}) with $k_{ij}$; departure functions $H-H^\circ$, $S-S^\circ$, $Z$
 & cubics degrade as the critical region is approached; $k_{ij}$ often needs to be made temperature-dependent over so wide a span. \\
\addlinespace
$\sim290$ to $374\,$\textdegree C \newline {\scriptsize mixed super/sub-critical}
 & the method itself must switch $\gamma$--$\phi$ $\rightarrow$ $\phi$--$\phi$ (one EoS for every phase)
 & above a species' $T_c$ the activity-coefficient picture loses meaning; no single $\gamma$ model spans the switch; near-critical accuracy is poor. \\
\addlinespace
$374$ to $1000\,$\textdegree C \newline {\scriptsize supercritical $+$ reaction}
 & $\phi$--$\phi$ EoS; wide-range $C_p^\circ(T)$ for the caloric terms
 & \textbf{high-T edge.} $P^{\mathrm{sat}}$ ceases to exist above $T_c$ (Choupo \emph{refuses} it, \S1.2); and the species thermally \emph{decompose} --- the fixed-component assumption fails: do they still exist? \\
\bottomrule
\end{tabularx}
\end{center}

The two edges --- the ones you asked to watch --- fail for opposite reasons but
teach the same lesson. At the \textbf{cold edge} there is simply no data: every
correlation is fitted to a window that starts well above $73\,$K, the species are
frozen, and a confident number there would be invention, not prediction. At the
\textbf{hot edge} the theory keeps producing numbers --- Antoine will gladly
report a vapour pressure at $800\,$\textdegree C --- but the \emph{physics} has
moved on: past each $T_c$ no saturation exists, and past a few hundred degrees
the molecules are no longer the molecules in the \code{.dat}. In both cases the
correct output is a \emph{flag}, not a figure.

\begin{pitfall}
No single model spans this axis. A run that quietly uses one $\gamma$ model and
one cubic from $-200$ to $1000\,$\textdegree C is not ``general'' --- it is
extrapolating through three regime changes in silence. The wide span is not one
problem; it is five problems wearing one temperature axis.
\end{pitfall}

\begin{intuition}
What Choupo ships \emph{today} covers the middle of this span honestly: cubic
EoS, \code{ideal}/\code{NRTL}/\code{Wilson}/\code{UNIFAC} activity, Antoine
$P^{\mathrm{sat}}$, Watson $\Delta H_{\mathrm{vap}}$, three-phase VLLE by Gibbs
minimisation, and azeotropic columns via the simultaneous MESH solver. It does
\emph{not} (yet) carry association models (PC-SAFT/CPA) for the hydrogen-bonding
non-ideality, nor electrolyte or wide-range near-critical packages. So for the
H-bonding mid-span and the two edges, Choupo's job is not to fake accuracy it
does not have --- it is to predict where it can and to \emph{say so} where it
cannot. That honesty is the product.
\end{intuition}

% =====================================================================
\section{The Choupo property architecture --- the complete map}
\label{ch:architecture}

Everything in this guide hangs off one architecture.  This chapter is the
complete map, stated once; every later chapter is a room in this house.

\subsection{The flow}

\begin{quote}
\texttt{constant/thermoPhysPropDict} \emph{declares} $\;\rightarrow\;$ the builder
\emph{loads, verifies and announces} (it never estimates) $\;\rightarrow\;$ the
engine \emph{computes} $\;\rightarrow\;$ the run log \emph{confesses}.
\end{quote}

Estimation is a curation-time act (the props bench writes a reviewable
\texttt{.dat}); assembly-time refusal is loud (a missing declared parameter
file, a $k_{ij}$ regressed for another cubic, a vapour method the liquid path
cannot serve).  The log then shows the whole assembly: the VLE world, the
reference rung of every group, and every pair parameter with its source file.

\subsection{The grammar --- the axes of a property package}

\begin{center}
\small
\begin{tabular}{@{}llp{6.6cm}@{}}
\toprule
\textbf{Axis} & \textbf{Slot} & \textbf{What it fixes}\\
\midrule
components & \texttt{components ( ... )} & identity; a salt's \texttt{dissociatesTo} names its ions\\
phases & \texttt{phases}, \texttt{phases/solid} data & which phases exist; $\rho_p$, $k_v$ for solids\\
model per PHASE & \texttt{equilibrium \{ formulation ...; liquid \{...\} vapour \{...\} \}} & ONE Gibbs surface per phase --- the \texttt{formulation} selects the VLE world\\
convention per GROUP & reference rungs (\texttt{referenceBasis}) & renormalisations of one surface: solvent on Raoult, solutes on Henry, ions on aqueous infinite dilution\\
correlation per COMPONENT & the component \texttt{.dat} (+ overrides) & pure-species data; legal for transport and pure correlations, never for the equilibrium surface\\
parameter per PAIR & \texttt{parameters \{ ...Pairs \{...\} \}} & declared FILE, verified at assembly, cited to its source\\
mixing per PROPERTY $\times$ PHASE & \texttt{mixing} slots + \texttt{effective\{\}} & how pures combine (Wilke, Grunberg--Nissan, van-der-Waals one-fluid); two-phase closures (a suspension's Thomas/Einstein viscosity) live in \texttt{effective\{\}}, because they are \emph{media closures}, not thermodynamic mixing\\
chemistry & \texttt{chemistry \{ salts (...); \}} & real equilibria with $K(T)$ and $\Delta H$\\
package per UNIT & unit \texttt{thermo \{\}} & the spatial axis: a unit selects ANOTHER whole package --- never a partial patch of single axes (that is the road to two surfaces in one phase); the model boundary is audited ($\Delta H$ at the seam)\\
phase equilibrium & \texttt{phaseEquilibrium \{\}} & for each coexisting phase pair, the imposed equality and the COMMON DATUM (e.g. \texttt{link muEquality; datum aqueousInfDil; transfer <pair-file>;}) --- the bridge between surfaces, first-class and declared, never implicit\\
\bottomrule
\end{tabular}
\end{center}

Provenance is an \emph{attribute of every value}, not an axis: each resolved
number carries a quality word
(\texttt{measured > fitted > correspondingStates > estimated > asserted}),
and an \texttt{asserted} value --- a pinned constant, a user polynomial, a user
table --- is refused without its written \texttt{why}.

\subsection{The four VLE worlds}

All selected by the \emph{liquid} method slot; each validated against measured
data in this guide:
\begin{center}
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{World} & \textbf{Liquid method} & \textbf{K-value} & \textbf{Natural home}\\
\midrule
$\gamma$--$\varphi$ & \texttt{activity.NRTL} (Wilson, UNIQUAC, UNIFAC) & $\gamma_i P^{sat}_i/P$ & organic mixtures\\
Henry & \texttt{solution.henryDilute} & $\gamma^*_i H_i/P$ & dissolved gases\\
$\varphi$--$\varphi$ & \texttt{eos.SRK} / \texttt{eos.PengRobinson} & $\varphi^L_i/\varphi^V_i$ & high pressure, near-critical\\
electrolyte & \texttt{electrolyte.pitzer} / \texttt{.eNRTL} & ion activities, osmotic & brines, salts, mixed solvent\\
\bottomrule
\end{tabular}
\end{center}

\subsection{The five component routes}

\begin{enumerate}
\item \textbf{Standard, by name} --- \texttt{data/standards/components/}
      (exact-name $O(1)$, every value cited, the engine refuses to write there).
\item \textbf{Local, by name} --- \texttt{data/local/components/} is the
      \emph{private} working tier (gitignored; shipped empty).  Records are
      runtime-usable, but every use is announced as
      \texttt{[local] UNVERIFIED}; a local record never shadows a standard
      one.  Integrity and provenance gates make a local record inspectable,
      not automatically authoritative.
\item \textbf{New, defined in the case} --- \texttt{myCase/constant/components/}:
      a new molecule (estimated and labelled), a new ion, a pseudo-component
      lump.  The case runs self-contained from day one; promotion into the
      standards tree is a \emph{requested, reviewed curation act} after
      validation.
\item \textbf{Standard + overlay} --- a partial case file merges field-by-field
      over the standard entry (sample-specific measured data, a new model
      block); the run announces the merge.
\item \textbf{Solids} --- solid reference rungs, solids on their own stream
      channel, and the iron rule: a dissolving salt's solid formation enthalpy
      is ION-DERIVED ($H_f^{solid}=\sum\nu_i h_{f,aq,i}-\Delta H_{soln}$),
      never a stored duplicate.
\end{enumerate}

Thus component resolution is \textbf{case-local overlay $>$ standard $>$
local (private, announced)}.  This is an evidence ladder, not merely a search
path: moving a file up the ladder requires review and validation, not a
rename.

\subsection{The ChemSep import route}

The ChemSep 8.3 importer (\texttt{bin/curate/chemsep\_to\_choupo.py}) is
deliberately reproducible and reviewable.  It runs CAS-based de-duplication
against the standard catalogue, then a structural/provenance audit and an
internal $P^{\mathrm{sat}}(T_b)$ closure check on every imported component
record --- a closure check on the imported constants and predictive
correlation, not an independent comparison with laboratory data.

The same import stages NRTL, UNIQUAC and Wilson binary-pair records, each
passing identity, direction, licence, source-digest and exact unit-conversion
checks.  Their source does not provide a captured temperature-validity range,
so they remain screening data and the runtime prints
\texttt{[local] UNVERIFIED} when one is selected.  Pair resolution is
\textbf{case-local/inline $>$ standard $>$ local $>$ announced ideal
default}; local data can therefore extend a case without silently changing
an approved standard.

The source boundary is intentional.  The unmodified third-party database and
its Artistic-2.0 licence are kept under \texttt{thirdParty/chemsep/}
(gitignored); the normalised, attributed records are written to the
\emph{private} \texttt{data/local/} tier --- they are never redistributed
with the public repository.  Literature values retain literature provenance;
Ambrose--Walton and other closures retain predictive provenance.  Missing
$C_p$, formation, electrolyte, adsorption, kinetic or transport data are not
inferred merely to make a record look complete.

The case additionally owns its \emph{equipment physics}: reaction kinetics in
\texttt{constant/reactions} (the heat of reaction always computed from the
formation datum and announced), per-machine measured curves (a dryer's
falling-rate data) in the unit's own folder, all inherited through the LOUD
fractal cascade (plant $\to$ sector $\to$ unit).

\subsection{The reference case}

The whole architecture is exercised by one designed flagship,
\texttt{tutorials/plant/lithiumBrinePlant} (under construction): salar brine
through Pitzer ponds and crystallisers, NF membrane Mg/Li separation,
solvent extraction with a student-defined extractant (the $\gamma$--$\varphi$
world and the audited Pitzer$\leftrightarrow$NRTL seam), electrodialysis,
60-bar CO$_2$ carbonation (the $\varphi$--$\varphi$ override and the Henry
world), a radical-kinetics burner satellite for thermal NO$_x$, batch
recrystallisation recipes, a control satellite, and the full outer ring
(fitting, sweeps, optimisation, sizing, costing, economics).  Later chapters
refer to its sectors by name.

\subsection{The principle}

No silent crutch, anywhere: every value is cited, estimated-and-labelled,
overlaid-and-announced, or asserted-with-a-why --- and the run log is where the
whole assembly confesses.  There is no ultimate model; hunting with the cat is
honest engineering, hiding the cat is fraud.

\section{The thermodynamic chain: how every equation here is
derived}\label{ch:thermo-chain}
% =====================================================================

Every K-value, activity coefficient and Henry constant in this guide falls
out of ONE chain of ideas.  This section walks it once, so that no equation
below ever has to be taken on faith.  (Full derivations: the Theory Guide,
chapters on flash, activity models, electrolytes and Henry's law.)

\subsection{1. Equilibrium is a Gibbs minimum}
The second law says an isolated system maximises entropy; recast at the
conditions a process engineer actually controls --- fixed $T$ and $P$ --- it
says a system minimises its \textbf{Gibbs energy} $G$.  For a multicomponent,
multiphase system,
\[
   dG \;=\; -S\,dT + V\,dP + \sum_i \mu_i\,dn_i,
   \qquad
   \mu_i \equiv \left(\frac{\partial G}{\partial n_i}\right)_{T,P,n_{j\neq i}},
\]
and at the minimum, moving a mole of $i$ between phases must change nothing:
\[
   \boxed{\;\mu_i^\alpha = \mu_i^\beta\;}\quad\text{for every component, across
   every pair of phases.}
\]
That single boxed statement is ALL the physics; everything else is bookkeeping
to make it computable.

\subsection{2. Fugacity: the workable currency}
$\mu_i$ is awkward: its absolute value is unknowable and it dives to $-\infty$
as $x_i \to 0$.  Lewis (1901) defined the \textbf{fugacity} $f_i$ through
$\mu_i = \mu_i^\circ + RT\ln(f_i/f_i^\circ)$ --- an ``effective pressure''
that equals $y_iP$ in an ideal gas.  Because $\ln$ is monotonic, the
equilibrium condition becomes
\[
   \boxed{\;f_i^{\,\alpha} = f_i^{\,\beta}\;}
\]
--- equal fugacities, phase by phase, component by component.  Every VLE
``world'' in Choupo is just a different, self-consistent way of WRITING
$f_i$ for a phase.

\subsection{3. Two routes to a mixture fugacity}
\textbf{Route A --- an equation of state} (works for any phase the EoS can
represent).  Define the fugacity coefficient $\varphi_i \equiv f_i/(y_iP)$;
thermodynamics gives it exactly from the volumetric behaviour,
\[
   RT\ln\varphi_i \;=\; \int_V^{\infty}\!\!\left[\left(\frac{\partial
   P}{\partial n_i}\right)_{T,V,n_j}\!\! - \frac{RT}{V}\right] dV
   \;-\; RT\ln Z ,
\]
so ONE model of $P(T,V,\mathbf n)$ --- a cubic like SRK with its $a(T)$,
$b$ and the pair correction $k_{ij}$ --- yields $\varphi_i$ for BOTH the
vapour root and the liquid root of the same cubic.

\textbf{Route B --- an activity model} (liquids only).  Split the liquid
fugacity into a reference times a correction,
$f_i^L = \gamma_i\,x_i\,f_i^{\circ}$, where the \textbf{activity coefficient}
$\gamma_i$ comes from a model of the excess Gibbs energy
($G^E$: NRTL, UNIQUAC, Wilson, Pitzer, eNRTL) via
$RT\ln\gamma_i = \partial G^E/\partial n_i$.
The price of Route B is choosing the \textbf{reference} $f_i^\circ$ --- and
that choice is a CONVENTION, not physics:

\begin{center}
\begin{tabular}{lll}
\textbf{Convention} & \textbf{Reference $f_i^\circ$} & \textbf{Normalised so} \\[2pt]
symmetric (Raoult)   & pure liquid: $P^{\mathrm{sat}}_i$ ($\times\,\varphi_i^{sat}\cdot$Poynting at high $P$) & $\gamma_i \to 1$ as $x_i \to 1$ \\
unsymmetric (Henry)  & infinite dilution: $H_{i,\mathrm{solv}} = \lim_{x\to 0} f_i/x_i$ & $\gamma_i^{*} \to 1$ as $x_i \to 0$ \\
electrolyte (aqueous)& ion at infinite dilution, $\mathrm{H^+(aq)}=0$ datum & $\gamma_\pm \to 1$ as $m \to 0$ \\
\end{tabular}
\end{center}

A supercritical solute HAS no pure liquid, so its only definable anchor is
the Henry limit --- which is why dissolved H$_2$/N$_2$ take the unsymmetric
convention, never an Antoine curve extrapolated above $T_c$.

\subsection{4. One Gibbs surface per phase --- the legality rule}
Gibbs--Duhem, $\sum_i x_i\,d\ln\gamma_i = 0$ at constant $T,P$, ties every
component's $\gamma$ in a phase to every other's: they must all derive from
ONE $G^E$ (or one EoS) surface.  Hence the rule that governs Choupo's
\code{thermoPhysPropDict} grammar: \emph{models per phase; conventions per
component-group within the phase; correlations per component; parameters per
pair}.  Mixing two $\gamma$-models (or two cubics) in one phase violates
Gibbs--Duhem and is refused.  Mixing CONVENTIONS in one phase is legal ---
solvent symmetric, solutes unsymmetric --- because a convention only
renormalises the same surface.

\subsection{5. The four worlds' K-values, assembled}
Setting $f_i^L = f_i^V$ with each route/convention gives every K-value this
guide uses:
\[
\begin{array}{llll}
\textbf{world} & f_i^L & f_i^V & K_i = y_i/x_i \\[3pt]
\gamma\text{-}\varphi & \gamma_i x_i P^{\mathrm{sat}}_i & y_i\varphi_i P
  & \dfrac{\gamma_i\,P^{\mathrm{sat}}_i}{\varphi_i\,P} \\[8pt]
\text{Henry (KK/KI)} & \gamma_i^{*} x_i H_i(T)\,e^{\bar v_i^\infty(P-P_s)/RT}
  & y_i\varphi_i P
  & \dfrac{\gamma_i^{*} H_i\,\mathcal P}{\varphi_i P} \\[8pt]
\varphi\text{-}\varphi & x_i\varphi_i^{L}P & y_i\varphi_i^{V}P
  & \dfrac{\varphi_i^{L}}{\varphi_i^{V}} \\[8pt]
\text{electrolyte} & \multicolumn{3}{l}{\text{ion activities from Pitzer/eNRTL on the aqueous rung;}}\\
 & \multicolumn{3}{l}{\text{full speciation (mass action + charge balance) where the case demands it}}
\end{array}
\]
Each row is one \code{equilibrium.formulation}
(\code{gammaPhi}, \code{diluteSolution}, \code{phiPhi},
\code{electrolyteGammaPhi}); the run log announces which world is live.  All four
are validated in the tutorials against measured data --- van~Ness-consistent
VLE sets ($\gamma$-$\varphi$), Wiebe/Larson solubilities (Henry, $\sim$2\%),
the Stryjek N$_2$--CH$_4$ isotherm ($\varphi$-$\varphi$, $\sim$2\%), and the
Hamer--Wu / calorimetric electrolyte sets.

\begin{intuition}
The whole edifice is: $G$ minimum $\to$ equal $\mu$ $\to$ equal $f$ $\to$
pick HOW to write $f$ per phase (an EoS or a $G^E$ model + a reference
convention) $\to$ a K-value.  When an equation in this guide looks arbitrary,
walk it back up this chain --- it never is.
\end{intuition}

% =====================================================================
\section{Why a separate guide}
% =====================================================================

The Theory Guide answers \emph{how a model works} --- how a flash splits, how
an EoS predicts $Z$, how Levenberg--Marquardt minimises a residual.  This guide
answers a question that comes \emph{before} that: \emph{where do the numbers
that feed the models come from?}  A simulation is only as good as the $T_c$,
the NRTL pair, the heat of formation it stands on, and those are not handed
down --- they are estimated, fitted, and tested.

Choupo treats that as a first-class phase with its own binary (\code{choupoProps})
and its own lifecycle:

\begin{quote}
\textbf{CREATE} a component (from groups) \;$\rightarrow$\; \textbf{CURATE} the
mixture models (fit pairs, test data, choose by evidence) \;$\rightarrow$\;
\textbf{WIRE} the flowsheet \;$\rightarrow$\; \textbf{SIMULATE}.
\end{quote}

A \code{propsDict} of \code{operations} drives the first two stages.  A case may
have \emph{no} \code{flowsheetDict} at all (a pure properties case --- the
compound bench), or carry one and use props to curate the thermo it will
simulate.  Same files, same engine; only the centre of the GUI differs.

\begin{intuition}
The usual reflex is to click a property method and trust a badge.  Choupo's
stance (the ``see, then decide'' principle) is the opposite: you make the
estimate or the fit \emph{visible} --- the group build-up, the model curve
against the lab points, the consistency residual --- and decide from the
evidence in front of you.  Estimation has error; the point is to see it.
\end{intuition}

% =====================================================================
\section{Where curated data lives, and how a case selects it}\label{ch:property-package}
% =====================================================================

You estimated a $T_c$, fitted an NRTL pair, tested a heat of solution.  Two
questions follow: \emph{where does that number land}, and \emph{how does a case
use it}?  Choupo's answer is \emph{declarative property packages with
OpenFOAM-like explicit files}.

A curated value lands in a glass-box \code{.dat} in a tree organised by
\emph{what kind of thing} it is --- the intrinsic property of one substance, the
parameter of a pair, the equilibrium of a reaction:

\begin{lstlisting}
data/standards/
  components/<name>.dat            intrinsic, one substance (Tc, Antoine, Cp)
                                   + dissociatesTo { Na 1; Cl 1; }  [salts]
  species/<name>.dat               one modelSpecies file per species
                                   (Na, Cl, CO2aq, O2): formula + charge +
                                   medium-tagged thermo; never fed to a flowsheet
  phases/solid/<phase>.dat         crystal phase (rho_p, k_v): halite, sylvite
  chemistry/.../<name>.dat         an equilibrium: dissolution, speciation
  parameters/.../<i>-<j>.dat       a pair: NRTL, Pitzer, eNRTL
  methods/<name>.dat           a model declaration (+ its reference basis)
  (the system is declared INLINE in the case's constant/thermoPhysPropDict
   -- THE CENTRE; there is no shared package catalogue)
\end{lstlisting}

The arity rule decides \emph{which} file: a number that ``stands alone with the
molecule'' (the $T_c$ of water) lives with the component; a number that needs
two names (an NRTL pair) cannot, and goes to \code{parameters/}.  The Developer
Guide (\emph{the thermophysical property architecture}) gives the full layout
and the eight data kinds.

\paragraph{Declaring the system.}  A case does not list loose models and
parameters.  It \emph{declares} its thermophysical system --- one record that
names the components, the formulation, and the parameter files the run needs
(active chemistry lives beside it in \file{constant/chemistryDict}):

\begin{lstlisting}
recordType    thermophysicalPropertySystem;
schemaVersion 2;

components ( water NaCl );          // NaCl's dissociatesTo names the ions

equilibrium
{
    formulation electrolyteGammaPhi;

    aqueous
    {
        solvent            water;
        apparentComponents ( NaCl );  // stream carries the salt; the model
        activityModel { model Pitzer; }        //   activates its ions
        compositionBasis molality;
    }

    vapour { fugacityModel idealGas; }
}
% Pitzer pairs resolve by (cation,anion) name from parameters/; declared
% binaryParameters/binaryInteractions files load from their `source` paths
% and refuse when missing.
\end{lstlisting}

The flow is one sentence: \textbf{\code{thermoPhysPropDict} declares,
\code{ThermoPackageBuilder} assembles, \code{ThermoPackage} computes.}  The
declaration is the recipe; the builder loads what it names and mounts the
runtime object the unit operations consume.  An ordinary molecular case is
the same record degenerate --- no dissociation, no chemistry, a
\code{gammaPhi} block with an NRTL or an EoS slot.

\begin{intuition}
The boundary you must respect: \emph{curation estimates, runtime does not}.  You
may estimate, fit, and derive at the props bench --- and you \emph{see} it
(\S\ref{ch:joback}, \S\ref{ch:pairs}).  Once a case runs, the builder only
\emph{loads} existing case, standard or local records; it never invents a
missing property.  For activity models, an unresolved binary pair may use the
documented ideal default, but that fallback is announced and appears in the
provenance.  A required record with no defined fallback stops the run with a
clear message.  So: estimate, look, promote --- and inspect every local or
defaulted rung used by the package.
\end{intuition}

% =====================================================================
\section{Creating a component: Joback group contribution}\label{ch:joback}
% =====================================================================

When a component is missing from the catalogue and from the literature, you can
\emph{estimate} its pure-component constants from its molecular structure ---
the groups it is built from.  Choupo implements the \textbf{Joback} method
(Joback \& Reid, 1987; tabulated in Poling, Prausnitz \& O'Connell, \emph{The
Properties of Gases and Liquids}, 5th ed.), the \code{estimateComponent}
property operation.

\subsection{The method}

Each first-order group contributes a tabulated increment.  Summing the
increments and substituting into Joback's correlations gives the constants
(with $n_A$ the total number of atoms):

\begin{align}
T_b      &= 198.2 + \textstyle\sum_i n_i\,\Delta T_{b,i} \\
T_c      &= T_b \,\big/\, \big(0.584 + 0.965\,\Sigma\Delta T_c - (\Sigma\Delta T_c)^2\big) \\
P_c      &= \big(0.113 + 0.0032\,n_A - \Sigma\Delta P_c\big)^{-2} \quad[\text{bar}] \\
V_c      &= 17.5 + \textstyle\sum_i n_i\,\Delta V_{c,i} \quad[\text{cm}^3/\text{mol}] \\
\Delta H_f^{\circ}(298) &= 68.29 + \textstyle\sum_i n_i\,\Delta H_{f,i} \quad[\text{kJ/mol, ideal gas}] \\
\Delta H_{\text{vap}}(T_b) &= 15.30 + \textstyle\sum_i n_i\,\Delta H_{v,i} \\
C_p^{\text{ig}}(T) &= (\Sigma a - 37.93) + (\Sigma b + 0.210)T
                    + (\Sigma c - 3.91\!\times\!10^{-4})T^2
                    + (\Sigma d + 2.06\!\times\!10^{-7})T^3 .
\end{align}

The acentric factor $\omega$ is \emph{not} a group sum; Choupo derives it from
$(T_b, T_c, P_c)$ by the Lee--Kesler correlation, with $T_{br}=T_b/T_c$ and
$P_c$ in atm:
\begin{equation}
   \omega = \frac{-\ln P_c - 5.97214 + 6.09648/T_{br} + 1.28862\ln T_{br} - 0.169347\,T_{br}^6}
                 {15.2518 - 15.6875/T_{br} - 13.4721\ln T_{br} + 0.43577\,T_{br}^6}.
\end{equation}

A few first-order group increments (Joback \& Reid; the full table is in the
code, \code{src/propertyOps/EstimateComponent.cpp}):
\begin{center}\small
\begin{tabular}{lrrrr}
\toprule
group & $\Delta T_b$ & $\Delta T_c$ & $\Delta P_c$ & $\Delta H_f$ (kJ/mol)\\
\midrule
$-$CH$_3$        & 23.58 & 0.0141 & $-0.0012$ & $-76.45$ \\
$>$CH$_2$        & 22.88 & 0.0189 & $\ \ 0.0000$ & $-20.64$ \\
$-$OH (alcohol)  & 92.88 & 0.0741 & $\ \ 0.0112$ & $-208.04$ \\
$>$C$=$O (ketone)& 76.75 & 0.0380 & $\ \ 0.0031$ & $-133.22$ \\
aromatic $=$CH$-$& 26.73 & 0.0082 & $\ \ 0.0011$ & $\ \ 2.09$ \\
\bottomrule
\end{tabular}
\end{center}

\subsection{The structure is identity: \code{jobackGroups} and \code{bin/estimate}}

The molecular structure belongs \emph{with the substance}, like the formula:
a component's \code{.dat} declares its groups once,
\begin{lstlisting}
jobackGroups { CH3 2; ketone 1; }     // acetone = 2 x CH3 + 1 x >C=O
\end{lstlisting}
(the sibling of the UNIFAC \code{groups\{\}} block --- a different table, the
same idea), and estimation needs no case at all:
\begin{lstlisting}
bin/estimate acetone            # case-local first, then the standard catalogue
bin/estimate path/to/new.dat    # or an explicit file
\end{lstlisting}
The command runs the Joback engine on the component's own groups, prints the
full glass-box build-up (each group's increment, the closure chain, deviations
against any \code{reference} block), and drops \emph{one} stable proposal ---
\code{<name>.estimated.dat} --- next to the source, for you to review and
promote by hand.  It never writes into the frozen \code{data/standards/} tree,
and the solver still refuses to estimate at run time: estimation is a
\emph{curation} act with a paper trail, not a silent fill-in.

The \code{estimateComponent} operation remains the pedagogical form --- declare
the groups inline in a \code{propsDict} when you want the estimate itself to be
the lesson, as in the worked example below.

\subsection{Worked example: acetone}

Acetone, $\mathrm{CH_3\!-\!CO\!-\!CH_3}$, decomposes into $2\times\code{CH3}$ and
$1\times\code{ketone}$.  The operation prints the build-up and validates against
a \code{reference} block:

\begin{center}\small
\begin{tabular}{lrrr}
\toprule
Property & Joback & Reference (NIST) & Dev. \\
\midrule
$T_b$   & 322.1\,K        & 329.2\,K  & $-2.2\%$ \\
$T_c$   & 500.6\,K        & 508.1\,K  & $-1.5\%$ \\
$P_c$   & 48.0\,bar       & 47.0\,bar & $+2.2\%$ \\
$V_c$   & 209.5\,cm$^3$/mol& 209\,cm$^3$/mol & $\approx 0$ \\
$\omega$& 0.295           & 0.307     & $-3.9\%$ \\
$\Delta H_f^{\circ}$ & $-217.8$\,kJ/mol & $-217.1$\,kJ/mol & $+0.3\%$ \\
\bottomrule
\end{tabular}
\end{center}

Within a few percent on the criticals and almost exact on $\Delta H_f$ --- and
you \emph{see} the error, because the reference sits beside the estimate.

\begin{intuition}
\textbf{Know the method's limits.}  Joback is weak on $T_b$ of strongly
hydrogen-bonding species (ethanol's $T_b$ comes out $\sim$14\,K low), and it
yields \emph{no} Antoine / vapour-pressure parameters --- those need a fit
(\S\ref{ch:pairs}) or a corresponding-states correlation, a separate step.  An
estimate is a starting point you then anchor to data, never a final answer.
\end{intuition}

\begin{tutoriallink}
\file{tutorials/props/estimate/estimate\_acetone} --- the example above, groups in,
validated estimate out.
\end{tutoriallink}

% =====================================================================
\section{Curating mixture models: fitting binary pairs}\label{ch:pairs}
% =====================================================================

A pure-component set is not enough for a mixture: the non-ideality lives in the
\emph{binary interactions} (NRTL, Wilson).  Those parameters are regressed
against experimental VLE data by Levenberg--Marquardt --- the \code{fitParameters}
operation.

\subsection{The activity-coefficient models}

For a binary, the \textbf{NRTL} model gives the liquid-phase activity
coefficients from three adjustable numbers per pair --- $\tau_{12},\tau_{21}$
(temperature-dependent) and the non-randomness $\alpha$:
\begin{equation}
   \tau_{ij} = a_{ij} + \frac{b_{ij}}{T}, \qquad G_{ij} = \exp(-\alpha\,\tau_{ij}),
\end{equation}
\begin{equation}
   \ln\gamma_1 = x_2^2\!\left[\tau_{21}\!\left(\frac{G_{21}}{x_1+x_2 G_{21}}\right)^{\!2}
   + \frac{\tau_{12}\,G_{12}}{\left(x_2+x_1 G_{12}\right)^2}\right],
\end{equation}
and $\ln\gamma_2$ by swapping the indices.  So a pair carries four fittable
coefficients $(a_{12},b_{12},a_{21},b_{21})$ plus a fixed $\alpha$ (often $0.3$).
\textbf{Wilson} is the two-parameter alternative,
\begin{equation}
   \ln\gamma_1 = -\ln(x_1+\Lambda_{12}x_2)
   + x_2\!\left[\frac{\Lambda_{12}}{x_1+\Lambda_{12}x_2}
   - \frac{\Lambda_{21}}{x_2+\Lambda_{21}x_1}\right],
\end{equation}
which cannot represent a liquid--liquid split (use NRTL when one is suspected).

\subsection{The residual the fit minimises}

The measurable is the bubble temperature.  At a fixed pressure $P$ and liquid
composition $x$, the bubble point is the $T$ that closes modified Raoult,
\begin{equation}
   \sum_i x_i\,\gamma_i(x,T)\,P_i^{\text{sat}}(T) = P
   \label{eq:bubbleT}
\end{equation}
(a Newton-1D in $T$; $P_i^{\text{sat}}$ from Antoine).  Levenberg--Marquardt
then drives the pair coefficients $\theta=(a_{12},b_{12},a_{21},b_{21})$ to
minimise the sum of squared bubble-temperature residuals against the data,
\begin{equation}
   \min_{\theta}\ \sum_k \big[\,T_{\text{bub}}(x_k;\theta) - T_{k}^{\text{exp}}\,\big]^2 .
\end{equation}

\begin{intuition}
\textbf{A low residual is not a good fit.}  At a \emph{single} pressure
$\tau_{ij}=a_{ij}+b_{ij}/T$ is nearly linear in both $a_{ij}$ and $b_{ij}/T$
over the narrow $T$ range of one isotherm/isobar, so $a$ and $b$ are
\emph{correlated} --- many $(a,b)$ pairs fit equally well and the optimum is a
valley, not a point.  \code{fitParameters} forms $J^{\mathsf T}J$ at the optimum
and reports the parameter correlation + confidence intervals so you SEE this;
the Theory Guide's \emph{Levenberg--Marquardt} and \emph{What is fittable}
chapters derive the diagnostic.  Fit across pressures, or fix $b$, to break it.
\end{intuition}

\subsection{See, then decide: the model overlay}

Before fitting, look.  Declare one scan \code{operation} per candidate model
(ideal, NRTL, Wilson), emitting the $(x, T_{\text{bubble}}, y_{\text{eq}})$
trio, and an \code{experimental \{\}} entry that overlays them --- and, with a
\code{dataset}, draws the measured points on top.  The student sees which model
the data actually picks: for ethanol/water at 1\,atm the ideal curve misses the
azeotrope entirely while NRTL lands on it.

\begin{tutoriallink}
\file{tutorials/props/compare/compare\_vle\_etoh\_water} --- ideal vs.\ NRTL bubble/dew
curves overlaid with measured VLE points (Carey \& Lewis, 1932).
\end{tutoriallink}

\subsection{Fit, then promote}

When a model earns its keep, \code{fitParameters} regresses its coefficients to
the data and writes a proposal to
\file{constant/parameters/<model>/<pair>.fit-<date>.dat}.  Promotion is a
deliberate, visible act: you inspect the parity plot and the identifiability
statistics (confidence intervals, parameter correlation) the fit reports ---
never a ``good fit'' badge --- and copy the proposal into the case's
\code{constant/} when you trust it.

% =====================================================================
\section{Testing the data itself: consistency}\label{ch:consistency-props}
% =====================================================================

Before fitting a model \emph{to} data, test whether the data is even
thermodynamically consistent --- a fit to bad data is a polished wrong answer.
The \code{vleConsistency} operation does this in three steps.

\subsection{Step 1 --- activity coefficients straight from the data}

It does not assume a model.  From each measured $(x_1, y_1, T)$ at the run
pressure $P$, modified Raoult gives the activity coefficient of each component
directly:
\begin{equation}
   \gamma_i = \frac{y_i\,P}{x_i\,P_i^{\text{sat}}(T)},
   \qquad P_i^{\text{sat}}(T)\ \text{from Antoine}.
\end{equation}
These are \emph{experimental} $\gamma_i$ --- the test now asks whether they obey
thermodynamics.

\subsection{Step 2 --- the Gibbs--Duhem point test}

The Gibbs--Duhem equation is the constraint every real mixture must satisfy at
constant $T,P$:
\begin{equation}
   \sum_i x_i\,\mathrm{d}\ln\gamma_i = 0
   \quad\Longrightarrow\quad
   x_1\frac{\mathrm{d}\ln\gamma_1}{\mathrm{d}x_1}
   + x_2\frac{\mathrm{d}\ln\gamma_2}{\mathrm{d}x_1} = 0 \ \ (\text{binary}).
   \label{eq:gd}
\end{equation}
The operation evaluates the left-hand side at each interior point (central
differences in $x_1$) and reports its max and mean magnitude --- the
\emph{point residual}.  A consistent set keeps it near zero.  (Isobaric data
adds a small $-H^{\mathrm E}/RT^2 \cdot \mathrm{d}T/\mathrm{d}x$ term, dropped
here; stated as a caveat, not hidden.)

\subsection{Step 3 --- the Herington area test}

Integrate \eqref{eq:gd} across the whole composition range: for an isothermal
set Gibbs--Duhem forces the net area of $\ln(\gamma_1/\gamma_2)$ to vanish.
Herington's isobaric form compares that net area to the absolute area and
corrects for the boiling-range:
\begin{equation}
   D = 100\,\frac{\left|\displaystyle\int_0^1 \ln\dfrac{\gamma_1}{\gamma_2}\,\mathrm{d}x_1\right|}
                 {\displaystyle\int_0^1 \left|\ln\dfrac{\gamma_1}{\gamma_2}\right|\mathrm{d}x_1},
   \qquad
   J = 150\,\frac{T_{\max}-T_{\min}}{T_{\min}},
\end{equation}
(both integrals by the trapezoid rule over the sorted points).  The set
\textbf{passes} when $|D-J| < 10$.  The GUI's consistency card plots
$\ln\gamma_1,\ \ln\gamma_2$ and $\ln(\gamma_1/\gamma_2)$ vs $x_1$ with the net
area shaded --- you SEE the $\pm$ areas cancel, not just the number.

Finally the infinite-dilution coefficients $\gamma_1^\infty,\gamma_2^\infty$
(the dilute-end limits) are reported --- the most stringent point of any model
later fitted to the set.

\begin{tutoriallink}
\file{tutorials/props/compare/compare\_vle\_etoh\_water} --- the \code{vleConsistency} op
on the Carey--Lewis ethanol/water data (Herington $|D-J|=4.6<10$, PASS).
\end{tutoriallink}

\vfill
\noindent\rule{\linewidth}{0.3pt}
\noindent\small This guide grows with the props pillar.  Planned: a second
group-contribution method (Constantinou--Gani) to compare against Joback;
group contribution for liquid viscosity; a structure (SMILES) front-end for the
compound bench.
\end{document}
