\documentclass[11pt,a4paper]{article}
\newcommand{\docTitle}{Tutorials Guide}
\input{preamble.tex}
\usepackage{adjustbox}

% ---------------------------------------------------------------------------
%  Per-tutorial entry macros (the generated files use these).
% ---------------------------------------------------------------------------
\newcommand{\tutentry}[2]{%
  \par\medskip\noindent
  \addcontentsline{toc}{subsubsection}{\texttt{#1}}%
  {\large\bfseries\texttt{#1}}\\[1pt]
  {\footnotesize\color{gray}\texttt{#2}}\par\smallskip}
\newcommand{\tutwhat}[1]{\noindent\textbf{What it does:}\ #1\par\smallskip}
\newcommand{\tutstory}[1]{\noindent{\small #1}\par}
\newenvironment{tutscheme}{%
  \par\smallskip\noindent\begin{center}\begin{adjustbox}{max width=\linewidth, max height=7cm}}{%
  \end{adjustbox}\end{center}\par\smallskip}
\newenvironment{tutnarrative}{\par\noindent\textbf{The story:}\ \small}{\par\normalsize\smallskip}
\newcommand{\tutreadme}[1]{\par\noindent{\small\emph{#1}}\par}
\newenvironment{tutgolden}{%
  \par\smallskip\noindent\textbf{What to expect (golden-locked):}\par\nopagebreak\vspace{2pt}
  \begin{footnotesize}\begin{tabular}{@{}llll@{}}\toprule
  kind & unit/stream & key & expected \\ \midrule}{%
  \bottomrule\end{tabular}\end{footnotesize}\par\medskip}
\newcommand{\goldrow}[4]{\texttt{#1} & \texttt{#2} & \texttt{#3} & \texttt{#4} \\}
\newcommand{\goldmore}[1]{\multicolumn{4}{l}{\emph{\dots and #1 more reference values in the case's \file{expected}.}} \\}

% The theory hook: the generated files call \grouptheory{<group>} right after
% each group heading; groups with a hand-written theory block get one via the
% \deftheory mechanism below, the rest expand to nothing.
\makeatletter
\newcommand{\grouptheory}[1]{\@ifundefined{grouptheory@#1}{}{\csname grouptheory@#1\endcsname}}
\newcommand{\deftheory}[2]{\expandafter\def\csname grouptheory@#1\endcsname{%
  \begin{theorybox}#2\end{theorybox}}}
\makeatother
\newenvironment{theorybox}{%
  \par\medskip\noindent\begin{minipage}{\linewidth}\small
  \textbf{The theory you need.}\ }{\end{minipage}\par\medskip}
\newcommand{\theoryref}[1]{\emph{Full derivation: Theory Guide, #1.}}

% ===========================================================================
%  THE THEORY BLOCKS -- hand-written, condensed, with the load-bearing
%  equations; each ends with a pointer to the full Theory Guide chapter.
%  (The per-tutorial text below each block is GENERATED from the cases' own
%  headers -- regenerate with bin/curate/gen_tutorials_guide.py.)
% ===========================================================================

\deftheory{flash}{%
An isothermal flash splits a feed of composition $z_i$ into vapour $y_i$ and
liquid $x_i$ at fixed $(T,P)$.  Phase equilibrium is $y_i = K_i x_i$ with
$K_i = \gamma_i P_i^{\mathrm{sat}}/P$ ($\gamma$--$\varphi$ formulation), and
the vapour fraction $\beta$ solves the Rachford--Rice equation
\[ \sum_i \frac{z_i\,(K_i - 1)}{1 + \beta\,(K_i - 1)} = 0, \]
monotone in $\beta$ -- Newton converges in a handful of iterations, all of
them visible at verbosity 3.  Expect: a bubble point when the equation's
root hits $\beta = 0$, a dew point at $\beta = 1$, and non-ideal systems
($\gamma_i \ne 1$: NRTL/Wilson/UNIQUAC/UNIFAC) bending $K_i$ with
composition.  Liquid--liquid and three-phase cases replace
successive-substitution with direct Gibbs-energy minimisation (the symmetric
$\gamma$-model trap: Newton collapses onto the trivial $K=1$ root).
\theoryref{the VLE and flash chapters}}

\deftheory{distillation}{%
A column is $N$ flash stages coupled by counter-current flows.  Each stage
obeys MESH equations: Material balance, Equilibrium ($y_i = K_i x_i$),
Summation ($\sum y = \sum x = 1$), and Heat balance.  Choupo ships two
solution strategies, selectable per column: Wang--Henke (bubble-point,
solving the tridiagonal material balances stage by stage; robust for
near-ideal systems, slow near azeotropes) and \code{simultaneous} (one
Newton on ALL the MESH residuals; $\sim$5--6 iterations even through
azeotropes).  Expect: reflux and boilup set the split; a minimum reflux
below which no stage count reaches the spec; and azeotropes as fixed points
no column can cross ($x_D$ pinned at the azeotropic composition).
\theoryref{the distillation chapter (McCabe--Thiele + MESH)}}

\deftheory{reactor}{%
Three idealised steady reactors.  Conversion reactor: stoichiometry only,
$n_{\mathrm{out}} = n_{\mathrm{in}} + \nu\,\xi$, duty from the ONE enthalpy
base $\Delta H_{\mathrm{rxn}}(T) = \sum_i \nu_i h_i(T)$ (formation datum --
no separate ``heat of reaction'' input, and any dict override is
cross-checked aloud).  CSTR: outlet at reactor conditions with rate
$r(c_{\mathrm{out}},T)$, so $F\,(c_{\mathrm{in}} - c_{\mathrm{out}}) =
r\,V$.  PFR: the same rate integrated along the volume,
$\mathrm{d}F_i/\mathrm{d}V = \nu_i\,r$.  Arrhenius kinetics
$k = A\,T^b e^{-E_a/RT}$, orders from the dict, reverse rates from the
equilibrium constant (thermo-consistent by construction).  Expect: CSTR
conversion below PFR at equal volume (back-mixing), exponential $T$
sensitivity, and equilibrium limits the kinetics can approach but never
cross. \theoryref{the reaction-engineering chapter}}

\deftheory{reactors}{%
Batch reactors integrate $\mathrm{d}n_i/\mathrm{d}t = \nu_i\,r\,V$ from a
charged initial state -- no in/out flows; time replaces residence time.
Adiabatic runs couple the energy balance
$\sum_i n_i \bar{c}_{p,i}\,\mathrm{d}T/\mathrm{d}t = -\Delta H_{\mathrm{rxn}} r V$,
so composition and temperature evolve together.  Stiffness appears the
moment fast and slow reactions coexist (radical chains: $\mu$s radical
lifetimes vs ms induction) -- the explicit RK4 integrator then needs
$\Delta t$ at the FASTEST scale or blows up, while Rosenbrock23 (L-stable,
linearly implicit) steps at the scale of the SOLUTION.  The stiffness ratio
is printed; watch it. \theoryref{the batch/ODE chapter}}

\deftheory{gibbs}{%
Equilibrium without a mechanism: minimise total Gibbs energy
$G = \sum_i n_i\,\mu_i$ over ALL species subject to elemental balances
$\mathbf{A}\,\mathbf{n} = \mathbf{b}$ (Lagrange multipliers per element --
printed as $\lambda_E$, the element potentials).  The species list in the
dict IS the model: what you include can form; what you omit cannot.  With
enough species this reproduces flame temperatures and radical pools with no
kinetics at all: expect the adiabatic flame temperature to land within a few
K of literature when radicals (H, O, OH\dots) are IN the pool, and to
overshoot by $\sim$100--250\,K when they are not (dissociation is
endothermic).  Above $\sim$1500\,K the catalogue's polynomial $c_p$ fits
extrapolate WRONG -- combustion cases carry case-local NASA-7 overrides,
and the case headers say so.

\medskip
\emph{Equilibrium maps.}  Sweeping the Gibbs solve over a $T\times P$ grid
gives an equilibrium map -- iso-lines of a product fraction over the
operating plane -- which shows why an industrial reactor fixes $T$ and $P$
in a narrow window (equilibrium wants one corner, the catalyst works in
another; the gap is the compromise).  The \code{temperatureApproach}
$\Delta T$ evaluates the \emph{reaction} equilibrium at $T+\Delta T$ while
enthalpy and the energy balance stay at the physical $T$ -- an empirical,
calibrated closeness-to-equilibrium knob (never predicted).
\emph{Gradeable exercise (ammonia):} on \code{nh3\_equilibrium\_map},
at $P = 200$\,bar and $\Delta T = 0$, find the temperature at which the
equilibrium NH$_3$ mole fraction crosses 30\,\%; report $T$ to the grid
resolution and verify against the case's anchor KPIs.  Unique because the
fraction is monotone in $T$ at fixed $P$. \theoryref{the Gibbs-equilibrium
chapter, \S equilibrium maps}}

\deftheory{flowsheets}{%
Recycle turns a flowsheet into a fixed-point problem: guess the tear stream,
march the units, compare.  Successive substitution converges linearly (and
diverges past unit gain); Wegstein accelerates it; Newton-on-tears solves
the tear residuals with a Jacobian.  Choupo announces the tear choice, every
iteration's residual, and any bound that binds -- no silent crutch.  Expect:
purge streams controlling inert build-up (close the loop with no purge and
watch the accumulation), and convergence behaviour that CHANGES with the
tear location. \theoryref{the flowsheet-convergence chapter}}

\deftheory{evaporation}{%
An evaporator boils solvent off a solution.  Two pieces of physics beyond a
flash: boiling-point elevation, $T_{\mathrm{boil}} =
T_{\mathrm{sat}}(P) + \mathrm{BPE}$ with the BPE from the electrolyte
activity of water ($\ln a_w = -\varphi\,\nu\,m\,M_w/1000$); and the heat of
dilution/concentration (the relative apparent molar enthalpy
$L_\varphi(m,T)$) entering the duty.  Expect: the BPE to GROW as the brine
concentrates, steam economy below 1 for a single effect, and -- for NaCl --
the full Silvester--Pitzer $T$-dependence of $\gamma_\pm$ shaping the answer
between 25 and 200$^\circ$C. \theoryref{the electrolyte chapters}}

\deftheory{crystallisation}{%
Crystallisation removes solute when the solution crosses saturation:
$m > m_{\mathrm{sat}}(T)$, with $K_{\mathrm{sp}}(T)$ from
$\Delta G_{\mathrm{diss}}$ and its $T$-dependence from
$\Delta H_{\mathrm{diss}}$ (van 't Hoff).  The heat of crystallisation is
ION-DERIVED -- $\Delta H_{\mathrm{cryst}} = -(\Delta H_{\mathrm{soln}}^\infty
+ \bar{L}_2(m_{\mathrm{sat}}))$ -- never a stored constant (a solid's
formation enthalpy is $\sum \nu_i h_{f,i}^{\mathrm{aq}} -
\Delta H_{\mathrm{soln}}$, derived, single-sourced).  Antisolvent cases add
the mixed-solvent electrolyte model (eNRTL): ethanol drops the dielectric
constant, $\gamma_\pm$ rises, salt crashes out.  Expect supersaturation as
the driving force and yields set by the solubility curve, not the kinetics
(these cases are equilibrium-staged). \theoryref{the crystallisation +
electrolyte-enthalpy chapters}}

\deftheory{membranes}{%
Solution-diffusion (Wijmans \& Baker 1995): inside a dense membrane the
pressure is uniform, so permeation is concentration-driven.  Linearised:
$J_w = A_w(\Delta P - \Delta\pi)$ and $J_s = B_s(c_m - c_p)$ -- water driven
by NET pressure, solute blind to pressure; $(A_w, B_s)$ are MATERIAL
constants from \file{data/standards/assets/}.  Concentration polarisation
(film model) makes the wall concentration $c_m = c_b\,e^{J_w/k}$ exceed the
bulk -- the flux--polarisation--osmotic feedback is THE nonlinearity of
RO/NF, solved per module segment.  Expect: rejection improving with
pressure ($c_p = B_s c_m/(J_w + B_s)$), polarisation eating driving force at
high flux, and charged-membrane (Donnan) effects on multivalent ions that
plain solution-diffusion cannot see -- that is what the DSPM-DE sibling
model is for. \theoryref{the membrane chapter}}

\deftheory{electrolyte}{%
Ions interact at long range (Debye--H\"uckel, $\sqrt{I}$ law) and short
range (ion--ion, ion--solvent).  Choupo's ladder: Davies (charge-only,
$I \le 0.5$), single-salt Pitzer ($\beta_0,\beta_1,C^\varphi$; to
$I \sim 6$, and with the Silvester--Pitzer $T$-functions to 300$^\circ$C),
multi-ion Pitzer--HMW (per-ion, mixing terms $\theta,\psi$ -- seawater-class
brines), and eNRTL (local-composition; single-salt, mixed-solvent, and the
2009 symmetric multi-salt with both published and Gibbs--Duhem-consistent
formulations).  Speciation solves mass action + charge balance by Newton in
log-molality space; saturation indices $\mathrm{SI} =
\log(\mathrm{IAP}/K_{\mathrm{sp}})$ flag scaling.  Expect: models to
DISAGREE above their validity ($\gamma$ divergence is the lesson, e.g.\
Davies vs Pitzer at RO recoveries), and every activity model to announce
itself. \theoryref{the electrolyte + speciation chapters}}

\deftheory{combustion}{%
Radical chain chemistry: initiation creates the first radicals (slow --
$\mathrm{H_2 + M}$, $E_a \sim 105$\,kcal), branching multiplies them
($\mathrm{H + O_2 \to O + OH}$: one in, two out), termination removes them.
Ignition is the race between branching and termination; the induction time
is set by the SLOW build-up, then heat release runs away.  Rate constants
$k = A T^b e^{-E_a/RT}$ come VERBATIM from cited mechanisms (GRI-Mech 3.0,
\'O Conaire 2004, Glarborg 2018) in their native units -- the engine
converts and announces.  Reverse rates from equilibrium: the kinetic
end-state must land on the Gibbs equilibrium (the two-engine handshake the
NOx tutorial pins at 98.9\,\%).  Expect stiffness ratios of $10^6$--$10^8$
(use Rosenbrock23), ignition delays in $\mu$s validated against shock-tube
data, and thermal NO forming over SECONDS -- the timescale gap behind
quench-based NOx control. \theoryref{the combustion/kinetics chapter}}

\deftheory{steam}{%
Water/steam properties from IAPWS-IF97: regions (liquid, vapour,
supercritical), $T_{\mathrm{sat}}(P)$, and $h(T,P)$/$s(T,P)$ from the
dimensionless Gibbs functions.  Power cycles are enthalpy bookkeeping around
them: pump work $\approx v\,\Delta P$, turbine work from isentropic
expansion $\times$ efficiency, cycle efficiency $\eta = W_{\mathrm{net}}/Q_{\mathrm{in}}$
bounded by Carnot. \theoryref{the steam/power chapter}}

\deftheory{power}{%
Gas (Brayton) and steam (Rankine) cycles, alone and combined.  Compressor
and turbine hardware is specified by SHAFT WORK and isentropic efficiency --
outlet pressure is a RESULT, not an input (the anti-DesignSpec credo).
Expect: pressure ratio setting Brayton efficiency, the combined cycle's gain
coming from bottoming the gas turbine's exhaust exergy, and every energy
flow visible in the ledger. \theoryref{the power-cycle chapter}}

\deftheory{heat}{%
Heat exchange couples streams without mixing them: $Q = U\,A\,\Delta
T_{\mathrm{lm}}$, effectiveness-NTU when outlets are unknown.  Utilities
(steam levels, cooling water) are allocated by temperature level -- a
reboiler cannot be heated by a colder utility, and the allocator says which
level it picked.  Heat-links (forward and feedback) couple distant units
into an energy tear the solver converges like any recycle.  Pinch analysis
finds the minimum utilities the composite curves allow.

\medskip
\emph{Exchanger DESIGN vs rating.}  A shell-and-tube poses two inverse
questions: RATE it (geometry given $\to$ find $U$, area, $\Delta P$, duty --
\code{model geometry;}) or DESIGN it (duty given $\to$ size the geometry --
\code{model design;}).  $U$ is BUILT from the two film coefficients
(Gnielinski tube, Kern shell) and the wall; the pressure drop (Kern both
sides) is the duty-vs-pumping half a real design lives by.  The standard
workflow -- establish the duty, THEN design, THEN rate, never invent the
geometry up front -- is walked end-to-end in the two-part sequence
\code{hxWorkflow1\_design\_from\_duty} $\to$ \code{hxWorkflow2\_rate\_designed},
and the design ends in a printable datasheet (GUI: the unit panel).
\theoryref{the heat-integration chapter, \S heat-exchanger design}}

\deftheory{estimate}{%
When a compound is missing, group contribution builds it: Joback sums
first-order group increments into $T_b, T_c, P_c, V_c, \Delta H_f^\circ,
\Delta H_{\mathrm{vap}}, c_p^{\mathrm{ig}}(T)$; Lee--Kesler closes $\omega$
from $(T_b, T_c, P_c)$; Ambrose--Walton gives $P^{\mathrm{sat}}(T)$ from
corresponding states; Rackett the liquid volume.  The estimate is a
PROPOSAL: a \file{.dat} you review against any reference data and promote
by hand -- the engine never estimates at runtime.  Molecular structure is
IDENTITY: components carry \code{jobackGroups\{\}} in their own record, and
\code{bin/estimate <name>} is the one-command curation act.  Expect
few-percent errors on criticals (and the guide's worked example shows
them), weaker $T_b$ for hydrogen-bonded species, and polymers via van
Krevelen / Yang on the repeat unit. \theoryref{the estimation chapter of
the Props Guide}}

\deftheory{scan}{%
Property scans sweep a model across $T$, $P$ or composition and plot what
the thermodynamics DOES: $P^{\mathrm{sat}}(T)$ curves, $\gamma_\pm(m)$,
VLE envelopes, ternary diagrams, phase boundaries.  Nothing is fitted --
these ops answer ``how does the selected model behave?'', the SEE step of
the props loop (CREATE $\to$ SEE $\to$ TEST $\to$ FIT $\to$ PROMOTE).
Expect model disagreements to be VISIBLE (overlay two methods and look), and
every curve exportable as CSV next to the case. \theoryref{the Props Guide}}

\deftheory{thermo}{%
Consistency tests interrogate DATA, not models: the van Ness point test
checks measured $(x,y,T,P)$ sets against the Gibbs--Duhem constraint
(residuals in $\ln(\gamma_1/\gamma_2)$); Herington's area test is its
integral cousin.  A dataset that fails is not ``bad'' -- the test says WHERE
and HOW MUCH, and the fit that follows should weight accordingly.  Expect
plain-language verdicts with severity, not a pass/fail stamp.
\theoryref{the consistency chapter of the Props Guide}}

\deftheory{compare}{%
Method comparison runs the SAME property through different models
side-by-side -- UNIFAC vs NRTL, Joback vs measured, ideal vs EOS.  The gap
between curves IS the modelling uncertainty; picking a method is a decision
the case author owns, made with eyes open. \theoryref{the Props Guide}}

\deftheory{still}{%
A batch still is Rayleigh distillation: $\mathrm{d}(n\,x_i)/\mathrm{d}t =
-\dot{V} y_i$ with $y_i$ in equilibrium with the pot -- the pot composition
DRIFTS as the light component leaves (contrast the steady column, where
stages hold profiles constant).  Expect the distillate to start rich and
fade, and the recipe layer (charge, heat, switch receivers) to drive the
run. \theoryref{the batch-distillation section}}

\deftheory{recipes}{%
A recipe is time-triggered discrete control over batch units: charge,
heat/cool, hold, transfer, sample.  Between events the DAE integrator runs;
at each event the state jumps (a transfer moves holdup between vessels).
Expect exact event times on the trajectory and holdup bookkeeping that
conserves mass across transfers.}

\deftheory{adsorber}{%
Adsorption in a vessel or a bed: the isotherm $q^{*}(T,p)$ (Langmuir
$q = q_{\max} b p/(1+bp)$, van't Hoff $b(T)$) is the curated PAIR datum;
the LDF rate $\mathrm{d}q/\mathrm{d}t = k\,(q^{*} - q)$ carries the
EQUIPMENT datum $k$ (this particle, this bed -- it lives in the case, never
the catalogue).  Work the cases as a staircase.  \code{batch09}: a closed
vessel relaxes to the inventory equilibrium -- one algebraic root, pinned
independently of the run.  \code{batch10}: two species compete for the
same sites (extended Langmuir) -- the selectivity is an equilibrium
statement, and the model's Gibbs--Duhem inconsistency for unequal
$q_{\max}$ is declared, not hidden.  \code{batch11}: in the Henry limit
the whole vessel is LINEAR and the integrator must match
$q_\infty(1-e^{-k_{\mathrm{eff}}t})$ to $10^{-8}$ -- the analytic gate.
\code{batch12}: declaring the headspace pressure AND the inventory
over-specifies a closed ideal-gas vessel -- the contradiction FATALs;
$P$ is a result.  \code{batch13}: the fixed bed -- the stoichiometric
time $t_{\mathrm{st}} = (L/u)(\varepsilon + \rho_b q^{*}/c_{\mathrm{in}})$
is announced BEFORE integration, breakthrough spreads around it, and the
constant-$u$ closure's fabricated carrier is printed, pinned and
discounted: the DECLARED model residual is itself the lesson.
\code{batch14}/\code{batch15} are the numerical hygiene: pure transport
vs the analytic advection--dispersion solution, and the mesh study that
earns $N = 100$. \theoryref{the adsorption chapter}}

\deftheory{adsorption}{%
The isotherm bench: evaluating a curated adsorption-equilibrium record
(Henry-limit, saturation, pin and unit-invariance gates run aloud) and
REGRESSING one from data -- the multi-temperature van't Hoff Langmuir fit
with identifiability verdicts: single-$T$ data refuses $\Delta
H_{\mathrm{ads}}$, Henry-regime data yields only the product $q_{\max}
b$, and a non-converged fit refuses its verdict by name.
\theoryref{the adsorption chapter}}

\deftheory{absorption}{%
Gas absorption transfers a solute from gas to liquid across stages; Henry's
law ($p_i = H_i x_i$) anchors the equilibrium at dilution.  Expect the
L/G ratio to control the operating line's slope, and a minimum solvent rate
at the pinch with the equilibrium line.
\theoryref{the absorption section}}

\deftheory{solids}{%
Solids carry mass and enthalpy without VLE: crystallisers, dryers, cyclones,
gas--solid splitters.  A dryer's air is a REAL stream (two inlets, adiabatic
mixing, outlet humidity from the psychrometrics); a cyclone's cut size comes
from the selected model (Lapple\dots Muschelknautz).  Expect solid streams
flagged separately ($F_{\mathrm{solid}}$) and energy balances that close
across phase change.}

\deftheory{hydraulics}{%
Pressure drop and flow distribution: Darcy--Weisbach with friction factors,
fittings as equivalent lengths, pump curves against system curves.  Expect
the operating point at the curves' intersection.}

\deftheory{utilities}{%
Utility streams (steam levels, cooling water, refrigerants, electricity)
carry duty into and out of the plant at declared temperature levels and
prices; allocation is by $T$-level feasibility, announced.}

\deftheory{rotating}{%
Compressors, turbines, pumps: specified by shaft work $W_s$ and isentropic
efficiency $\eta_s$; the outlet state follows from
$h_{\mathrm{out}} = h_{\mathrm{in}} + W_s/\dot{n}$ with the isentropic
reference path.  Outlet pressure is a RESULT (hardware, not a spec).}

\deftheory{optimisation}{%
The outer driver wraps the whole simulation in $f(\mathbf{x})$: sweeps map
it, Nelder--Mead descends it, parameter estimation fits it to data
(Levenberg--Marquardt, with identifiability diagnostics --
$\chi^2_{\mathrm{red}}$, correlation, conditioning).  Expect every function
evaluation to be a full, honest flowsheet pass.}

\deftheory{molecular}{%
Molecular activity models on display: NRTL/Wilson/UNIQUAC correlate binary
$\gamma$'s from pair parameters; UNIFAC PREDICTS them from group
counts.  Expect UNIFAC within a few percent of fitted models for families it
was trained on -- and to be honestly worse outside them.
\theoryref{the activity-model chapter}}

\deftheory{esterification2sector}{%
The self-contained two-sector demo: reaction (esterification CSTR) +
separation, fractal case layout (each sector its own folder), recycle
between them.  The showcase for the dignified-unit layout.}

\deftheory{twoSectorDemo}{%
The zippable two-sector demonstration plant -- the case used to rehearse
the online landing.  Same physics as the esterification demo.}

\deftheory{ChemicalPlantTutorial}{%
The full multi-sector plant showcase: feed treatment, reaction, separation,
utilities -- the fractal flowsheet at its largest, run-only (no golden), for
reading and dissecting rather than reproducing numbers.}

\deftheory{polycaprolactonePlant}{%
Ring-opening polymerisation plant: monomer to polycaprolactone with
step-growth KPIs (Carothers, Flory) on the PFR, polymer properties from
group contribution on the repeat unit.}

\deftheory{userops}{%
User-defined operations: the extension seam.  A custom property op or unit
compiled into the binary via the explicit factory -- the tutorial shows the
registration pattern (\code{registerBuiltins}) end to end.}

\deftheory{economics}{%
Sizing (stirred tanks, shell-and-tube areas) feeds costing (Guthrie
modules); prices are CASE data (\file{constant/economics}, dated and
cited) -- the engine refuses to invent a price.  Expect honest refusal when
a price is missing, never a silent default.}

% ===========================================================================
\title{}
\author{}
\date{}

\begin{document}

\manualfrontmatter{Tutorials Guide}{
Choupo is a trademark of TalentGround Lda.

Anthropic, Claude, and Claude Code are trademarks or registered trademarks of
Anthropic, PBC.
}

{\small\setcounter{tocdepth}{3}\setcounter{secnumdepth}{0}\tableofcontents}
\clearpage

% =====================================================================
\section*{How to use this guide}
\addcontentsline{toc}{section}{How to use this guide}

Every Choupo tutorial is a complete, runnable case:
\begin{verbatim}
   source etc/bashrc
   runCase tutorials/<category>/<group>/<name>
\end{verbatim}
This guide gives each tutorial the CONTEXT the case file cannot: the theory
block at the head of each group states the physics and the equations you
need (condensed -- the Theory Guide carries the full derivations), and each
entry tells you \textbf{what the case does} and \textbf{what to expect}
before you run it.  The per-tutorial text is assembled from the cases' own
headers -- the case file is always the source of truth, and reading it IS
part of the exercise.

Three reading rules:
\begin{enumerate}[leftmargin=*,nosep]
  \item \textbf{Run first, read second, re-run third.}  The log at
        verbosity 3 shows every Newton iteration, every K-value, every
        announcement -- the glass box is the point.
  \item \textbf{Expect the announcements.}  Choupo never silently helps:
        initial guesses, bounds that bind, model fallbacks, unit overrides
        -- all are printed.  If a run is quiet about something, it did not
        happen.
  \item \textbf{Golden numbers are commitments.}  Cases carrying an
        \file{expected} file pin their KPIs; the validation AADs quoted in
        the entries are locked by the regression suite and re-verified on
        every change.
\end{enumerate}

% =====================================================================
\section*{Learning paths}
\addcontentsline{toc}{section}{Learning paths}

\paragraph{The MCFT path (computational transfer phenomena).}
Start with \texttt{flash01\_benzene\_toluene} (the Rachford--Rice loop,
visible), then \texttt{flash} non-ideal cases (activity models bend the
K-values), \texttt{distillation} (stages couple; Wang--Henke vs
simultaneous), one \texttt{flowsheets} recycle case (tears and convergence),
and close with a \texttt{gibbs} flame (equilibrium without mechanism).
Then the batch side: an \texttt{ignition} case for stiffness -- run the
\texttt{\_rk4} twin and watch the explicit method die -- and a \texttt{ctrl}
loop for feedback.

\paragraph{The electrolytes path.}
\texttt{pitzer01\_nacl} ($\gamma_\pm$, $\varphi$, $a_w$ of one salt), then
\texttt{pitzer\_gamma\_hamer\_wu} (the catalogue vs NBS data, zero refit),
\texttt{pitzer\_nacl\_sp77\_hot} (temperature, to 200$^\circ$C),
speciation (\texttt{scaling} cases: SI, pH, Davies-vs-Pitzer forking an
engineering answer), the eNRTL series
(\texttt{enrtl\_mixed\_nacl\_ethanol\_esteso} mixed-solvent;
\texttt{enrtl\_multisalt\_nacl\_kcl} two salts, two formulations, an
experimental verdict), and the process payoff: evaporators and
crystallisers where all of it sets duties and yields.

\paragraph{The combustion path.}
\texttt{gibbs06\_h2\_flame\_radicals} (the radical pool at equilibrium; the
$c_p$ extrapolation trap), \texttt{ignition01\_h2o2} (GRI kinetics, the
stiffness lesson via the \texttt{\_rk4} twin), \texttt{ignition02\_h2air\_slack}
(quantitative: $\tau_{\mathrm{ign}}$ vs shock-tube data),
\texttt{nox01\_zeldovich\_postflame} (thermal NO; kinetics meets Gibbs at
98.9\,\%).

\paragraph{The props-bench path.}
\texttt{estimate\_acetone} (CREATE: groups $\to$ constants, error shown),
a \texttt{scan} case (SEE), \texttt{vle01\_consistency} (TEST the data),
a \texttt{fit} case (Levenberg--Marquardt with identifiability), then the
promote-and-use loop into a steady case.  The Props Guide is this path's
companion.

% =====================================================================
\clearpage
\section*{Category I: Steady-state flowsheets (choupoSolve)}\addcontentsline{toc}{section}{Category I: Steady-state flowsheets (choupoSolve)}\markboth{Steady-state flowsheets}{Steady-state flowsheets}
Solved as one nonlinear system $F(\mathbf{x}) = \mathbf{0}$: units in
sequence, recycles as tears, everything announced.
\input{tutorialsGuide-steady}

\clearpage
\section*{Category II: The property bench (choupoProps)}\addcontentsline{toc}{section}{Category II: The property bench (choupoProps)}\markboth{The property bench}{The property bench}
Questions asked of the thermodynamics BEFORE a simulation uses it --
CREATE, SEE, TEST, FIT, PROMOTE.
\input{tutorialsGuide-props}

\clearpage
\section*{Category III: Batch and time-dependent (choupoBatch)}\addcontentsline{toc}{section}{Category III: Batch and time-dependent (choupoBatch)}\markboth{Batch and time-dependent}{Batch and time-dependent}
$\mathrm{d}\mathbf{Y}/\mathrm{d}t = f(\mathbf{Y})$ integrated in real time,
with recipes for discrete events; trajectories, not steady states.
\input{tutorialsGuide-batch}

\clearpage
\section*{Category IV: Dynamics and control (choupoCtrl)}\addcontentsline{toc}{section}{Category IV: Dynamics and control (choupoCtrl)}\markboth{Dynamics and control}{Dynamics and control}
Continuous processes in time with feedback: PID loops, setpoint schedules,
disturbances and responses.
\input{tutorialsGuide-ctrl}

\clearpage
\section*{Category V: Electrochemical systems}\addcontentsline{toc}{section}{Category V: Electrochemical systems}\markboth{Electrochemical systems}{Electrochemical systems}
\input{tutorialsGuide-electrochem}

\clearpage
\section*{Category VI: Multi-sector plant showcases}\addcontentsline{toc}{section}{Category VI: Multi-sector plant showcases}\markboth{Plant showcases}{Plant showcases}
The fractal flowsheet at full size -- read them like worked examples.
\input{tutorialsGuide-plant}

\end{document}
