\documentclass[11pt,a4paper]{article}
\newcommand{\docTitle}{User Guide}
\input{preamble.tex}

\title{}
\author{}
\date{}

\begin{document}

\manualfrontmatter{User Guide}{
Choupo is a trademark of TalentGround Lda.

Anthropic, Claude, and Claude Code are trademarks or registered trademarks of
Anthropic, PBC.

ParaView is a trademark of Kitware Inc.
}

% =====================================================================
%   Table of contents
% =====================================================================
\pagenumbering{roman}
\setcounter{tocdepth}{2}
\markboth{Contents}{Contents}
\tableofcontents
\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}

% =====================================================================
%   Preface
% =====================================================================
\section*{Preface --- read this first}
\addcontentsline{toc}{section}{Preface --- read this first}

Engineers used to text-driven, file-first tools will feel at home.
Newcomers will feel curious, then liberated.  Choupo is an
\emph{educational} process simulator: every Newton iteration, every
$K$-value, every Wegstein step is visible both in the source and at
runtime.  If something surprises you, the answer is always one
\texttt{grep} away.

This guide walks you from a cold install to running, reading, modifying,
and building your own cases: first the case anatomy (the plain-text
dictionaries and the per-stream state files), then the thermodynamic
foundation, then every family of unit operations with a minimal working
dict, then the outer layers --- sweeps, design specifications,
optimisation, sizing and costing --- and finally the time-dependent
binaries, the property workbench, and the browser front-end.  Each
section opens with a \emph{What you'll learn} note; boxes mark
crystal-clear takeaways, common pitfalls, and small experiments you
should run as you read.

Four binaries cover four problem classes, and one dispatcher
(\code{runCase}) picks the right one for each case:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}llX@{}}
\toprule
\textbf{Binary} & \textbf{Problem class} & \textbf{Mathematics} \\
\midrule
\code{choupoSolve} & steady-state flowsheets & $F(x)=0$ root-finding; Newton-on-tears recycles \\
\code{choupoBatch} & batch vessels + recipes & $\mathrm{d}Y/\mathrm{d}t = f(Y)$, RK4 or stiff Rosenbrock \\
\code{choupoCtrl}  & dynamic continuous + control & $\mathrm{d}Y/\mathrm{d}t = f(Y,u,t)$ with controllers writing $u$ \\
\code{choupoProps} & property evaluation & points, scans, consistency tests, LM fits \\
\bottomrule
\end{tabularx}
\end{center}

% =====================================================================
\section{Quick start}

\begin{whatyoulllearn}
How to build the simulator once, set up your shell environment, and
run your first tutorial in under five minutes.
\end{whatyoulllearn}

Choupo needs only a C\texttt{++}17 compiler (\code{g++}) and GNU
\code{make} --- no external libraries.  Build once, source the shell
environment once, then run any tutorial:

\begin{lstlisting}
cd /path/to/Choupo
make all                                       # release build
source etc/bashrc                              # adds bin/ + project root to PATH

listCases                                      # available tutorials
runCase tutorials/steady/flash/flash01_benzene_toluene     # run one
less tutorials/steady/flash/flash01_benzene_toluene/log.choupoSolve
\end{lstlisting}

Tutorials live under \file{tutorials/<category>/<family>/<name>/}, where
the category is one of \file{steady/} (\code{choupoSolve}),
\file{batch/} (\code{choupoBatch}), \file{ctrl/} (\code{choupoCtrl}),
\file{props/} (\code{choupoProps}), \file{electrochem/}, or
\file{plant/} (multi-sector cases).  \code{runCase} reads each case's
\code{controlDict.application} and dispatches to the right binary
automatically --- you never pick the executable by hand.  After your
first build a symlink per binary (e.g.\ \file{./choupoSolve}) points to
the matching \file{build/<platform>/} executable.

\subsection{Permanent shell setup}

Append this line to \file{\textasciitilde/.bashrc} so every new terminal
has the environment ready:

\begin{lstlisting}
source /path/to/Choupo/etc/bashrc
\end{lstlisting}

A fresh terminal then greets you with:

\begin{lstlisting}
Choupo environment loaded.
  CHOUPO_HOME      = /path/to/Choupo
  CHOUPO_WORKSPACE = /path/to/Choupo/cases   (your cases; newCase writes here)
  Commands available  : choupoSolve, runCase, runApplication, newCase,
                        listCases, cleanCase, Choupo, cdTut, cdWork
\end{lstlisting}

\subsection{Build modes}

\begin{lstlisting}
make all                # default --- release, -O2
make MODE=debug all     # -O0 -g, parallel tree under build/<platform>Debug
make print              # show resolved build variables
make clean              # remove current mode's objects
make distclean          # remove the whole build/ tree (and the WASM build)
\end{lstlisting}

Both modes coexist; the \file{choupoSolve} symlink points to the
\emph{last} binary built.

\begin{keyidea}
You never recompile to switch cases.  Build once, source once, then
all you do is edit text dictionaries and re-run.
\end{keyidea}

\begin{tryit}
Run \code{runCase tutorials/steady/flash/flash01\_benzene\_toluene} and then
\code{less} the \file{log.choupoSolve} it writes.  Notice the
per-iteration trace under ``Rachford-Rice Newton'': that is the
simulator showing its work.
\end{tryit}

% =====================================================================
\section{The shell helpers}

\begin{whatyoulllearn}
A handful of short commands cover every interaction you'll have with
the simulator from the shell.
\end{whatyoulllearn}

After sourcing \file{etc/bashrc} you have:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Command} & \textbf{Purpose} \\
\midrule
\code{choupoSolve <case>}    & Run a case directly; write everything to stdout. \\
\code{runCase [<case>]}      & Run + redirect to \file{log.<binary>} in the case dir; dispatches by \code{controlDict.application}.  Defaults to the current dir. \\
\code{runApplication <binary> <args>} & Run any binary, redirect to \file{log.<binary>} in the cwd. \\
\code{listCases}             & Tabulate all tutorials with one-line descriptions. \\
\code{newCase <name> [note]} & Scaffold a fresh case in \code{\$CHOUPO\_WORKSPACE}. \\
\code{cleanCase [<case>]}    & Remove \file{log.*} and generated outputs (never touches source). \\
\code{Choupo}                & \code{cd} to the project root. \\
\code{cdTut}                 & \code{cd} to \file{tutorials/}. \\
\code{cdWork}                & \code{cd} to your case workspace. \\
\bottomrule
\end{tabularx}
\end{center}

Two equivalent workflows:

\begin{lstlisting}
# From anywhere (use $CHOUPO_HOME):
runCase $CHOUPO_HOME/tutorials/steady/flash/flash01_benzene_toluene

# From inside the case dir:
cd tutorials/steady/flash/flash01_benzene_toluene
runCase                                 # writes log.choupoSolve here
\end{lstlisting}

Two further tools appear later in this guide: \code{bin/choupo-init0}
(materialise the initial stream state, \S\ref{sec:init0}) and
\code{bin/choupo-import} (seal a case's property data so it runs
self-contained, \S\ref{sec:self-contained}).  \code{bin/runTests} runs
the full regression over every bundled tutorial.

% =====================================================================
\section{Anatomy of a case directory}\label{sec:anatomy}

\begin{whatyoulllearn}
Every case is a folder of plain-text files with a fixed division of
labour: \file{system/} says how the problem is organised,
\file{constant/} carries the models and data, and --- in a steady case
--- the \file{0/} directory carries the complete initial state of every
stream, one file per stream, OpenFOAM-style.  The solver writes its
answer to \file{converged/}.
\end{whatyoulllearn}

The tree below is a \emph{steady} (\code{choupoSolve}) case --- the most
common kind, and the one whose anatomy the rest of this section
dissects.  The other binaries reuse the same \file{system/} +
\file{constant/} skeleton but not the stream-state directories: a batch
or control case has no flowsheet streams to seed --- its vessels carry
\code{initial \{\}} charges and the run writes \file{trajectory.csv}
(\S\ref{sec:batch}, \S\ref{sec:ctrl}) --- and a props case replaces
\file{flowsheetDict} with a \file{propsDict} altogether
(\S\ref{sec:props}).

\begin{lstlisting}
caseName/
|-- caseName.cho             empty marker file --- the GUI's "openable" entity
|-- system/
|   |-- controlDict          REQUIRED  application, verbosity, reports{}
|   |-- flowsheetDict        REQUIRED  TOPOLOGY ONLY: units + connections
|   |-- solverDict           optional  per-unit-op solver options
|   |-- outerDict            optional  sweep / designSpec / optimisation
|   `-- postDict             optional  sizing + costing chain
|-- constant/
|   |-- propertyDict         REQUIRED  components + thermodynamic models
|   |-- reactions            optional  named-reaction library
|   |-- crystallisation      optional  crystallisation-kinetics library
|   |-- dryingKinetics       optional  drying-curve library
|   |-- components/          optional  case-local component overlays
|   `-- propertyData/        optional  sealed data snapshot (self-contained cases)
|-- 0/                       REQUIRED (steady only)  initial stream state,
|   |-- feed   liquid   vapor                        ONE FILE PER STREAM
`-- (written by the run)
    |-- converged/           the steady solution, same one-file-per-stream shape
    |-- reports/             CSV + .ods reports selected in controlDict
    `-- log.choupoSolve      written by runCase
\end{lstlisting}

\begin{keyidea}
A standalone unit operation is just a \file{flowsheetDict} of length 1.
There is no special ``single-unit'' mode --- one consistent case format
for everything, from one flash drum to a multi-sector plant.
\end{keyidea}

\subsection{\file{controlDict}}

The essential keys:

\begin{lstlisting}[language=dict]
application   choupoSolve;      // | choupoBatch | choupoCtrl | choupoProps
description   "Isothermal flash benzene/toluene at 370 K, 1 bar (Raoult)";
verbosity     3;                // 0 silent / 1 +warnings / 2 +summary
                                // 3 info (Newton iterations visible) / 4 debug
reports
{
    streamTable    { }          // see the Reports section
    massBalance    { }
    energyBalance  { }
}
\end{lstlisting}

For pedagogical use you want \code{verbosity 3} --- that prints every
Newton iteration, every Wegstein step, every $K$-value refresh.  Set it
lower only when an outer driver sweeps the simulator hundreds of times.
Time-dependent cases (\code{choupoBatch} / \code{choupoCtrl}) add time
controls --- see \S\ref{sec:batch}.

\subsection{\file{flowsheetDict} --- topology only}\label{sec:flowsheet}

The \file{flowsheetDict} declares \emph{what is connected to what} ---
units, their wiring, and (for recycles) the tear declaration.  It
carries \textbf{no stream numbers}: state lives in \file{0/}
(\S\ref{sec:zero}).  The first tutorial in full:

\begin{lstlisting}[language=dict]
// stream state lives on disk in 0/ (per-stream files);
// flowsheetDict is TOPOLOGY only

units
(
    {
        name        flash01;
        type        isothermalFlash;
        in          feed;
        outputs     ( liquid  vapor );

        operation   { T 370.0 K;  P 1.0 bar; }
    }
);
\end{lstlisting}

Every unit entry follows the same slot order: \code{name} and
\code{type} (required), an optional \code{model} (the sub-model
selector, e.g.\ \code{simultaneous} on a column), then \code{in} (one
inlet) or \code{inputs (...)} (several), \code{outputs (...)}, the
\code{operation \{...\}} block of numeric knobs, and finally optional
\code{reaction} / \code{crystallisation} / \code{dryingCurve}
references and an optional per-unit \code{thermo \{\}} override
(\S\ref{sec:thermo-override}).

\paragraph{Identifying streams: inputs by name, outputs by \emph{position}.}%
\label{sec:flowsheetDict-streams}
This is the one rule that trips up every new author, so learn it once.
A unit's \code{inputs} and its \code{outputs} are matched to physical
streams by \emph{different} mechanisms.  An \textbf{input} is wired
\emph{by name}: you write \code{in feed;} (or \code{inputs ( hot cold );})
and the unit reads each stream by the name you gave it (a
\code{heatExchanger} names its tube-side inlet with \code{tubeStream}; a
multi-feed column maps each named feed to a stage in
\code{operation.feeds}).  An \textbf{output}, by contrast, is matched
\emph{by position}.  Every unit produces its streams in a \emph{fixed
internal order}, and your \code{outputs} list does nothing more than
\emph{rename that ordered list}: \code{outputs[$k$]} becomes the name of
the $k$-th stream the unit emits.

\begin{center}
\begin{tikzpicture}[>=Stealth, font=\footnotesize, line width=0.4pt]
  \node[draw, fill=gray!5, minimum width=2.4cm, minimum height=1.6cm]
        (u) at (0,0) {\code{flash01}};
  \draw[<-] (u.west) -- ++(-1.4,0) node[left]{\code{feed}};
  % produced streams, in emission order
  \draw[->] (u.east)++(0,0.45) -- ++(1.4,0)
        node[right]{\code{producedStreams()[0]} $\to$ \code{liquid}};
  \draw[->] (u.east)++(0,-0.45) -- ++(1.4,0)
        node[right]{\code{producedStreams()[1]} $\to$ \code{vapor}};
  \node[align=center, font=\scriptsize\itshape, text=gray!60!black]
        at (0,-1.55) {\code{outputs ( liquid vapor )} renames by index --- the ORDER matters};
\end{tikzpicture}
\end{center}

\noindent The engine does literally \code{s = produced[k]; s.name = outputs[k]}.
So the order of your \code{outputs} list must follow the unit's emission
order (\code{isothermalFlash}: liquid, vapour; \code{crystalliser}:
crystals, mother liquor; \code{cyclone}: clean gas, captured solids;
\code{distillationColumn}: distillate, bottoms, then side draws by
ascending stage --- \S\ref{sec:advanced-distillation}).  Getting the
order wrong sticks the label on the \emph{wrong} physical stream and the
engine does \emph{not} warn --- the counts still match.  The only guard
is arity: declaring more names than the unit produces is a loud error.
To learn a unit's order, read its \code{producedStreams()} in the source
(glass-box) or its entry in the unit-operation catalogue
(\S\ref{sec:units}).

\paragraph{Recycles: tear streams.}  A recycle is an ordinary stream
whose producer sits downstream of its consumer.  Declare the
cycle-breaking stream (and, optionally, the recycle solver knobs) at the
top level of \file{flowsheetDict} (or in \file{solverDict}):

\begin{lstlisting}[language=dict]
tearStreams        ( recycle );
recycleSolver       Newton;      // Newton on the tears (default); or Wegstein
recycleTol          1e-5;
recycleMaxIter      120;
\end{lstlisting}

The tear's initial guess is its \file{0/} file, like any other stream
(\S\ref{sec:zero}).  \code{recycleSolver Newton} builds a
finite-difference Jacobian over the tear variables $(F, z_i, T)$ and
converges quadratically; \code{Wegstein} keeps the classical
fixed-point accelerator --- run both on
\file{tutorials/steady/flowsheets/process03\_recycle} and compare the
iteration counts.

\paragraph{Sectors --- the fractal flowsheet.}  A big plant is
organised as \emph{sectors} (composite sub-flowsheets), each a folder
with its own \file{system/flowsheetDict} and, when it needs one, its own
\file{constant/propertyDict} (a sector may run a different
thermodynamic world --- e.g.\ an electrolyte model in a brine section
and NRTL in a solvent-recovery section).  The root dict then assembles
members and named connections:

\begin{lstlisting}[language=dict]
sectors ( BRINE  EXTRACTION  CARBONATION );    // UPPERCASE folders
// or:   units ( cryst1 cryst2 recovery );     // dignified unit folders

connections
{
    liRichBrine { from BRINE/liRichBrine;  to EXTRACTION/liRichBrine; }
    brineFeed   { to BRINE/brineFeed; }        // to-only   = plant inlet
    tails       { from EXTRACTION/raffinate; } // from-only = plant outlet
}
\end{lstlisting}

The connection \emph{key} is the stream's one stable identity (and its
\file{0/} filename); \code{from}/\code{to} are producer/consumer ports.
A stream's \emph{role} is inferred from the shape of its edge --- a
\code{to}-only edge is an inlet, \code{from}+\code{to} is internal,
\code{from}-only is an outlet.  There is no boundary mini-language to
maintain.  The whole tree is flattened into one solver problem
(\code{plant.sector.unit} names); the hierarchy is for authoring and
namespacing, never a solver-within-a-solver.  See
\file{tutorials/plant/} for worked multi-sector cases.

\subsection{The \file{0/} directory --- one file per stream}\label{sec:zero}

In a steady case a stream's state lives in its own file on disk ---
never inside \file{flowsheetDict}.  The \file{0/} directory holds the
\textbf{complete initial state}: for an \emph{inlet} the file is your
boundary specification; for an \emph{internal} or \emph{outlet} stream
it is an initial estimate the solver overwrites.  (This contract
belongs to the steady domain --- \code{choupoSolve}; the time-dependent
and props binaries have no stream registry to seed.)  The file for the
flash tutorial's feed:

\begin{lstlisting}[language=dict]
// 0/feed
componentMolarFlows
{
    benzene    40 kmol/h;
    toluene    60 kmol/h;
}

T    370 K;
P    1.0 bar;
\end{lstlisting}

\paragraph{Material --- pick exactly one form.}  Four equivalent ways to
state what flows; a second form in the same file is a fatal conflict:

\begin{lstlisting}[language=dict]
componentMolarFlows { water 86 kmol/h;  NaCl 12 kmol/h; }             // (a)
componentMassFlows  { water 1549.3 kg/h;  NaCl 701.3 kg/h; }          // (b)
molarFlow 100 kmol/h;   moleFractions { water 0.86;  NaCl 0.14; }     // (c)
massFlow 2250.6 kg/h;   massFractions { water 0.6884; NaCl 0.3116; }  // (d)
\end{lstlisting}

Forms (a)/(b) are the natural process-design description (``mix
40~kmol/h of A with 60~kmol/h of B''); (c)/(d) are the textbook total +
composition.  Mass converts through each component's MW.  Fractions
must close: if $|\sum_i x_i - 1| > 10^{-6}$ the reader stops with a
fatal error naming the stream and the actual sum --- there is no silent
renormalisation of a state file.

\paragraph{Thermodynamic closure --- T and P.}  Give \code{T} and
\code{P}; the phase split is then a \emph{result} (Duhem's theorem: a
stream of known composition is fixed by exactly two state variables).
Do not write a decorative \code{vaporFraction 0;} or \code{1;} --- the
engine recovers the phase itself.  Two honest pins exist for the cases
$(T,P,z)$ cannot close:

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
  \item \code{vaporFraction 0.30;} --- a genuine \emph{two-phase split}
        on the saturation manifold, where $T$ and $P$ are not
        independent and the split is a real degree of freedom;
  \item \code{phase gas;} (or \code{phase liquid;}) --- a
        \emph{phase-intent} pin for a feed the cheap screen cannot
        classify (e.g.\ steam in a hot gas mixture: all vapour, but $T$
        is below water's $T_c$ so ``permanent gas'' cannot be proved).
        Hot air or combustion gas above every component's $T_c$ needs
        \emph{nothing} --- the engine's $T_c$ screen recovers
        $vf = 1$ on its own.
\end{itemize}

\paragraph{Multiphase decomposition.}  A slurry or partly-solid stream
adds a \code{phases \{\}} block that decomposes the overall material and
must sum back to it exactly (validated on read); a particle-size
distribution rides in a \code{psd \{\}} block.  A utility stream may
carry \code{category <word>;} so the utilities report can aggregate
consumption by header (\S\ref{sec:reports}).

\paragraph{The completeness contract.}  $N$ streams in the flowsheet
graph $\;=\;$ $N$ files in \file{0/}.  A missing or an orphan file is
\textbf{fatal} for \code{choupoSolve} --- there is no partial state.
The run announces the check:

\begin{lstlisting}
[state] seeded 3 stream(s) from 0/ via the canonical manifest
[state] 0/ complete: 3 graph stream IDs == 3 state files
\end{lstlisting}

\paragraph{\file{converged/} --- the answer, in the same shape.}  After
a converged solve the engine writes the full solution as
\file{converged/<stream>} files --- same grammar, one file per stream,
now carrying the solved numbers (and the resolved \code{phase}).  The
log confirms: \code{[state] wrote converged/}.  A drilled-in sector case
is seeded from its parent's \file{converged/} state, so a sub-case is
reproducible from disk alone.

\subsection{Materialising \file{0/}: \code{bin/choupo-init0}}\label{sec:init0}

You author only the \emph{domain inlets} (and a seed for each declared
tear); the tool writes every remaining stream file by explicit
propagation through the topology:

\begin{lstlisting}
bin/choupo-init0 <caseDir>            # writes the missing 0/ files
bin/choupo-init0 <caseDir> --force    # regenerates internal/outlet estimates
\end{lstlisting}

The rules are strict and announced: an authored inlet file is never
touched; an unseeded recycle is a hard error naming the file you must
author (feed propagation is only honest on an acyclic graph); and an
incomplete \file{0/} handed to \code{choupoSolve} stays fatal --- this
tool is the sanctioned way out.  The generated files are estimates the
solver overwrites; they exist so that the initial state is explicit,
inspectable, and on disk --- never an invisible in-memory guess.

\begin{keyidea}
\file{flowsheetDict} = topology.  \file{0/} = numbers.
\file{converged/} = the answer.  Every stream's state is a file you can
open, diff, and version-control --- the same glass-box rule that governs
the solver's iterations governs its data.
\end{keyidea}

\subsection{Units on every scalar}\label{sec:units-suffixes}

Choupo uses canonical SI internally: Pa, kmol/s, K, m, m\textsuperscript{2},
m\textsuperscript{3}, J, kg.  You almost never want to type those raw,
so any scalar in any dict may carry a unit suffix.  The parser converts
the value to SI on the fly and remembers its physical dimension --- a
typo in the unit is caught with a clear error before the solver runs.

\paragraph{Three accepted forms.}

\begin{lstlisting}[language=dict]
P    1     bar;                    // (a) named-unit shortcut
A_w  [-1  2  1  0  0]  1.5e-12;    // (b) bracket form:
                                   //     [M L T Theta N] exponents + SI value
R    0.5;                          // (c) raw SI (no declaration)
\end{lstlisting}

Form (a) is what you will type 95\,\% of the time.  Form (b) is the
escape hatch for quantities with no short named unit (membrane
permeabilities, exotic transport coefficients).  Form (c) says ``trust
me, the number is in canonical SI'' --- the parser cannot check
anything, so a typo here passes silently.

\paragraph{Named units shipped by Choupo.}  The most useful aliases
(the full registry of $\sim$90 units lives in \file{src/core/Units.cpp}):

\begin{center}\small
\begin{tabular}{lll}
\toprule
Quantity            & Canonical SI   & Named-unit aliases \\
\midrule
pressure            & Pa             & \code{Pa, kPa, MPa, bar, atm, psi, mmHg, torr} \\
molar flow          & kmol/s         & \code{kmol/s, kmol/h, mol/s, mol/h} \\
mass flow           & kg/s           & \code{kg/s, kg/h, g/s, g/h, t/h} \\
volumetric flow     & m\textsuperscript{3}/s & \code{m3/s, m3/h, L/s, L/h, L/min} \\
temperature         & K              & \code{K, degC, C, degF, F} \\
length              & m              & \code{m, mm, cm, km, in, ft} \\
area / volume       & m\textsuperscript{2} / m\textsuperscript{3} & \code{m2, cm2, mm2;~ m3, L, mL} \\
time                & s              & \code{s, min, h, day} \\
mass                & kg             & \code{kg, g, t, ton, tonne} \\
energy / power      & J / W          & \code{J, kJ, MJ;~ W, kW, MW\_power} \\
heat-transfer coeff.& W/(m\textsuperscript{2}K) & \code{W/m2/K, W/(m2.K), kW/(m2.K)} \\
velocity            & m/s            & \code{m/s, cm/s, mm/s} \\
membrane $A_w$      & m/(s$\cdot$Pa) & \code{m/s/Pa, m/s/bar} \\
viscosity           & Pa$\cdot$s     & \code{Pa.s, cP} \\
concentration       & kmol/m\textsuperscript{3} & \code{kmol/m3, mol/m3, mol/L, M} \\
molality            & mol/kg         & \code{mol/kg, mmol/kg} \\
density             & kg/m\textsuperscript{3}   & \code{kg/m3, g/cm3, g/mL, mg/L, ppm} \\
molar energy        & J/mol          & \code{J/mol, kJ/mol, J/kmol, kJ/kmol} \\
dimensionless       & ---            & \code{-}, \code{\%} \\
\bottomrule
\end{tabular}
\end{center}

Note the deliberate quirks: power is \code{MW\_power} (a bare \code{MW}
would collide with the mega-prefix logic of \code{MPa}), and
\code{degC}/\code{degF} are affine --- use \code{K} for temperature
\emph{differences}.

\paragraph{What you see on a bad unit.}  Asking for a pressure but
typing \code{P 1 kg;} produces:

\begin{lstlisting}
ERROR: Dictionary 'feed' ():
  parameter 'P' expected dimensions [1 -1 -2 0 0] (Pa)
  but the dict declared [1 0 0 0 0] (kg).  Check the unit
  suffix on this entry.
\end{lstlisting}

The same check guards every parameter the unit operations read
(\code{T}, \code{V\_R}, \code{area}, \code{W\_shaft},
\code{P\_perm},~\dots).

\paragraph{Output preferences.}  By default Choupo prints flows in
kmol/h, pressures in bar, temperatures in K --- the chem-eng
convention.  To override, drop a \code{units} block in
\file{controlDict}; an optional \code{precision} block tunes decimal
digits per quantity:

\begin{lstlisting}[language=dict]
units      { pressure atm;  flow kmol/s;  temperature degC; }
precision  { pressure 3;  temperature 2;  flow 4;  composition 4; }
\end{lstlisting}

The internals stay in SI; only the printed output changes.

\subsection{Optional dicts}

\begin{itemize}[leftmargin=*]
  \item \dict{solverDict} --- solver defaults per unit-op type,
        overridable by each unit: e.g.\
        \code{isothermalFlash \{ tolerance 1e-8;\ maxOuterIter 50; \}}.
  \item \dict{outerDict} --- sweeps, design specs, optimisation
        (\S\ref{sec:outer}).
  \item \dict{postDict} --- sizing + costing chain (\S\ref{sec:post}).
  \item \dict{constant/reactions} --- the named-reaction library
        referenced by reactors: \code{reaction esterification;} or
        \code{reactions ( r1 r2 );} (\S\ref{sec:reactors}).
\end{itemize}

% =====================================================================
\section{The thermodynamic foundation: \file{constant/propertyDict}}\label{sec:thermo}

\begin{whatyoulllearn}
How a case declares its components and selects the models that turn
$(T,P,z)$ into $K$-values, enthalpies, and transport properties --- and
how the run log announces every choice it makes.
\end{whatyoulllearn}

Every case carries a \file{constant/propertyDict} declaring the
component list and the thermodynamic models.  The minimum viable file:

\begin{lstlisting}[language=dict]
components       ( benzene  toluene );

activityModel    { model ideal; }        // gamma_i = 1  (Raoult)
equationOfState  { model idealGas; }     // phi_i = 1
\end{lstlisting}

Components are resolved \emph{by name}, with an explicit evidence order:
\textbf{case-local overlay $>$ standard $>$ local}.  There are exactly two
data homes, and the distinction is deliberate:
\begin{itemize}
  \item \textbf{\file{data/standards/}} --- \emph{curated and public}.  The
    committee-managed reference set (194 species); shipped in the repository,
    every value carrying primary provenance.
  \item \textbf{\file{data/local/}} --- \emph{yours and private}.  A
    \emph{gitignored} working tier for data you import, license or estimate.
    The repository ships it \textbf{empty} (a \file{README} and a
    \file{.gitkeep}); the engine reads it when present.  A value resolved from
    it is announced \code{[local]\ UNVERIFIED} and carries its source, licence,
    import digest and method --- it is never presented as equivalent to a
    standard.
\end{itemize}
Data is thus either curated-and-public or yours-and-private: there is no public
``proposed'' middle tier (retired 2026-07).  A case-local file in
\file{<case>/constant/components/} overlays either home and always wins
(\S\ref{sec:build-own}).

You populate \file{data/local/} yourself --- for example with the reproducible
ChemSep~8.3 importer (\file{bin/curate/chemsep\_to\_choupo.py}), which writes
there.  This is why the public repository redistributes \emph{no} third-party
property values: for coverage beyond the curated standards it ships Choupo's own
open group-contribution estimates under \file{data/groupEstimative/} (clearly
flagged as estimates), and you import any measured or third-party data into your
private \file{data/local/}.
Imported literature constants and predictive closures stay distinguishable in
the provenance --- an imported boiling point is not silently upgraded because
an Ambrose--Walton closure makes the component flashable, and missing heat
capacities or formation data stay missing (the gap report says what the model
still needs).

\subsection{Activity models (liquid $\gamma$)}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\code{model}  & \textbf{Notes} \\
\midrule
\code{ideal}  & $\gamma_i = 1$ --- the Raoult limit.  Hydrocarbons, dilute systems. \\
\code{NRTL}   & Renon--Prausnitz.  Handles LLE.  Needs binary pairs. \\
\code{Wilson} & Cheaper than NRTL but structurally \emph{cannot} represent a liquid split. \\
\code{UNIQUAC}& Abrams--Prausnitz; the model many published kinetic studies regressed against. \\
\code{UNIFAC} & \emph{Predictive} --- $\gamma$ from molecular groups, no fitted binaries.  The group decomposition lives in each component's \file{.dat}. \\
\bottomrule
\end{tabularx}
\end{center}

NRTL/Wilson/UNIQUAC pairs are supplied inline or resolved by name from
the pair catalogues.  Resolution is \textbf{inline/case-local $>$ standard
$>$ local $>$ ideal default} (the same two-tier model as components: the
public curated \file{data/standards/\allowbreak binaryPairs/\allowbreak <model>/}
bank, then your private \file{data/local/} bank, which the repository ships
empty).  A pair absent from every tier is the announced ideal default
($\tau=0$), never a silent one:

\begin{lstlisting}[language=dict]
activityModel
{
    model  NRTL;
    pairs
    (
        { i ethanol;  j water;
          a_ij -0.8009;  b_ij  246.18;      // tau_ij = a_ij + b_ij/T
          a_ji  3.4578;  b_ji -586.0809;
          alpha 0.30; }
    );
}
\end{lstlisting}

A proposed pair never shadows a standard pair.  The 748 staged ChemSep pairs
have passed identity, unit-conversion and provenance audits, but have not all
been independently validated against experimental data; the log therefore
marks their use \code{[proposed] UNVERIFIED}.  A pair with no parameters
anywhere does not abort the run --- it defaults to ideal ($\tau = 0$) so you
can build the foundation pair-by-pair --- but it is never silent: each
ideal-defaulted pair is announced with a \code{[thermo]} log line, and the GUI
colours the pair matrix by where each pair resolved.

\subsection{Equation of state (vapour $\varphi$ --- and, for a cubic, the liquid too)}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\code{model} & \textbf{Notes} \\
\midrule
\code{idealGas} & $Z = 1$, $\varphi_i = 1$.  Adequate at low pressure. \\
\code{SRK}      & Soave--Redlich--Kwong cubic; van der Waals one-fluid mixing; optional \code{binaryInteractions ( \{ i CO2; j CH4; kij 0.0973; \} )}. \\
\code{PR}       & Peng--Robinson cubic; better near-critical behaviour (CO$_2$). \\
\bottomrule
\end{tabularx}
\end{center}

The cubics read each component's $(T_c, P_c, \omega)$ from the
catalogue and deliver $Z$, residual $H$/$S$, and \emph{both} roots ---
so a cubic can also serve as the \emph{liquid} model in a $\varphi$--$\varphi$
package (below).

\subsection{Per-component correlations}

Vapour pressure defaults to Antoine
($\log_{10} P^{\mathrm{sat}} = A - B/(T+C)$), stored per component; a
corresponding-states alternative (\code{vaporPressure \{ model
AmbroseWalton; \}}) covers estimated components.  Ideal-gas and liquid
heat capacities are polynomials in $T$; species used at flame
temperatures carry two-range NASA-7 forms (\code{model NASA7}).  Each
component's \code{gibbsFormation \{ dHf\_298; s\_298; phase; \}} block
pins the formation datum --- the \code{phase} keyword (\code{gas} /
\code{liquid} / \code{solid}) declares on which rung of the enthalpy
ladder the value is tabulated, and every energy balance that crosses a
reactor requires it (\S\ref{sec:reactors}).

\subsection{Multiple liquid phases (LLE / VLLE)}

For liquid--liquid work, declare the phases explicitly:

\begin{lstlisting}[language=dict]
components ( water  nButanol );
phases
(
    { name vapour;   type vapor;   eos      { model idealGas; } }
    { name aqueous;  type liquid;  activity { model NRTL; pairs (...); } }
    { name organic;  type liquid;  activity { model NRTL; pairs (...); } }
);
\end{lstlisting}

The flash then searches for the liquid split by direct Gibbs-energy
minimisation (\S\ref{sec:flash-family}) --- the legacy single-liquid
form builds only one liquid and will not split.

\subsection{Transport properties}

Units that need viscosity, conductivity, diffusivity, or surface
tension (pipes, dryers, tray hydraulics, exchanger rating) read them
from a \code{transport} block of selectable sub-models:

\begin{lstlisting}[language=dict]
transport
{
    viscosity            { model Chung; }        // gas mu, from Tc/Pc/omega/MW
    thermalConductivity  { model Eucken; }
    diffusivity          { model Fuller; }       // needs diffusionVolume
    liquidViscosity      { model Vogel; }        // or Andrade
    liquidConductivity   { model SatoRiedel; }   // or Latini
    liquidDiffusivity    { model WilkeChang; }   // or Scheibel
    surfaceTension       { model BrockBird; }
}
\end{lstlisting}

When a unit needs a property the package does not supply, the solver
stops with a clear error naming the missing model --- it never invents
a value.

\subsection{Dissolved gases: Henry's law}

A gas solute (CO$_2$, NH$_3$, O$_2$, H$_2$S, SO$_2$,~\dots\ in water)
takes the infinite-dilution Henry reference: declare
\code{solvent water;} and the engine consults the Henry-pair catalogue
(\file{data/standards/henrysLaw/<solute>-<solvent>.dat}, van~'t~Hoff
temperature dependence).  The absorber/stripper and the dilute-solution
flash world run on this.  Each engaged pair is announced with its
constants and source (\code{[henry]} log lines).

\subsection{The \code{propertyPackage} manifest and the four VLE worlds}\label{sec:vle-worlds}

The flat \file{propertyDict} above is the everyday form.  Cases that
need a declared, verifiable assembly --- Henry worlds, $\varphi$--$\varphi$
high-pressure work, electrolytes --- write the same file as a typed
manifest: \code{recordType propertyPackage;} plus a
\code{propertyMethods} block whose \emph{liquid slot selects the whole
VLE structure}:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}llX@{}}
\toprule
\code{propertyMethods.liquid} & \textbf{World} & \textbf{K-value} \\
\midrule
\code{activity.<Model>}   & $\gamma$--$\varphi$ & $K_i = \gamma_i P_i^{\mathrm{sat}} / (\varphi_i P)$ \\
\code{solution.henryDilute} & dilute solution & solvent on Raoult; each solute on the Henry form $y\varphi P = x\gamma^* H(T)\,e^{v_\infty (P-P_s)/RT}$ \\
\code{eos.<Model>}        & $\varphi$--$\varphi$ & $K_i = \varphi_i^L/\varphi_i^V$ --- the same cubic's two roots \\
\code{electrolyte.pitzer} / \code{electrolyte.eNRTL} & speciation & ionic activity + osmotic coefficients on the aqueous-ion reference \\
\bottomrule
\end{tabularx}
\end{center}

\begin{lstlisting}[language=dict]
recordType     propertyPackage;
schemaVersion  1;
name           flash08_co2Water;
components     ( water CO2 );
propertyMethods
{
    liquid  solution.henryDilute;
    vapour  builtin.idealGas;
}
solution   { solvent water;  solutes ( CO2 ); }
parameters
{
    henryPairs { CO2-water "data/standards/henrysLaw/CO2-water.dat"; }
}
\end{lstlisting}

Declared parameter files are \emph{verified at assembly} --- a
declared-but-missing pair refuses loudly, naming the entry to add ---
and the run announces the selected world
(\code{[builder] VLE world: phi-phi (SRK both phases)}).  Mixing two
different cubics across the two phases is refused: one Gibbs surface
per phase, never two.  Reference tutorials:
\file{flash08\_co2\_water\_package} (Henry) and
\file{flash09\_n2ch4\_stryjek} ($\varphi$--$\varphi$ with a declared
$k_{ij}$).

\subsection{Pure fluids: IAPWS-IF97 steam}

For water-dominated power cycles the package can route a pure component
through the hand-rolled IAPWS-IF97 steam kernel:

\begin{lstlisting}[language=dict]
pureFluids { water { method IF97; } }
\end{lstlisting}

The Rankine tutorial \file{tutorials/steady/power/rankine02\_water}
closes its first law to $10^{-12}$ on IF97 --- and its ideal-gas twin
\file{rankine01\_water} is kept on purpose, to show what an ideal-gas
water \emph{cannot} do (there is no two-phase dome to cross).

\subsection{Per-unit \code{thermo \{\}} override and model boundaries}\label{sec:thermo-override}

A unit may carry a \code{thermo \{...\}} block that \emph{replaces} the
global models for that unit only (components stay global):

\begin{lstlisting}[language=dict]
{ name turbine;  type turbine;  in feed;  outputs ( expanded );
  operation { W_shaft -3 kW;  eta 0.8; }
  thermo    { equationOfState { model SRK; } } }    // real gas here only
\end{lstlisting}

A stream crossing from model X into model Y is re-interpreted at the
boundary: $(T,P,z)$ are held and each unit recomputes its own enthalpy,
so $H$ \emph{steps} at the boundary --- visibly, in the printed
enthalpy.  \textbf{$H$ is the conserved truth; $T$ is the
model-dependent readout.}  The solver never silently nudges $T$ to
paper over the step; an opt-in model-boundary audit prints
$\Delta H = H_{\mathrm{down}} - H_{\mathrm{up}}$ into the first-law
ledger.  Default to one consistent global model; reach for the override
only when the physics demands it.  The run log is loud either way ---
every unit prints \code{thermo: inherited (global package)} or
\code{LOCAL override --- ...}.

\subsection{Self-contained cases}\label{sec:self-contained}

A case can be sealed so it runs with \emph{no} installation catalogue:

\begin{lstlisting}
bin/choupo-import <case>     # materialise the dependency closure into
                             # constant/propertyData/ + a sha256 manifest
\end{lstlisting}

When \file{constant/propertyData/} is present the runtime reads the
frozen snapshot only (the log states ``installation catalogue NOT
consulted'') --- the definitive test is moving the case out of the repo
and watching it still run.  Research cases and anything you archive
with a thesis should be sealed.

% =====================================================================
\section{Built-in unit operations}\label{sec:units}

\begin{whatyoulllearn}
The complete catalogue of steady-state unit operations --- what each
\code{type} does, which \code{model} sub-selectors exist, and what the
\code{operation} block must contain.  Aliases
(\code{flash}, \code{column}, \code{FUG}, \code{REquil},
\code{boiler}, \code{condenser}, \code{MHeatX}, \code{membraneSW},
\code{extract}) name the same classes.
\code{UnitOperation::availableTypes()} prints the live registry.
\end{whatyoulllearn}

\subsection{Quick reference}

\begin{center}\footnotesize
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash}p{0.26\linewidth}
                                 >{\raggedright\arraybackslash}X
                                 >{\raggedright\arraybackslash}p{0.27\linewidth}@{}}
\toprule
\code{type}            & \textbf{What it does}    & \textbf{Key spec} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Flash and saturation}} \\
\code{isothermalFlash} \textit{(alias \code{flash})} & VL / LL / VLLE flash at fixed $(T,P)$; duty $Q$ is a result & \code{T}, \code{P} optional (inherit feed); \code{phaseSet VL|LL|VLLE} \\
\code{adiabaticFlash}  & $H_{\mathrm{out}} = H_{\mathrm{in}}$; solves $T$ & \code{P} \\
\code{bubbleT} / \code{dewT} & Saturation temperature at fixed $P$ & \code{P} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Reactors (\S\ref{sec:reactors})}} \\
\code{conversionReactor} & Specified per-pass conversion; stoichiometric & \code{conversion} (or \code{conversions (...)}) \\
\code{equilibriumReactor} \textit{(alias \code{REquil})} & Named reactions driven to simultaneous $K_p(T)$ equilibrium & \code{reactions (...)}; \code{T} \\
\code{gibbsReactor}    & Equilibrium by Gibbs minimisation over species & \code{elements}, \code{species}; models \code{elementPotential} / \code{reactiveFlash} / \code{directMin} \\
\code{cstr}            & Stirred tank; kinetics; 1 or N reactions & \code{V\_R}; \code{thermalMode} \\
\code{pfr}             & Plug flow, RK4 along volume; 1 or N reactions & \code{V\_R}, \code{nSteps}; \code{thermalMode} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Heat transfer (\S\ref{sec:heat-family})}} \\
\code{heater}          & 1-stream duty device; $T_{\mathrm{out}}$ is a result & \code{Q} \\
\code{phaseChanger} \textit{(aliases \code{boiler}, \code{condenser})} & Crosses the saturation dome; $Q$ a result & exactly one of \code{Q}, \code{outletT}, \code{outletQuality}, \code{outletState}, \code{superheat}, \code{subcool}; or \code{model geometry} \\
\code{heatExchanger}   & 2-stream $\varepsilon$-NTU rating & \code{area}, \code{U}; or \code{model geometry} / \code{model design} \\
\code{multiStreamHX} \textit{(alias \code{MHeatX})} & N hot + M cold in one shell; balance + internal pinch & per-stream \code{outlet \{...T...\}} targets \\
\code{evaporator}      & Single-effect; steam chest is a real stream & \code{area}, \code{U} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Rotating equipment (\S\ref{sec:rotating})}} \\
\code{compressor} / \code{turbine} & $W_{\mathrm{shaft}}$ + $\eta$; $P_{\mathrm{out}}$, $T_{\mathrm{out}}$ results & \code{W\_shaft}, \code{eta} \\
\code{pump}            & Incompressible; closed form & \code{eta} + one of \code{P\_out} / \code{dP} / \code{W\_shaft} \\
\code{electricLoad}    & Sink for shaft work via an energy wire & \code{energyInputs}, \code{eta} \\
\bottomrule
\end{tabularx}
\end{center}

\begin{center}\footnotesize
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash}p{0.26\linewidth}
                                 >{\raggedright\arraybackslash}X
                                 >{\raggedright\arraybackslash}p{0.27\linewidth}@{}}
\toprule
\code{type}            & \textbf{What it does}    & \textbf{Key spec} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Columns and staged separations (\S\ref{sec:columns})}} \\
\code{distillationColumn} \textit{(alias \code{column})} & Rigorous MESH; models \code{WangHenke} / \code{simultaneous}; multi-feed, side draws, Murphree, hydraulics, reactive & \code{nStages}, \code{feedStage}, \code{refluxRatio} or \code{distillateRate}, \code{P} \\
\code{shortcutColumn} \textit{(alias \code{FUG})} & Fenske--Underwood--Gilliland design & keys, recoveries, \code{refluxFactor}, \code{P} \\
\code{absorber} / \code{stripper} & Counter-current staged gas--liquid contact (Henry) & \code{stages} \\
\code{extractor} \textit{(alias \code{extract})} & Counter-current staged LL extraction (Gibbs-split stages) & \code{stages} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Membranes and sorption separations (\S\ref{sec:membranes})}} \\
\code{spiralWoundModule} \textit{(alias \code{membraneSW})} & NF/RO element or train (solution-diffusion + film model) & \code{membrane}, \code{area}, \code{P\_perm} \\
\code{ionExchanger}    & Cation-exchange softener (Gaines--Thomas) & \code{resin}, \code{CEC} \\
\code{electrodialysisStack} & N cell pairs; Faraday transfer, Nernst + ohmic voltage & \code{N\_cellpairs}, \code{current}, geometry \\
\code{psa}             & Pressure-swing adsorption (equilibrium-theory shortcut) & adsorbent isotherms, \code{bedCapacity}, \code{eta} \\
\midrule
\multicolumn{3}{@{}l}{\emph{Solids (\S\ref{sec:solids})}} \\
\code{cyclone}         & PSD-aware gas--solid classifier; 5 models (\code{Lapple},~\dots, \code{Muschelknautz}) & \code{bodyDiameter},~\dots \\
\code{bagFilter}       & Cake filtration; $\Delta P$ headline & \code{filterArea},~\dots \\
\code{gasSolidSplitter}& Ideal spec'd split (PSD unchanged) & \code{solidsRecovery} \\
\code{pneumaticConveyor} & Dilute-phase conveying line; $\Delta P$ + saltation check & \code{geometry \{ D L dz \}} \\
\code{crystalliser}    & Models \code{equilibrium} / \code{MSMPR} / \code{FVM}; magma or cake+liquor outputs & \code{operatingTemperature}; kinetics for MSMPR \\
\code{sprayDryer} / \code{solidDryer} / \code{evaporativeDryer} & Gas--solid drying (\S\ref{sec:advanced-spray}) & atomiser, chamber; sorption isotherm \\
\midrule
\multicolumn{3}{@{}l}{\emph{Flow primitives (\S\ref{sec:primitives})}} \\
\code{mixer}           & $N \to 1$ adiabatic merge (or declared-$T$) & --- \\
\code{splitter}        & $1 \to N$ tee at spec'd fractions & \code{fractions (...)} \\
\code{valve}           & Isenthalpic throttle & \code{P} \\
\code{pipe}            & Darcy--Weisbach line; $\Delta P$ a result; two-phase aware & \code{geometry \{ D L roughness dz \}} \\
\bottomrule
\end{tabularx}
\end{center}

Each unit also accepts top-level solver-tolerance overrides (same keys
as \file{solverDict}) and the per-unit \code{thermo \{\}} override
(\S\ref{sec:thermo-override}).

% ---------------------------------------------------------------------
\subsection{Flash and saturation}\label{sec:flash-family}

The isothermal flash is the pedagogical core: Newton on the
Rachford-Rice equation in $V/F$, wrapped in a Wegstein outer loop that
refreshes $\gamma_i(x)$.  A flash is a \emph{knob-less separator} ---
with an empty \code{operation \{\}} it inherits $T$ and $P$ from the
feed and just separates; pin either or both to hold the drum state:

\begin{lstlisting}[language=dict]
operation { T 370 K;  P 1 bar;  phaseSet VL; }   // VL (default), LL, or VLLE
\end{lstlisting}

\paragraph{Degrees of freedom --- $Q$ is a result.}  A flash is fixed by
exactly two numbers (Duhem's theorem), here $T$ and $P$.  The duty is
therefore a \emph{result, never a spec}: a flash \emph{gives} a $Q$
(read the \code{Q\_kW} KPI), it never \emph{takes} one.  To impose a
known heat and then separate, chain \code{heater(Q)} $\to$
\code{flash}; for a target outlet temperature, wrap the heater in a
DesignSpec on \code{\$Q} (\S\ref{sec:designspec}).  The
\code{adiabaticFlash} is the one duty that needs no external device:
$Q = 0$, spec \code{P}, and the outer Newton finds $T$.

\paragraph{LL and VLLE.}  For symmetric $\gamma$-models a
liquid--liquid flash solved on the $K$-value fixed point collapses onto
the trivial $x_\alpha = x_\beta$ saddle --- so Choupo's LL and VLLE
paths minimise the Gibbs energy of mixing \emph{directly}
(multi-start Nelder--Mead on the phase-split simplex).  Run
\file{tutorials/steady/absorption/extraction01\_waterButanol} (LL) and
\file{tutorials/steady/thermoTest/vlle03\_audit\_artificial} (a
constructed three-phase global minimum) to watch it work.
\code{bubbleT} and \code{dewT} are the 1-D saturation solves
($\sum K_i x_i = 1$ and $\sum y_i/K_i = 1$).

% ---------------------------------------------------------------------
\subsection{The reactor family}\label{sec:reactors}

\begin{whatyoulllearn}
Six reactors, one ladder of knowledge: specify the outcome
(\code{conversionReactor}), let thermodynamics
decide (\code{equilibriumReactor}, \code{gibbsReactor}), or let kinetics
decide (\code{cstr}, \code{pfr} --- and, in time,
\code{batchReactor}, \S\ref{sec:batch}).  Pick the rung that matches
what you actually know.
\end{whatyoulllearn}

All reactors share two contracts.  First, \textbf{one reaction
grammar}: stoichiometry and kinetics live once, in
\file{constant/reactions}, and a unit references them ---
\code{reaction <name>;} for one, \code{reactions ( r1 r2 );} for a
network --- never repeated inside the unit.  Second, \textbf{one
enthalpy datum}: the heat of reaction is computed from each species'
\code{gibbsFormation} block (the elements datum),
$\Delta H_{\mathrm{rxn}}(T) = \sum_i \nu_i h_i(T)$, in every reactor,
steady or dynamic.  There is no \code{dH\_rxn} input to a steady
reactor --- a reactions-dict \code{dH\_rxn} key is an announced override
honoured only for toy components that deliberately lack formation data.

\subsubsection{\code{conversionReactor} --- the outcome is a spec}

You state the per-pass conversion of the limiting reactant; the outlet
follows stoichiometry.  No kinetics, no residence time --- the honest
choice for a gas-phase reactor whose conversion is set by the catalyst,
and the right reactor when a \emph{recycle} is the point of the
flowsheet (it keeps a finite per-pass conversion).

\begin{lstlisting}[language=dict]
operation { conversion 0.75;  T 895 K; }     // single reaction
reaction  toluene_hda;                        // stoichiometry + limitingReactant
\end{lstlisting}

With several reactions, each extent is a spec --- selectivity becomes a
declared property of the catalyst:

\begin{lstlisting}[language=dict]
operation
{
    T 650 K;
    conversions
    (
        { reaction main;  conversion 0.60; }    // fraction of the FEED limiting reactant
        { reaction side;  extent 0.04; }        // or a direct extent
    );
}
reactions ( main side );
\end{lstlisting}

Tutorials: \file{conversion01\_hda},
\file{conversion02\_parallel\_selectivity}.

\subsubsection{\code{equilibriumReactor} (alias \code{REquil}) --- named reactions to equilibrium}

You give a \emph{set of reactions}; each is driven to simultaneous
chemical equilibrium via its $K_p(T)$, computed from the species'
formation data ($K = e^{-\Delta G^\circ/RT}$) --- no kinetics, no
residence time.  The coupled system
$K_{p,j} = \prod_i (p_i/P^\circ)^{\nu_{ij}}$ over all extents is solved
in log form by the multivariate Newton.  This is the reforming / shift
/ synthesis workhorse --- and the tool to use when you want to restrict
\emph{which} reactions equilibrate (something the Gibbs reactor, which
sees only species, cannot do).

\begin{lstlisting}[language=dict]
{ name reformer;  type equilibriumReactor;
  in feed;  outputs ( effluent );
  operation { T 1000 K; }
  reactions ( steamReforming  waterGasShift ); }   // stoichiometry only; no kinetics
\end{lstlisting}

KPIs: \code{Kp\_<name>}, \code{extent\_<name>\_kmol\_h},
\code{conversion\_<name>}, \code{Q\_kW}.  Tutorial:
\file{tutorials/steady/reactors/\allowbreak equil01\_reforming}
(steam-methane reforming and shift together).

\subsubsection{\code{gibbsReactor} --- equilibrium without naming reactions}

Direct Gibbs-energy minimisation over a declared species pool with
atom-balance constraints --- no reactions at all.  Three selectable
models: \code{elementPotential} (default; Lagrange multipliers),
\code{reactiveFlash} (multi-condensable + NRTL), and \code{directMin}
(multi-start Nelder--Mead on the extent null space).

\begin{lstlisting}[language=dict]
operation
{
    T 800 K;   P 1 bar;
    elements ( C H O );
    species
    (
        { name CH4; atoms ( 1 4 0 ); }
        { name H2O; atoms ( 0 2 1 ); }
        { name CO;  atoms ( 1 0 1 ); }
        { name CO2; atoms ( 1 0 2 ); }
        { name H2;  atoms ( 0 2 0 ); }
    );
}
\end{lstlisting}

An optional \code{approachTemperature <dT>;} evaluates the equilibrium
at $T + \Delta T$ --- the classical way to model an incomplete approach
for reversible systems without kinetics.  For the adiabatic flame
problem, wrap the reactor in a DesignSpec that varies \code{\$T} until
$H_{\mathrm{out}} = H_{\mathrm{in}}$
(\file{gibbs05\_adiabatic\_flame}: lean CH$_4$/air $\to$ 1492 K).

\subsubsection{\code{cstr} and \code{pfr} --- kinetics}

The kinetic pair.  The CSTR solves the design equation(s)
$\xi_j = r_j(\boldsymbol{\xi})\,V_R$ --- Newton-1D for one reaction, the
multivariate Newton for a network; the PFR marches
$\mathrm{d}F_i/\mathrm{d}V = \sum_j \nu_{ij} r_j$ with a manual RK4.
Both compute concentrations on the liquid molar volume --- for a
gas-phase reactor prefer \code{conversionReactor} or
\code{equilibriumReactor}.

\begin{lstlisting}[language=dict]
{ name reactor;  type cstr;   in feed;  outputs ( out );
  operation { V_R 0.004 m3; }
  reaction  esterification; }               // or:  reactions ( r1 r2 );
\end{lstlisting}

A reaction entry in \file{constant/reactions} carries stoichiometry
(with per-species orders) and an Arrhenius \code{kinetics} block;
\code{reversible true;} adds the reverse leg by detailed balance
($k_r = k_f / K_{eq}(T)$).  Multi-reaction networks are what make
\emph{selectivity} modellable --- and the CSTR/PFR contrast visible: for
the series network A$\to$B$\to$C at equal $\tau$,
\file{pfr03\_series\_selectivity} keeps 79\,\% of the intermediate
where \file{cstr03\_series\_selectivity} keeps 53\,\% --- back-mixing
eats the intermediate, a result you \emph{see}, not memorise.

\paragraph{Non-isothermal operation and multiplicity.}  Both take
\code{thermalMode isothermal} (default) $\cdot$ \code{adiabatic} $\cdot$
\code{heatExchange \{ UA; T\_coolant; \}} (CSTR) /
\code{heatExchange \{ U; areaPerVolume; T\_coolant; \}} (PFR):

\begin{lstlisting}[language=dict]
operation { V_R 0.004 m3;  thermalMode adiabatic;  T_guess 350; }
\end{lstlisting}

Because the reaction heat already lives inside $H$ on the elements
datum, an adiabatic reactor is simply $H_{\mathrm{out}} = H_{\mathrm{in}}$
--- no source term.  The CSTR eliminates the extents first (at fixed
$T$ the design equations are monotone), leaving one equation
$\varphi(T) = H_{\mathrm{out}}(T) - H_{\mathrm{in}} - Q_{\mathrm{ext}}(T)$
--- the textbook heat-generated vs heat-removed balance, which can have
\textbf{three roots}.  Choupo does not hide that: it \emph{scans} the
bracket, prints every steady state it finds, and reports the one
nearest \code{T\_guess}.  The PFR marches total enthalpy beside the
species and recovers $T(V)$ by inversion --- a hot spot is something you
see in the axial profile.

\begin{tryit}
Run \file{tutorials/steady/reactors/cstr05\_multiplicity}: three steady
states at 300.4 / 348.4 / 391.3~K, all printed, the \code{steadyStates}
KPI counting them.  Change \code{T\_guess} and the same reactor answers
differently --- that is the physics of ignition and extinction, not a
numerical defect.  Then run \file{cstr04\_adiabatic} and
\file{pfr04\_adiabatic} on the same esterification: +82~K vs +83~K ---
the difference is the back-mixing the CSTR pays for.
\end{tryit}

% ---------------------------------------------------------------------
\subsubsection{Catalytic kinetics: writing an LHHW rate law}\label{sec:lhhw}

\begin{whatyoulllearn}
How to write a Langmuir--Hinshelwood--Hougen--Watson rate law in
\file{constant/reactions}, what \code{basis activity} obliges you to,
and the one experiment that actually demonstrates product inhibition
(it is not the one you would try first).
\end{whatyoulllearn}

The same block is read by \code{cstr}, \code{pfr}, \code{batchReactor}
and \code{dynamicCSTR}: one \code{RateLaw}, four reactors, no
possibility of them disagreeing.  The Theory Guide \S\,``The rate law''
derives where the denominator comes from; here is how to write it.

\begin{lstlisting}
esterification_meoac
{
    limitingReactant   methanol;

    stoichiometry
    (
        { component aceticAcid;     nu -1;  order 1; }   // forward orders
        { component methanol;       nu -1;  order 1; }
        { component methylAcetate;  nu  1; }             // orderRev defaults to nu
        { component water;          nu  1; }
    );

    kinetics
    {
        type       LHHW;
        basis      activity;        // a_i = gamma_i x_i    (else `concentration`)

        A          8.497e6;         // k_f0   [mol/(g_cat.s)]
        Ea         6.047e4;         // J/mol
        A_rev      6.127e5;         // the REGRESSED reverse pair ...
        Ea_rev     6.373e4;         // ... or `reversible true;` for detailed balance

        adsorption
        {
            exponent    2;          // sites the rate-determining step occupies
            vacantSite  0;          // 1 (default) gives the classical `1 + sum K c`
            species
            (
                { component aceticAcid;     K0 0.05245454; }   // K_i = K0 exp(B/T)
                { component methanol;       K0 0.17601898; }   // B defaults to 0
                { component methylAcetate;  K0 0.05602127; }
                { component water;          K0 0.29086872; }
            );
        }
    }
}
\end{lstlisting}

Read it against the rate-law equation in the Theory Guide:
\code{exponent} is $p$, \code{vacantSite} is $v$, \code{K0}/\code{B}
are $K_i(T)$, and the \code{order} / \code{orderRev} entries are
$\alpha_i$ / $\beta_i$.  Omit the \code{adsorption} block entirely and
the denominator collapses to $1$: you are back to \code{type Arrhenius},
the plain power law, byte for byte.

\paragraph{Two things the engine refuses.}  You may not ask for
\code{reversible true} \emph{and} an \code{A\_rev}/\code{Ea\_rev} pair
--- pick the one you can defend.  And you may not ask for
\code{reversible true} on \code{basis activity}: detailed balance runs
through $K_c$, a concentration-basis equilibrium constant, and
pretending otherwise would put the equilibrium in the wrong place.
Give the regressed reverse pair instead.

\paragraph{\code{basis activity} obliges you to an activity model.}
The rate constants of a published LHHW law were regressed \emph{with} a
particular $\gamma$-model.  P\"opken's were regressed on UNIQUAC with
the binary parameters of their own Table 3 --- which are in the
standard catalogue, cited --- so \file{cstr07\_lhhw\_methylAcetate}
declares \code{activityModel \{ model UNIQUAC; \}} and nothing else
will do.  Swap in UNIFAC and $k_1$ stops being their $k_1$.

\paragraph{Rate constants per gram of catalyst.}  Add
\code{catalystLoading <kg/m\^{}3>;} to \code{operation}.  The reactor
prints it back and converts; with concentrations in
kmol/m\textsuperscript{3} the loading \emph{is} the factor.  Omit it
and the constants are taken as already volumetric.

\begin{pitfall}
The obvious way to see product inhibition --- delete the
\code{adsorption} block, keep $k_f$, compare --- \textbf{shows
nothing}.  The reaction gets \emph{faster}, because the adsorbed
activities are small and dividing by their square is a multiplication.
The constants were fitted for that denominator.

The honest experiment: hold every reactant fraction fixed and
\emph{charge the product}.  In \file{batch07\_lhhw\_methylAcetate},
starting from \code{methanol 0.4;\ aceticAcid 0.4;\ water 0.2;} instead
of \code{...\ water 0.0;} halves the initial rate,
$58.3 \to 29.0$~mmol/(L\,s).  The water is not a reactant.  It is
sitting on the sites.
\end{pitfall}

\begin{tryit}
Run \file{cstr07\_lhhw\_methylAcetate}: $X = 49\,\%$, kinetically
limited.  Raise \code{V\_R} to $1$~m\textsuperscript{3} and it climbs
to $76\,\%$ and stops.  Then run \file{batch07\_lhhw\_methylAcetate},
which integrates the \emph{same} law in time: it reaches $76.6\,\%$.
Nothing in either dict asserts that number.  It is where the rate law's
numerator vanishes --- the activity equilibrium $K_a = 25.50$ that
P\"opken's independent Eq.~17 recovers from the same eight constants.
Two reactors agreeing is how you check you typed the parameters right.
\end{tryit}

% ---------------------------------------------------------------------
\subsection{Heat transfer}\label{sec:heat-family}

\paragraph{\code{heater} --- the unit that \emph{takes} a $Q$.}
Single-stream, sensible-only; \code{operation \{ Q 100 kW; \}} (negative
cools) and $T_{\mathrm{out}}$ is the result.  ``I need
$T_{\mathrm{out}} = 360$~K'' is a DesignSpec on \code{\$Q}
(\S\ref{sec:designspec}) --- the outer loop is visible, the duty is
never back-computed in secret.

\paragraph{\code{phaseChanger} (aliases \code{boiler}, \code{condenser})
--- crossing the dome.}  Where a heater must stay off the saturation
line, the phase changer lands its outlet \emph{anywhere} on the
$(T, vf)$ surface --- subcooled, inside the dome, or superheated ---
wrapping the flash kernel and the dome-aware enthalpy
$h = (1-vf)h_f + vf\,h_g$.  $P$ is held from the inlet (or set); then
pick \emph{exactly one} outlet coordinate (a degrees-of-freedom guard
rejects zero or two):

\begin{lstlisting}[language=dict]
operation
{
    // pick exactly ONE:
    Q             -630 kW;            // duty given -> T, vf results
    outletT        368 K;             // -> Q result
    outletQuality  0.4;               // partial condenser -> Q result
    outletState    saturatedVapour;   // or saturatedLiquid
    superheat      25 K;              // T = Tsat(P) + dT
    subcool        10 K;              // T = Tsat(P) - dT
}
\end{lstlisting}

The labels \emph{boiler} / \emph{partial condenser} /~\dots\ are
emergent from the converged $vf$ and the sign of $Q$ --- not a model
switch.  With \code{model geometry;} the duty \emph{emerges} from real
heat-transfer physics instead: a \code{coolant \{\}} block selects
Nusselt (1916) laminar-film condensation on a tube bundle; a
\code{boiling \{\}} + \code{heatingMedium \{\}} pair selects Rohsenow
nucleate pool boiling with the Zuber critical-heat-flux ceiling --- a
design above CHF is hard-refused (burnout), and the Rohsenow surface
constant \code{Csf} has \emph{no default} and demands a
\code{citation} (a $\pm 100\,\%$ number must state its provenance).
Tutorials: \file{phasechange01\_partial\_condenser},
\file{condenser01\_film\_nusselt}, \file{reboiler\_water\_copper},
\file{rankine02\_water}.

\paragraph{\code{heatExchanger} --- rating, geometry, design.}  The
two-stream exchanger has three faces of one model:

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
  \item \textbf{default (\code{model epsNTU})} --- give \code{area} and
        \code{U}; duty and outlet temperatures are results
        ($\varepsilon$-NTU, \code{flow counter|co}).
  \item \textbf{\code{model geometry}} --- give the tube-bundle geometry
        (\code{tubeID/OD/Length}, \code{nTubes}, \code{passes},
        \code{shellID}, \code{baffleSpacing}, \code{tubePitch},
        \code{wallK} or \code{wallMaterial}) and the per-side
        correlations (\code{Gnielinski} tube-side, \code{Kern}
        shell-side by default) \emph{compute} $U$ and the area --- the
        glass-box answer to ``where does $U$ come from?''.  Both-side
        pressure drops are reported too.
  \item \textbf{\code{model design}} --- give the duty target (an outlet
        temperature) and the tube choices; the engine \emph{sizes} the
        bundle (bisects the tube count so the delivered $\varepsilon$-NTU
        duty meets the target) and then rates its own design.
\end{itemize}

The two-part tutorial pair
\file{hxWorkflow1\_design\_from\_duty} $\to$
\file{hxWorkflow2\_rate\_designed} is the standard industrial sequence:
converge the balance with a duty target, design the exchanger, then
rate the designed geometry (and ask the off-design question the design
step cannot answer).

\paragraph{\code{multiStreamHX} (alias \code{MHeatX}).}  N hot + M cold
streams in one adiabatic shell.  Every inlet carries its own outlet-$T$
target; the unit verifies the energy balance (a non-zero imbalance is
\emph{reported}, never absorbed into an invented outlet) and builds the
hot/cold composite curves to report the internal pinch
$\Delta T_{\min}$ --- a temperature cross is a hard warning.  Tutorial:
\file{mheatx01\_two\_hot\_one\_cold}.

\paragraph{\code{evaporator}.}  Single-effect with a \emph{real} steam
chest: \code{inputs ( feed steam )}, hardware \code{area} + \code{U};
the boiling temperature, vapour split, and operating pressure are all
results.  Multi-effect trains are just chained evaporators --- see
\file{evaporator02\_triple\_effect\_sugar} and the DesignSpec that sizes
them (\S\ref{sec:designspec}).

% ---------------------------------------------------------------------
\subsection{Rotating equipment and electric loads}\label{sec:rotating}

The credo: rotating hardware is specified by \emph{shaft power and
efficiency}, and the pressure it achieves is a result:

\begin{lstlisting}[language=dict]
operation { W_shaft  10 kW;   eta 0.75; }     // compressor: W > 0
operation { W_shaft  -3 kW;   eta 0.80; }     // turbine:    W < 0
\end{lstlisting}

The compressor/turbine run an outer Newton on $P_{\mathrm{out}}$
matching the isentropic enthalpy change scaled by $\eta$.  The
\code{pump} is incompressible and closed-form, so it accepts \code{eta}
plus exactly \emph{one} of \code{P\_out} / \code{dP} /
\code{W\_shaft} --- specifying the header it feeds is a local one-shot
inversion, not a DesignSpec.  A let-down is a \code{valve}, not a pump.

\paragraph{Energy wires.}  Shaft work travels on typed scalar wires
declared on the \emph{consumer}:

\begin{lstlisting}[language=dict]
{ name T1;  type turbine;  in steam;  outputs ( exhaust );
  operation { W_shaft -1500 kW;  eta 0.85; }
  energyOutputs ( { name shaft;  kind work;  expression "-W_shaft";  unit W; } ) }

{ name G1;  type electricLoad;
  operation { eta 0.97; }
  energyInputs ( { from T1.shaft;  kind work;  target W_shaft; } ) }
\end{lstlisting}

Several inputs to the same target \emph{sum} --- the gas-turbine
shaft-split, where one generator nets the turbine's gross work against
the compressor's load
(\file{tutorials/steady/power/combined02\_brayton\_rankine\_shaft}).
Typed ports are checked: a \code{work} wire cannot feed a \code{heat}
input.  Heat duties follow a different rule --- see
\S\ref{sec:energy-utilities}.

% ---------------------------------------------------------------------
\subsection{Distillation, absorption, extraction}\label{sec:columns}

\subsubsection{The rigorous column}

\code{distillationColumn} solves the full MESH stage equations with two
selectable numerical strategies (the \code{model} slot):

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
\item \code{WangHenke} (default) --- the sequential bubble-point method:
      a tridiagonal (Thomas) component-balance sweep at frozen $K$,
      then a bubble-point $T$ update per stage.  Cheap per pass,
      inspectable, but O(100--500) outer iterations, and it can step
      \emph{through} an azeotrope non-physically.  Use it for ideal /
      wide-boiling systems.
\item \code{simultaneous} (aliases \code{MESH},
      \code{NaphtaliSandholm}) --- Newton on the whole block at once
      with Armijo backtracking.  Quadratic ($\approx$5--6 iterations)
      and stable through azeotropes --- the required choice for
      non-ideal systems (NRTL ethanol/water:
      \file{column03\_azeotrope\_mesh}).
\end{itemize}

\begin{lstlisting}[language=dict]
{ name column01;  type distillationColumn;      // model WangHenke is the default
  in feed;  outputs ( distillate  bottoms );
  operation
  {
      nStages          12;
      feedStage         6;
      refluxRatio      2.5;                     // XOR with distillateRate
      distillateRate   50 kmol/h;
      P                1.01325 bar;
  } }
\end{lstlisting}

A column carries two intrinsic heat ports --- the reboiler (heating, at
the bottoms $T$) and the condenser (cooling, at the top $T$), reported
as \code{Q\_reboiler\_kW} / \code{Q\_condenser\_kW} and sized to plant
utilities by temperature level (\S\ref{sec:energy-utilities}).

\subsubsection{Multiple feeds, side draws, reactive stages}\label{sec:advanced-distillation}

\begin{whatyoulllearn}
How the \code{simultaneous} column handles \emph{several feeds},
\emph{side draws}, and a \emph{reaction on catalytic stages} --- with a
worked, paper-validated reactive-distillation case.
\end{whatyoulllearn}

\paragraph{Feeds are flowsheet streams.}  A multi-feed column declares
its feeds as \emph{streams} (\code{inputs ( feed acidFeed )});
\code{operation.feeds} only maps each stream to a stage.  This is
deliberate (\emph{information follows the streams}): a feed hidden
inside the operation block would be a side channel the plant-boundary
mass and energy balance never sees --- and the first law would not
close.

\paragraph{Side draws, and the order of \code{outputs}.}  A side draw
is an \emph{output} --- a liquid or vapour bleed from an intermediate
stage, its rate set in \code{operation.sideDraws}.  Because outputs
bind \emph{by position} (\S\ref{sec:flowsheetDict-streams}), the
column's emission order is fixed and your \code{outputs} list must
match it:
\[
   \underbrace{\code{distillate}}_{[0]},\;
   \underbrace{\code{bottoms}}_{[1]},\;
   \underbrace{\text{side draws by \emph{ascending stage}}}_{[2],[3],\dots}
\]
So a four-product column that draws liquid at stage 13 then 23 is wired
\code{outputs ( distillate bottoms drawC4 drawC5 )}: \code{drawC4}
(position 2) is the stage-13 draw, \code{drawC5} (position 3) the
stage-23 draw.  Swap the two names and the label lands on the wrong
stream --- the engine does \emph{not} warn.

\begin{center}
\begin{tikzpicture}[>=Stealth, font=\footnotesize, line width=0.4pt]
  \draw[fill=gray!5] (0,0) rectangle (2.0,6);
  \foreach \y in {1,2,3,4,5} \draw[gray!40] (0,\y) -- (2.0,\y);
  % feeds (by name, to a stage)
  \draw[<-] (0,4.6) -- ++(-1.5,0) node[left]{\code{feed1} \scriptsize(stage 10)};
  \draw[<-] (0,2.2) -- ++(-1.5,0) node[left]{\code{feed2} \scriptsize(stage 20)};
  % products (by position)
  \draw[->] (2.0,5.7) -- ++(1.5,0) node[right]{\code{[0] distillate}};
  \draw[->] (2.0,3.9) -- ++(1.5,0) node[right]{\code{[2] drawC4} \scriptsize(stage 13)};
  \draw[->] (2.0,1.9) -- ++(1.5,0) node[right]{\code{[3] drawC5} \scriptsize(stage 23)};
  \draw[->] (2.0,0.3) -- ++(1.5,0) node[right]{\code{[1] bottoms}};
  \node[align=center, font=\scriptsize\itshape, text=gray!60!black]
        at (1.0,-0.6) {feeds in \emph{by name}; products out \emph{by position}};
\end{tikzpicture}
\end{center}

\begin{lstlisting}[language=dict]
operation
{
    nStages 15;  refluxRatio 3.0;  distillateRate 42 kmol/h;  P 1.013 bar;
    feeds      ( { stream feed;  stage 5;  quality 1.0; }
                 { stream feed2; stage 11; quality 1.0; } );
    sideDraws  ( { stage 8;  phase liquid;  rate 15 kmol/h; } );
}
\end{lstlisting}

\paragraph{A reaction on the stages.}  A \code{reaction \{\}} block
turns the column reactive (\code{simultaneous} only; restricted to
mole-conserving $\sum \nu = 0$ reactions so the flow profile is
untouched).  The extent on each named catalytic stage is closed either
by chemical \code{equilibrium} (activity-based $K_a(T)$, van~'t~Hoff)
or by \code{kinetics} (a rate law $\times$ catalyst mass).

\paragraph{Worked case --- methyl acetate by reactive distillation.}
The classic industrial esterification:
\[
   \text{MeOH} + \text{HOAc} \;\rightleftharpoons\; \text{MeOAc} + \text{H}_2\text{O}.
\]
Acetic acid is fed \emph{above} the catalytic zone and methanol
\emph{below}, so they flow counter-currently through it; the light
product rises to the distillate, water sinks to the bottoms, and
removing both products pulls the equilibrium forward --- conversion far
above a single equilibrium reactor.

\begin{center}
\begin{tikzpicture}[>=Stealth, font=\footnotesize, line width=0.4pt]
  % column body
  \draw[fill=gray!4] (0,0) rectangle (1.8,9);
  % reactive (catalytic) zone
  \fill[blue!12] (0,2.2) rectangle (1.8,5.6);
  \node[align=center, text=blue!55!black] at (0.9,3.9)
       {\scriptsize catalytic\\[-1pt]\scriptsize zone\\[-1pt]\scriptsize (11--19)};
  % trays
  \foreach \y in {0.6,1.2,...,8.4} \draw[gray!50] (0.15,\y) -- (1.65,\y);
  % condenser
  \draw[fill=cyan!12] (0.4,9.4) rectangle (1.4,10.0);
  \node at (0.9,9.7) {\scriptsize cond.};
  \draw[->] (0.9,9.0) -- (0.9,9.4);
  \draw[->] (0.6,9.4) -- (0.6,8.7);
  \draw[->] (1.4,9.7) -- (3.0,9.7) node[right, align=left]
       {\scriptsize \textbf{distillate}\\[-2pt]\scriptsize methyl acetate ($x{=}0.93$)};
  % reboiler
  \draw[fill=orange!15] (0.4,-1.0) rectangle (1.4,-0.4);
  \node at (0.9,-0.7) {\scriptsize reb.};
  \draw[->] (0.9,-0.4) -- (0.9,0.0);
  \draw[->] (1.4,-0.7) -- (3.0,-0.7) node[right, align=left]
       {\scriptsize \textbf{bottoms}\\[-2pt]\scriptsize water ($x{=}0.83$)};
  % feeds
  \draw[->, red!70!black] (-1.6,6.0) -- (0,6.0)
       node[midway, above] {\scriptsize HOAc} node[midway, below]{\scriptsize (st.\ 10)};
  \draw[->, blue!70!black] (-1.6,1.8) -- (0,1.8)
       node[midway, above] {\scriptsize MeOH} node[midway, below]{\scriptsize (st.\ 20)};
  \node[red!60!black]  at (2.4,5.3) {\scriptsize $\downarrow$\,HOAc};
  \node[blue!60!black] at (2.4,2.6) {\scriptsize MeOH\,$\uparrow$};
  \node[align=center] at (0.9,10.4) {\scriptsize 25 equilibrium stages};
\end{tikzpicture}
\end{center}

The dict says exactly this --- two input streams mapped to stages
(their state authored in \file{0/feed} and \file{0/acidFeed}), plus the
reaction:

\begin{lstlisting}[language=dict]
units ( {
    name rdColumn;  type distillationColumn;  model simultaneous;
    inputs  ( feed  acidFeed );              // both feeds are flowsheet streams
    outputs ( distillate  bottoms );
    operation {
        nStages 25;  refluxRatio 2.1;  distillateRate 30.1 mol/h;  P 1.017 bar;
        feeds ( { stream feed;     stage 20; quality 1.0; }   // methanol below the zone
                { stream acidFeed; stage 10; quality 1.0; } );// acetic acid above it
    }
    reaction {
        stoichiometry ( { component methanol;      nu -1; } { component aceticAcid; nu -1; }
                        { component methylAcetate; nu  1; } { component water;      nu  1; } );
        reactiveStages ( 11 12 13 14 15 16 17 18 19 );
        equilibrium { Ka298 38.7; dHrxn -5670; }   // activity-based K_a(T), van't Hoff
    }
} );
\end{lstlisting}

Running it
(\file{tutorials/steady/distillation/column05\_reactive\_methylacetate})
converges by homotopy (solve the column non-reactive, then switch the
reaction on) and gives \textbf{96\,\% methanol conversion}, a
methyl-acetate-rich distillate and a water-rich bottoms --- the
\emph{equilibrium ceiling}.  The measured value (P\"opken \emph{et
al.}, \emph{Ind.\ Eng.\ Chem.\ Res.}\ 40 (2001) 1566) is 87.6\,\%: the
real column is slightly kinetically limited, so the equilibrium model
correctly brackets it from above (replace \code{equilibrium} with a
\code{kinetics} block to model the finite rate).  Because both feeds
are flowsheet streams, the plant-boundary energy balance closes
exactly.

\subsubsection{Real trays: Murphree efficiency}\label{sec:murphree}

Add \code{murphreeEfficiency 0.70;} to \code{operation}
(\code{WangHenke} only --- \code{simultaneous} refuses the keyword
rather than silently returning ideal stages) and \code{nStages} stops
counting ideal stages and starts counting \emph{real trays}:
$y_j = E\,K_j x_j + (1-E)\,y_{j+1}$, folded into an effective $K$.
$E$ is a \emph{declared} number --- O'Connell's correlation, a chart, or
plant data; the engine never invents it.  \file{column11\_murphree} is
\file{column09} plus that one line: the distillate falls from
$98.12\,\%$ to $95.05\,\%$ benzene, and at $E = 1$ every existing case
reproduces bit for bit.

\subsubsection{Tray hydraulics: will the column you converged actually run?}\label{sec:tray-hydraulics}

\begin{whatyoulllearn}
How to ask a converged column whether its trays can pass the traffic
they carry --- and how the same block, with one keyword removed, stops
rating the column and starts sizing it.
\end{whatyoulllearn}

A converged column tells you what separates.  It says nothing at all
about how wide it is, because nothing in the stage equations depends on
the diameter.  Two columns of very different width give the identical
stream table --- and one of them floods, which means in practice that it
separates nothing.  The Theory Guide \S\,``Tray hydraulics'' derives
why.

Add a \code{hydraulics \{\}} block inside \code{operation} (either
column model) and the column runs a rating pass over its own converged
profile:

\begin{lstlisting}[language=dict]
hydraulics
{
    trayType               sieve;
    // diameter            1.10;     // GIVE it -> rating.  OMIT it -> design.
    traySpacing            0.50;     // m
    weirHeight             0.050;    // m
    holeDiameter           5.0e-3;   // m
    holeAreaFraction       0.10;     // A_holes / A_active
    downcomerAreaFraction  0.12;     // A_downcomer / A_tower
    weirLengthFraction     0.77;     // l_weir / D
    orificeCoefficient     0.84;     // C_o -- YOU read this off the chart
    floodFraction          0.80;     // design target
    K2                     30.7;     // optional: enables the weep check
}
\end{lstlisting}

\paragraph{The thermo package must supply the surface tension.}
Entrainment is a contest between the vapour's drag on a droplet and the
surface tension holding it together, so the Souders--Brown flooding
velocity cannot be evaluated without $\sigma$.  Add
\code{transport \{ surfaceTension \{ model BrockBird; \} \}} to the
\file{propertyDict}, or declare a constant \code{sigma <N/m>;} inside
the hydraulics block.  The column refuses to guess; it will stop and
say so.

\paragraph{Two modes, one block.}  Give \code{diameter} and the column
\emph{rates}: it reports how close each tray runs to its flooding
velocity.  Omit \code{diameter} and it \emph{designs}: it finds the
width at which no tray exceeds \code{floodFraction} of flood, and the
widest tray sets the column.  This is the same rating/design duality as
the spray dryer (\S\ref{sec:spray-design}) --- one set of equations, run
in whichever direction you need.

\paragraph{Two numbers you must own.}  The orifice coefficient $C_o$
and the weep constant $K_2$ are charts in the literature, not formulae.
Choupo does not invent charts.  $C_o$ is a declared input, printed back
at you with its provenance.  $K_2$ is optional: give it and the weep
check runs, leave it out and the column tells you plainly that the
check was not performed.  If a tray lands where the flooding fit is
known to be unreliable ($F_{LV} < 0.03$), the column says that too.

\paragraph{What it prints.}  Run
\file{tutorials/steady/distillation/column09\_tray\_hydraulics} at
\code{verbosity 3}:

\begin{lstlisting}
--- Sieve-tray hydraulics (DESIGN: the diameter is the result) ---
D = 1.322 m    tray spacing 0.500 m    weir 50.000 mm    holes 5.000 mm ...
C_o = 0.840  (declared -- the Liebson chart value for this A_h/A_p and t/d_h)
sized so no tray passes 80.000 % of its flooding velocity

stage    F_LV    C_SB    u_v    u_flood   %flood    h_t     dP      h_b   h_b,max
 1     0.0389  0.091   1.002   1.575    63.6     112.9   886.1   178.1   275.0
 ...
14     0.0826  0.085   1.081   1.351    80.0     135.3  1017.8   215.8   275.0
worst tray: 80.0 % of flood on stage 14   column dP = 13.26 kPa   ...
\end{lstlisting}

The last stripping tray sets the diameter, sitting exactly on the
$80\,\%$ it was asked for.  Notice that the flow parameter $F_{LV}$
jumps at the feed --- more liquid, less vapour below it --- which lowers
the capacity parameter $C_{SB}$ and so the flooding velocity.  Which
tray ends up tightest is not obvious in advance; you find out by
looking.

\begin{tryit}
Run \file{column10\_flooding}.  It is the \emph{same} column, built
$1.10$~m wide instead of $1.322$~m.  Compare the stream tables: they
are identical to the last digit, the MESH converges just as fast, the
distillate is still $98.12\,\%$ benzene.  Then read the hydraulics:
stage~14 runs at $115.6\,\%$ of flood, and the downcomer backs up to
$271$~mm against a $275$~mm limit.  On paper it is a perfectly good
column.  It would not work.  This is exactly why the check exists.
\end{tryit}

\begin{pitfall}
Hydraulics is a \emph{rating pass on the converged profile}, not a
coupled unknown.  The $13$~kPa of pressure drop it reports is
\textbf{not} fed back into the stage temperatures --- the column was
solved at one pressure.  For a low-pressure column where the drop is a
large fraction of the operating pressure, that matters, and Choupo says
so rather than quietly absorbing it.  The KPIs \code{diameter},
\code{floodApproach\_max}, \code{floodStage}, \code{dP\_column\_kPa}
and \code{downcomerBackup\_max\_mm} are available to a sweep or an
optimiser, so you can trace how far you can push the reflux before a
tray floods.
\end{pitfall}

\subsubsection{Shortcut column, absorber, stripper, extractor}

\code{shortcutColumn} (alias \code{FUG}) is the
Fenske--Underwood--Gilliland \emph{design} method (constant $\alpha$,
no azeotropes): declare the key components, their recoveries and
\code{refluxFactor} ($R/R_{\min}$); $N_{\min}$, $R_{\min}$, $N$, and
the feed stage are results.  Use it to size first, then hand the
numbers to \code{distillationColumn} for a rigorous rating.

\code{absorber} / \code{stripper} are counter-current staged
gas--liquid contactors solved by the Thomas tridiagonal; the solute(s)
need a Henry pair (\S\ref{sec:thermo}).  \code{extractor} (alias
\code{extract}) is the liquid--liquid analogue: each theoretical stage
is one LL equilibrium computed by the \emph{same} Gibbs-minimisation
flash kernel the flash uses (never re-implemented), the cascade swept
by successive substitution.  Hardware in, results out: \code{stages}
is the knob; recovery and the distribution ratio are results.
Tutorials: \file{absorber01\_NH3\_water},
\file{extract01\_ethanol\_water\_benzene}.

% ---------------------------------------------------------------------
\subsection{Membranes and electrochemical separations}\label{sec:membranes}

\paragraph{\code{spiralWoundModule} (alias \code{membraneSW}).}  The
NF/RO element: solution-diffusion permeation + film-model concentration
polarisation, with selectable sub-models for feed-channel mass transfer
(\code{SchockMiquel} or \code{constant}), pressure drop, and osmotic
pressure (\code{vanHoff} or \code{Pitzer}):

\begin{lstlisting}[language=dict]
operation
{
    membrane         SW30HR;        // -> data/standards/membranes/<name>.dat
    area             35 m2;         // per element
    length           1 m;
    elements         1;             // >1 for a train in one vessel
    P_perm           1.01325 bar;
    dP_feed_total    2 bar;
    massTransfer  { model SchockMiquel;  channelHeight 0.7 mm;
                    diffusivity 1.6e-9 m2/s;  viscosity 1.0e-3 Pa.s; }
    osmotic       { model Pitzer; }
}
\end{lstlisting}

Inputs \code{( feed )}, outputs \code{( retentate permeate )}.  KPIs:
\code{J\_w\_avg}, per-solute observed rejection, \code{water\_recovery},
\code{dP\_feed}.  Four membranes ship in the catalogue (a seawater-RO
archetype, a loose-NF archetype, a pore-model NF, and an ion-exchange
pair); the solutes carry \code{nonvolatile true;} and a van~'t~Hoff
\code{dissociation} factor.  A \code{scaling \{ minerals (calcite
gypsum); \}} block runs the aqueous speciation solver per module at the
bulk \emph{and at the wall}, so you see concentration polarisation as
the scaling driver (\file{membrane07\_scaling\_si}).  Start with
\file{membrane01\_RO\_NaCl\_seawater}, then the concentration-polarisation
sweep \file{membrane03\_concentration\_polarization}.

\paragraph{\code{ionExchanger}.}  A cation-exchange softener: the inlet
water analysis is fully equilibrated with a fixed-capacity Na-form
resin (Gaines--Thomas mass action), so Ca/Mg drop onto the resin and Na
rises equivalent-for-equivalent.  \code{operation \{ resin SAC\_Na;
CEC 2.0 eq/L; \}}; hardness removal is a \emph{result} (a
target-hardness key is refused --- that is an outer-driver job).  Every
run carries a non-suppressible banner: this is the fresh-resin
equilibrium limit, not a bed in service (breakthrough is transient).
Tutorial: \file{membrane08\_softened\_scaling} --- softener upstream of
the RO train, the scaling driver removed at the source.

\paragraph{\code{electrodialysisStack}.}  N cell pairs of alternating
cation/anion-exchange membranes: an applied current drives ions out of
the diluate channel into the concentrate.  Glass-box electrochemistry,
all printed: Faraday transfer per cell pair ($\dot n_i = \xi I / z_i F$),
Nernst membrane potential from real ion activities, the Cowan--Brown
limiting current density (a loud warning --- never a clamp --- when
$i > i_{\lim}$), ohmic drops from the membrane and channel resistances,
and the stack power published on an energy wire.

\begin{lstlisting}[language=dict]
{ name edStack;  type electrodialysisStack;
  inputs ( diluateFeed concentrateFeed );  outputs ( diluate concentrate );
  operation
  {
      N_cellpairs 100;   current 8.0;        // A
      xi 0.9;                                 // current efficiency (announced)
      membraneArea 0.2 m2;  channelThickness 0.5 mm;
      channelLength 0.4 m;  linearVelocity 0.05 m/s;
      E_electrodes 1.5;  membrane CMX_AMX;
  }
  energyOutputs ( { name power; kind work; expression "W_electric"; unit W; } ) }
\end{lstlisting}

Tutorials: \file{tutorials/electrochem/ed01\_nacl\_desalination} and
\file{ed02\_over\_limiting\_current} (what the $i_{\lim}$ warning looks
like when you cross it).

\paragraph{\code{psa}.}  Pressure-swing adsorption as a steady
flowsheet unit --- the equilibrium-theory (local-equilibrium) shortcut:
the swing between $P_{\mathrm{high}}$ and $P_{\mathrm{low}}$ on a
competitive-Langmuir adsorbent sets the captured flow per species, with
a declared equilibrium-utilisation derating \code{eta} and a purge
fraction.  What the shortcut does \emph{not} model (breakthrough, cycle
time, thermal effects) is announced in the run header.  A
result-named key (\code{recovery}, \code{purity}) in \code{operation}
is refused --- hitting a target is an outer-driver job.  Tutorial:
\file{tutorials/steady/separation/psa01\_h2\_psa}.

% ---------------------------------------------------------------------
\subsection{Solids: separation, conveying, crystallisation}\label{sec:solids}

Solids ride streams as a \code{phases \{ solid \{...\} \}} decomposition
plus an optional particle-size distribution (\code{psd}); the solids
units read and transform both.

\paragraph{Gas--solid separation.}  \code{cyclone} is a PSD-aware
classifier with five selectable models (\code{Lapple} default,
\code{LeithLicht}, \code{IoziaLeith}, \code{Barth},
\code{Muschelknautz} --- the last adds the solids-loading effect);
\code{bagFilter} is cake-dominated near-total collection whose headline
KPI is \code{dP\_filter}; \code{gasSolidSplitter} is the ideal spec'd
split (PSD passes unchanged --- it does \emph{not} classify).

\paragraph{\code{pneumaticConveyor}.}  A dilute-phase conveying line:
the total $\Delta P$ is the sum of announced contributions (gas +
solids acceleration, wall friction, static head), particle velocity by
Hinkle, and the \emph{Rizk saltation velocity} reported with a loud
warning if the gas velocity falls below it (settling / blockage risk)
--- never a silent clip.

\paragraph{\code{crystalliser}.}  Three models on one saturation
foundation:

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
\item \code{equilibrium} (default) --- cooling crystalliser; the mother
      liquor leaves saturated, yield $=$ solute in $-\,c_{\mathrm{sat}}(T)\cdot$solvent.
\item \code{MSMPR} --- steady continuous population balance by the
      method of moments: nucleation/growth kinetics from
      \file{constant/crystallisation}, supersaturation $S$ a result,
      and the full crystal-size distribution (mean sizes, CV, magma
      PSD) reported.
\item \code{FVM} --- the discretised population balance on a size grid
      (size-dependent growth).
\end{itemize}

All three read the same saturation --- a solubility curve
(\code{solubility \{\}} on the component) or, for a salt, the
ion-activity $K_{sp}$ of the electrolyte package (including antisolvent
drowning-out; \file{crystalliser08\_nacl\_antisolvent\_msmpr}).
Multiple feeds are first-class (\code{inputs ( brine antisolvent );})
and \code{operation \{ solute <name>; \}} names which salt this unit
crystallises when the feed carries several
(\file{crystalliser09\_KHT\_KCl\_series}).  The declared
\code{outputs} pick the product shape: \code{( magma )} for one slurry
(required by MSMPR/FVM so the PSD travels), or
\code{( crystals motherLiquor )} for a solid--liquid split, where the
cake leaves \emph{wet} (\code{cakeMoisture}, default 0.10 kg liquor per
kg dry solid --- perfectly dry solids are unphysical) and
\code{solidsRecovery} lets fines escape with the liquor.

% ---------------------------------------------------------------------
\subsection{Drying}\label{sec:drying}

Three dryers, one honesty rule --- the drying air is a \emph{real
stream} (it brings the heat and carries the moisture; the outlet
temperature is a result, there is no imposed duty):

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
\item \code{sprayDryer} --- solution/suspension $\to$ powder
      (\S\ref{sec:advanced-spray});
\item \code{solidDryer} --- polish a \emph{hygroscopic} powder toward
      the equilibrium moisture set by the air humidity and the solid's
      GAB sorption isotherm (a \code{sorption \{\}} block on the
      component, typically a case-local overlay --- sorption is
      sample-specific);
\item \code{evaporativeDryer} --- dry a \emph{non-sorbing} crystalline
      solid of its free surface moisture: no isotherm exists or is
      consulted; constant-rate evaporation until the first of three
      announced limits binds (water gone, exhaust saturates, air runs
      out of heat).
\end{itemize}

\subsubsection{Advanced spray drying: the atomiser library}\label{sec:advanced-spray}

\begin{whatyoulllearn}
How the \code{sprayDryer} atomises a solution into droplets, dries them
to a powder of a real particle size, and where each piece of physics
--- the droplet size, the liquid properties, the drying curve --- comes
from.  ``All models are wrong, but some are useful'': every correlation
is an algebraic fit (no CFD) and announces its source and validity
envelope at run time.
\end{whatyoulllearn}

A spray dryer takes a liquid feed and hot drying air (both flowsheet
\emph{streams}) and returns a powder plus the humid exhaust.  Three
pieces of physics compose: (i)~\emph{atomisation} sets the droplet
Sauter diameter $d_{32}$; (ii)~\emph{single-droplet drying}
(Ranz--Marshall
$\mathrm{Nu}=2+0.6\,\mathrm{Re}^{1/2}\mathrm{Pr}^{1/3}$, a wet-bulb
surface) sets the rate; (iii)~the droplet shrinks to a powder particle
$d_p=d_{32}\,(x_{\text{solid}}\rho_L/\rho_{\text{solid}})^{1/3}$
carrying a residual moisture set by the material's sorption isotherm.

\paragraph{The atomiser is a selectable sub-model.}

\begin{lstlisting}[language=dict]
operation
{
    atomiser
    {
        model   pressureNozzle;     // rotary | pressureNozzle | twinFluid
        deltaP  40 bar;             // pressure-nozzle hardware
    }
    chamberDiameter  1.0 m;         // -> residence time is a RESULT
    chamberHeight    5.0 m;
    flow             co;            // co | counter current
}
\end{lstlisting}

\noindent Omit the \code{atomiser} block and the default \code{rotary}
wheel reads the legacy \code{wheelSpeed}/\code{wheelDiameter} keys ---
older cases are unchanged.  The families and their hardware:

\begin{center}\small
\begin{tabular}{@{}lll@{}}
\toprule
\code{model} & correlation ($d_{32}$) & hardware \\
\midrule
\code{rotary} & Friedman/Marshall wheel & \code{wheelSpeed} [rpm], \code{wheelDiameter} [m] \\
\code{pressureNozzle} & $9575\,\Delta P^{-1/3}\,\mu$m (class curve) & \code{deltaP} [Pa] \\
              & \emph{or} Wang--Lefebvre (two-term) & \code{correlation wangLefebvre} \\
\code{twinFluid} & Nukiyama--Tanasawa airblast & \code{relativeVelocity}, \code{airLiquidMassRatio} \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Droplet size follows the liquid properties.}  $d_{32}$ rises
with surface tension ($\sigma^{0.1}$--$\sigma^{0.6}$) and viscosity
($\mu_L^{0.2}$), so a concentrated, viscous feed sprays coarser.
Choupo pulls $\sigma$ and $\mu_L$ from the property layer when the
\file{propertyDict} declares the models (\code{surfaceTension},
\code{liquidViscosity}), with an announced fallback otherwise --- the
$\sigma$ source is printed in the report.

\paragraph{Drying kinetics --- equilibrium vs.\ curve.}  The powder
does not leave bone dry: it stops at the residual moisture set by
(a)~the equilibrium $X_e$ from the solid's \code{sorption} isotherm at
the exhaust-air humidity, and (b)~the falling-rate approach past the
critical moisture $X_c$ (the \code{dryingCurve} in
\file{constant/dryingKinetics}).  Co- vs.\ counter-current changes the
residence time and, in the trajectory models, the droplet thermal
history.

\paragraph{Chamber model --- named by its authors, not ``rigorous''.}
The unit-level \code{model} slot picks the chamber model:
\code{marshall} (default) --- the lumped inlet/outlet mass + energy
balance and single-droplet constant-rate time (Marshall 1954);
\code{langrishKockel} --- the 1-D axial droplet trajectory with the
Characteristic Drying Curve, which reports the profiles of $X$, $d$,
$T_g$, $T_p$ \emph{along} the chamber (a \file{profile.csv} you can
plot); or \code{chen} --- the same trajectory with the Reaction
Engineering Approach kinetics, a smooth activation-energy rate with no
critical-moisture break.  Reported KPIs: $d_{32}$, $d_p$, Nu, Sh, $h$,
$k_c$, $T_{\mathrm{out}}$, $T_{\mathrm{wetbulb}}$, the constant-rate
droplet time and $X_{\mathrm{final}}$ --- all feeding
sensitivity/sizing/costing.

\subsubsection{Simulation vs.\ design --- the model runs both ways}\label{sec:spray-design}

It can feel odd to supply a wheel speed and a chamber size just to
\emph{simulate} a dryer.  That feeling is correct, and it has a name:
what you are doing is \emph{rating} ---
geometry~$\rightarrow$~performance, ``I have this dryer, what does it
do?''.  The opposite question --- ``I want this powder, what dryer do I
need?'' --- is \emph{design}, targets~$\rightarrow$~geometry.  Choupo
does not split these into two black boxes; it keeps \textbf{one model
and runs it in both directions}.

\emph{Rating} is the \code{sprayDryer} unit itself.  \emph{Design} is a
\file{postDict} sizing pass carrying a \code{design \{\}} block, which
\textbf{inverts the run you just rated} --- it does not re-derive the
correlations, it \emph{scales from the simulated point}, so every step
is visible:

\begin{lstlisting}[language=dict]
sizing
{
    units
    (
        {
            unitName    dryer;
            type        sprayDryer;
            material    SS304;
            designRules
            {
                design
                {
                    targetSize       40;    // um  -- target powder Sauter mean d32
                    residenceFactor  3.0;   // chamber residence >= 3 x drying time
                }
            }
        }
    );
}
\end{lstlisting}

The pass then reports the geometry the targets demand: the wheel speed
from the target drop size (Friedman's $d_{32}\!\sim\!N^{-0.6}$) and the
chamber size from the required residence, with the Tanasawa regime
warning still telling you whether that design lands in the clean
ligament regime.  In \file{tutorials/steady/drying/sprayDryer07\_design}
the sugar dryer rated at 20\,000~rpm must spin to $\approx$40\,000~rpm
for a 40~\textmu m powder, and its $1\times5$~m chamber turns out
$\sim$30$\times$ oversized --- the fast sugar drying needs only 0.24~s
of residence.  \textbf{You simulate with the unit and design with the
pass: two directions of one model, both on the page, neither hidden.}

% ---------------------------------------------------------------------
\subsection{Flow primitives: mixer, splitter, valve, pipe}\label{sec:primitives}

\paragraph{\code{mixer}.}  Adiabatic merge:
$P_{\mathrm{out}} = \min(P_{\mathrm{in}})$, energy balance on the
dominant phase, solids summed, Newton-1D in $T_{\mathrm{out}}$.  A
mixer \emph{mixes} --- it never runs an internal flash (separators
separate).  Declare gas inlets with \code{phase gas;} in their
\file{0/} files so the energy balance picks the right basis; an
optional \code{operation \{ T 330 K; \}} declares the mix temperature
and skips the adiabatic balance (announced in the log).

\paragraph{\code{splitter}.}  One inlet, N outlets at
\code{fractions ( 0.6 0.4 );} --- a flow split, not a separation: all
branches share $T$, $P$, composition, $vf$.

\paragraph{\code{valve}.}  Isenthalpic throttle to
\code{operation \{ P 1 bar; \}}; $T_{\mathrm{out}}$ and the vapour
fraction are results (a real-gas EoS flashes the let-down).  Use it
upstream of a flash to set the drum pressure; it warns if
$P \ge P_{\mathrm{in}}$.

\paragraph{\code{pipe}.}  The pressure \emph{drop} is a computed result
of geometry + flow --- never a spec (a pipe has no knob; a pump raises
$P$, a valve sets a let-down).  Darcy--Weisbach friction + minor losses
+ static head, with the friction-factor sub-model in the \code{model}
slot (\code{Churchill} default --- one explicit correlation across all
regimes; \code{Haaland}; \code{Colebrook}):

\begin{lstlisting}[language=dict]
{ name P1;  type pipe;  model Churchill;
  in feed;  outputs ( downstream );
  operation {
    geometry {
      D 0.1 m;  L 50 m;  roughness 4.6e-5 m;  dz 2 m;    // rise costs P
      fittings ( { K 0.9; count 2; }      // two elbows
                 { K 10.0;         } );   // one globe valve
    }
  }
}
\end{lstlisting}

When the feed flashes two-phase at $(T,P)$, the pipe automatically
switches to a gas--liquid model, selected in an \code{operation \{
twoPhase \{ model ...; \} \}} sub-block: \code{LockhartMartinelli}
(default; Chisholm multiplier + Butterworth holdup),
\code{homogeneous}, \code{Friedel}, or \code{BeggsBrill} (flow-pattern
map + inclination).  KPIs include the five-way $\Delta P$ breakdown,
velocities, Reynolds, regime, holdup and void fraction.  Tutorials:
\file{pipe01\_water\_line}, \file{pipe02\_airwater\_twophase}.

% ---------------------------------------------------------------------
\subsection{Heat duties, utilities, heat integration}\label{sec:energy-utilities}

One rule covers heat: \textbf{a duty is carried by a physical stream
when one is wired to it, otherwise it is sized to a plant utility by
temperature level}.  Shaft work is a direct scalar wire
(\S\ref{sec:rotating}); no abstract ``Q stream'' ever travels the
flowsheet.

\paragraph{Q then allocate (the default).}  Most units just report
their duty and let a report size the utility.  Add to
\file{controlDict}:

\begin{lstlisting}[language=dict]
reports { utilityAllocation { dTmin 10; } }
\end{lstlisting}

For each duty $Q$ at process temperature $T$ the report picks the
lowest-grade catalogue utility that still works (a steam grade with
$T_{\mathrm{in}} \ge T + \Delta T_{\min}$ for heating; a coolant with
$T_{\mathrm{in}} \le T - \Delta T_{\min}$ for cooling) and converts
$|Q|$ to a mass flow, a load, and a cost.  Nine utilities ship in
\file{data/standards/utilities/}: \code{steamLP/MP/HP},
\code{coolingWater}, \code{chilledWater}, \code{dowthermA},
\code{hitecSalt}, \code{refrigerationPG}, and \code{electricity} ---
the power tier tallies pump/compressor motors against turbine
generators, so a combined cycle shows a net electricity credit.

\paragraph{Heat integration.}  Reuse a column's condenser heat by
wiring the duty on the consumer, like any energy link:

\begin{lstlisting}[language=dict]
{ name preheater;  type heater;  in coldFeed;  outputs ( warmFeed );
  operation { }                                   // Q comes from the link
  energyInputs ( { from column01.condenser;  kind heat;  target Q; } ) }
\end{lstlisting}

Both ends then read \emph{(carried)} in the allocation --- no
double-counting, and the report shows the saving.  A \emph{forward}
link runs in one pass; a \emph{feedback} link (the condenser preheating
the column's own feed) is an energy tear --- the flowsheet detects it,
turns the recycle loop on, and converges the duty by successive
substitution.  Tutorials: \file{heatlink01\_condenser\_to\_heater},
\file{heatlink02\_condenser\_feed\_preheat}.

% =====================================================================
\section{Reports}\label{sec:reports}

\begin{whatyoulllearn}
One \code{reports \{\}} block in \file{controlDict} turns a converged
run into CSV tables and a consolidated spreadsheet --- stream tables,
balances, utility bills, profiles, costs.
\end{whatyoulllearn}

Each key names a report object; after a converged solve each writes
into the case's \file{reports/} directory and announces the file:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Report} & \textbf{What it writes} \\
\midrule
\code{streamTable}   & one row per stream: role, $F$, $F_{\mathrm{mass}}$, $T$, $P$, $vf$, compositions \\
\code{massBalance}   & per-component in/out/net + global closure \%, and a per-unit table \\
\code{energyBalance} & per-unit $H_{\mathrm{in}}$ / $H_{\mathrm{out}}$ / $\Delta H$ (the implied duty) + the plant boundary \\
\code{utilities}     & consumption aggregated by stream \code{category} (kg/h, MW, cost) \\
\code{utilityAllocation} & each duty sized to a catalogue utility by temperature level (\S\ref{sec:energy-utilities}) \\
\code{energyStreams} & the W/Q wire ledger \\
\code{computed}      & the values of computed \code{\$variables} (below) \\
\code{profiles}      & per-unit internal profiles (column stages, PFR axis, dryer trajectory) \\
\code{design} / \code{economics} & equipment realisation + cost tables (with a \file{postDict}, \S\ref{sec:post}) \\
\code{spreadsheet}   & one consolidated, coloured \file{report.ods} (multi-sheet) \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{Computed variables.}  For post-processing arithmetic,
declare \code{compute} expressions in the \code{variables} block of
\file{flowsheetDict} --- they resolve against unit KPIs, stream fields,
and earlier variables, and land in \file{reports/computed/values.csv}:

\begin{lstlisting}[language=dict]
variables
{
    Q_total    { compute "effect1.Q + effect2.Q + effect3.Q";  unit kW; }
    economy    { compute "water_evap / effect1.F_steam"; }
}
\end{lstlisting}

% =====================================================================
\section{Sweeps, design specs, optimisation}\label{sec:outer}

\begin{whatyoulllearn}
How to wrap the simulator in an outer loop without writing a single
line of \cpp: trend one knob (\code{sweep}), map two
(\code{gridSweep}), meet N specs with N knobs (\code{designSpec}), or
search for the best value (\code{optimization}).
\end{whatyoulllearn}

Drop a \file{system/outerDict} into a case and the binary switches from
one pass to the driver loop automatically.  One driver per case;
\code{type} selects it.  Ask the right question of the right tool:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}Xl@{}}
\toprule
\textbf{Question} & \textbf{Tool} \\
\midrule
``What if $T$ were 360 K?'' (one point)          & edit the dict, re-run \\
``How does $V/F$ trend with $T$?''               & \code{sweep} \\
``What does the $(T,P)$ envelope look like?''    & \code{gridSweep} \\
``What area gives exactly 4500 kg/h?''           & \code{designSpec} \\
``Which reflux minimises $V_{\mathrm{strip}}$?'' & \code{optimization} \\
``Fit NRTL to my bubble-$T$ data''               & \code{fitParameters} under \code{choupoProps} (\S\ref{sec:props}) \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{Paths.}  Drivers mutate the case between passes via
dot-and-bracket paths (\code{units[0].operation.refluxRatio},
\code{variables.A}, \code{streams.feed.F} --- the last steered through
an explicit stream-override channel, since stream state lives in
\file{0/}).  Values written this way are \emph{raw canonical SI}.
Results are read back as \code{<unit>.<kpi>} or
\code{<stream>.\{F,T,P,vf\}}.

\subsection{\code{type sweep;} --- trend over one parameter}

\begin{lstlisting}[language=dict]
type        sweep;

parameter
{
    target    units[0].operation.T;        // any scalar dict path
    range     ( 360.0  385.0 );            // (min max), raw SI of the target
    nPoints   26;
}

responses
(
    flash01.V_over_F       // <unit>.<kpi>
    flash01.Q_kW
);

report { format csv;  file sweep_flashT.csv; }
\end{lstlisting}

One row per point; a point that fails to converge is reported and
written as \code{nan} --- the sweep continues.  Canonical case:
\file{tutorials/steady/flash/flash06\_sweep\_T} (the $V/F$ S-curve
across the bubble--dew window).  \code{type gridSweep;} is the 2-D
version --- a \code{parameters ( \{...\} \{...\} )} list of two targets,
producing the long-form CSV a contour plot reads
(\file{flash07\_gridsweep\_TP}: the flash envelope over $(T,P)$;
non-converged cells stay honest gaps, never interpolated).

A sweep may declare CONSTRAINTS --- the same block the SQP optimizer
uses, same grammar, same refusals:
\begin{lstlisting}[language=dict]
constraints
(
    { kpi column01.x_D_LK;   atLeast 0.985; }
    { kpi column01.V_strip;  atMost  0.065;  }
);
\end{lstlisting}
The sweep becomes a FEASIBILITY MAP: every row gains, per constraint,
the value, the bound, the signed normalised residual and a
\code{satisfied} flag, plus a final \code{feasible} verdict --- and no
point is ever filtered.  The infeasible rows are the lesson: where the
spec fails and where the budget binds, with the feasible window read
directly off the CSV
(\file{sensitivity02\_feasibility\_map}: the reflux window between the
purity spec and the boilup budget).

Two objectives that pull against each other build a PARETO FRONT by
epsilon-constraint --- \code{type paretoSweep;} --- with no
multi-objective solver: the second objective becomes a swept bound and
each front point is one full, auditable SQP run,
\begin{lstlisting}[language=dict]
type paretoSweep;   method sqp;
variables ( { path units[0].operation.refluxRatio;
              min 1.5; max 6.0; initial 3.0; } );
objective { kind kpi;  path column01.V_strip;  sense minimise; }
tradeoff  { kpi column01.x_D_LK;  bound atLeast;
            range ( 0.96  0.994 );  nPoints 5; }
report    { file pareto.csv; }
\end{lstlisting}
The CSV records, per point, the bound, the achieved objective and
trade-off values, the variables at the optimum and the solver's own
converged/feasible verdicts --- a failed point stays on the table,
marked.  Where the bound is slack the point is the unconstrained
optimum; where it binds, the SQP's Lagrange multiplier is the marginal
price of the second objective
(\file{pareto01\_purity\_energy}: the purity--energy front of the
benzene/toluene column, $R^*$ rising 1.6\,$\to$\,3.8 as the purity
bound tightens).

An INTEGER decision --- how many stages? --- is declared, never
relaxed: \code{integer true;} on a variable makes the optimizer
enumerate every value in \code{min..max}, each one a NEW run (a changed
\code{nStages} rebuilds the column; a late parameter mutation on a
sized object would lie), solving the inner continuous problem at each.
The FULL table is emitted --- failures included, best marked
(\file{optim06\_integer\_stages}: the classic $N$-vs-$R$ trade, boilup
falling 1.03\,$\to$\,0.045\,kmol/s from 9 to 14 stages at the same
purity spec, with $N=8$ refused aloud because the feed cannot sit on
the bottom stage).

\begin{tryit}
\code{runCase tutorials/steady/optimisation/sensitivity01\_column\_reflux}
produces \file{sweep\_results.csv}.  Plot $V_{\mathrm{strip}}$ against
$R$ with any tool; you'll see the classic energy asymptote as the
reflux ratio grows past the minimum.
\end{tryit}

\subsection{\code{type designSpec;} --- meet N specs with N knobs}\label{sec:designspec}

The DesignSpec is the visible version of ``back-compute the input from
the target''.  Each knob is a case-level \code{\$variable}, declared in
\file{flowsheetDict} and referenced by the unit --- the driver iterates
the variable, a square Newton meets the targets:

\begin{lstlisting}[language=dict]
// system/flowsheetDict
variables { A 100 m2; }                       // initial guess; DesignSpec iterates it
units ( { name effect1; type evaporator;  ...
          operation { area $A; ... } }
        { name effect2; type evaporator;  ...
          operation { area $A; ... } } );     // the SAME $A: equal areas by construction
\end{lstlisting}

\begin{lstlisting}[language=dict]
// system/outerDict
type         designSpec;

manipulate
(
    { variable A;   initial 100.0 m2;   min 20.0 m2;   max 500.0 m2; }
    { variable streams.utilitySteam_200kPa.F;
                    initial 400 kmol/h; min 100 kmol/h; max 1500 kmol/h; }
);

targets                        // SAME COUNT as knobs (square Newton)
(
    { path  L3.F_mass;   value 4500.0 kg/h;   tol 1.0 kg/h; }   // value form
    { path  effect3.P;   value 15 kPa;        tol 50 Pa;    }
    // equality form: { lhs effect1.area;  rhs effect2.area;  tol 0.001; }
);

options { maxIter 40;  tolF 1.0;  fdStep 1.0e-4; }
report  { file designspec_history.csv; }
\end{lstlisting}

All knob entries carry units, converted on read; a \code{manipulate}
variable may also be a stream field (as above --- the steam flow the
plant must supply).  After convergence the simulator is replayed once
at the design point so the case ends with full converged output.
Canonical cases: \file{compressor02\_designspec\_pout} ($1\times1$),
\file{designSpec01\_triple\_equal\_areas} ($2\times2$: one area + one
steam flow meet a capacity and a vacuum spec at once).

\subsection{\code{type optimization;} --- search for the best value}

\begin{lstlisting}[language=dict]
type    optimization;
method  nelderMead;                        // or sqp (adds nonlinear constraints)

variables
(
    { path units[0].operation.refluxRatio;  min 3.0;  max 5.0;  initial 4.0; }
    { path units[0].operation.feedStage;    min 6;    max 12;   initial 8;   }
);

objective
{
    kind     kpi;             // kpi | stream | cost | costTotal
    path     column01.V_strip;
    sense    minimise;        // minimise (default) | maximise
}

options { maxIter 80;  tolX 5e-3;  tolF 1e-4; }
report  { file optimization_history.csv; }
\end{lstlisting}

\code{nelderMead} is the derivative-free simplex with hard box bounds
--- an evaluation that fails to converge gets an announced infinite
penalty and the simplex contracts away from it.  The \code{cost} /
\code{costTotal} objective kinds run the \file{postDict} sizing +
costing chain on every evaluation (\S\ref{sec:post}) --- see
\file{optim01\_column\_reflux} and \file{optim02\_process\_cost}.

\code{method sqp;} switches to a hand-rolled line-search Sequential
Quadratic Programming (damped BFGS on the Lagrangian, L1 merit,
active-set QP) and accepts a \code{constraints (...)} list with
unit-aware \code{atLeast} / \code{atMost} / \code{equals} bounds on any
KPI.  The trace is glass-box: per-iteration KKT residuals, the active
set, the Lagrange multipliers (shadow prices), and a closing five-line
KKT certificate.  Worked example:
\file{optim04\_membrane\_recovery\_scaling} --- brackish-RO recovery
maximised subject to wall-scaling constraints, with the calcite
constraint active at the optimum and its multiplier printed (the
shadow price of scaling head-room).  Asking Nelder--Mead for
constraints is a loud error --- add a penalty term or switch method.

\paragraph{Parameter estimation.}  Regression of model parameters
against experimental data lives in \code{choupoProps} as the
\code{fitParameters} operation (\S\ref{sec:props}); the old
\code{fitBinaryPair} outer driver was retired, and asking for it points
you to the replacement.

% =====================================================================
\section{Equipment sizing and costing}\label{sec:post}

\begin{whatyoulllearn}
How a single \file{postDict} turns a converged flowsheet into a vessel
list, wall thicknesses, weights, and bare-module costs.
\end{whatyoulllearn}

Drop a \file{system/postDict} and the simulator result is
post-processed after the pass:

\begin{lstlisting}[language=dict]
sizing
{
    units
    (
        {
            unitName    reactor;
            type        stirredTank;
            material    SS316;
            designRules
            {
                L_over_D        2.5;
                pressureDesign  3.0;
                corrosionAllow  0.003;        // m
                jointEfficiency 1.0;
            }
        }
        {
            unitName    heater;
            type        shellTubeHX;
            material    SS304;
            designRules { U 600.0;  LMTD 10.0;  pressureDesign 2.0; }
        }
    );
}

costing
{
    method      Guthrie;
    year        2026;
    cepci       820;
    cepci2001   397;
    usdToEur    0.92;
}
\end{lstlisting}

\code{type} selects an \code{EquipmentSize} model
(\code{stirredTank}, \code{shellTubeHX}, and the \code{design \{\}}
inversions such as the spray dryer's, \S\ref{sec:spray-design});
materials come from \file{data/standards/materials/} (carbon steel,
SS304, SS316, aluminium --- each with density, Guthrie material factor,
yield stress, ratings).  The \code{Guthrie} costing model uses
Turton-style purchased-cost correlations, the ASME wall-thickness
pressure factor for vessels, then
$C_{BM} = C_p (B_1 + B_2 F_M F_P)$ and $C_{TM} = 1.18\,C_{BM}$,
printing a per-unit table:

\begin{lstlisting}
========================  Equipment Costing  =========================
  Method:  Guthrie    Year:  2026.0    CEPCI: 820.0

  unit      equipment       F_M   F_P    C_purchased   C_bare_mod    C_total_mod
  heater    shellTubeHX     2.34  1.09     365 249      2 146 753    2 533 169
  reactor   stirredTank     3.05  1.00       2 055         16 033       18 918
  TOTALS (EUR)                            367 304      2 162 786    2 552 087
\end{lstlisting}

Cost indices, currency rates and prices are \emph{case data} --- dated
and cited in the dict, never invented by the engine.  Run
\file{tutorials/steady/flowsheets/process02\_with\_design} for the full
chain, and \file{optim02\_process\_cost} for an optimiser driving it.

% =====================================================================
\section{Batch simulation: \code{choupoBatch}}\label{sec:batch}

\begin{whatyoulllearn}
How the time-dependent binary integrates closed vessels ---
\code{batchReactor}, \code{batchStill} (plain Rayleigh or a staged
rectifier), \code{batchCrystalliser} --- and how a \code{recipe} block
scripts a campaign of transfers and setpoint changes.
\end{whatyoulllearn}

A batch case is the same case shape with time controls in
\file{controlDict}:

\begin{lstlisting}[language=dict]
application    choupoBatch;
startTime      0;         // s
endTime        600;       // s
deltaT         1.0;       // s   (RK4 step)
writeInterval  30;        // s   (trajectory snapshot interval)
\end{lstlisting}

Each vessel carries an \code{initial \{\}} charge instead of an inlet
stream; the run writes \file{trajectory.csv} (one column per state
variable, one row per \code{writeInterval}).

\subsection{\code{batchReactor}}

\begin{lstlisting}[language=dict]
{ name reactor;  type batchReactor;
  initial
  {
      T 350 K;  P 1.013 bar;  V 0.001;  totalMoles 0.0185;
      molarComposition { ethanol 0.5;  aceticAcid 0.5; }
  }
  operation { mode isothermal;  T_setpoint 350 K; }   // or: mode adiabatic
  reaction  esterification; }                          // or: reactions ( A_to_B B_to_C );
\end{lstlisting}

\code{mode isothermal} holds $T$; \code{adiabatic} integrates the
energy balance alongside the species --- \file{batch02\_adiabatic} shows
a cold start, a slow induction, then thermal runaway from 345 to
482~K in a hundred seconds.  The kinetics vocabulary is shared with the
steady reactors: \code{Arrhenius}, the gas-kinetics forms
(\code{modifiedArrhenius}, \code{thirdBody}, \code{falloff} --- real
combustion mechanisms run here, see \file{ignition02\_h2air\_slack}),
and \code{LHHW} --- the \emph{same} \code{RateLaw} class
(\S\ref{sec:lhhw}), so a batch vessel and a CSTR can never disagree
about what an adsorption denominator means.

\paragraph{Stiff chemistry: adaptive time-stepping.}  The default is
the fixed-step RK4 loop.  Add \code{timeStepping adaptive;} to
\file{controlDict} and the stiff, error-controlled Rosenbrock23
integrator takes over --- the step starts low (\code{deltaT0}), grows
through quiet stretches, and shrinks through the runaway, with the step
history printed at verbosity 3.  Trajectory rows still land on the
clean \code{writeInterval} grid.  Compare
\file{batch06\_adaptive\_runaway} with its fixed-step twin, and
\file{ignition01\_h2o2} (real GRI-Mech H/O ignition, $\lambda \sim
10^{15}$) with \file{ignition01\_h2o2\_rk4}, which blows up to NaN at
the same \code{deltaT} --- the stiffness lesson on real chemistry.

\subsection{\code{batchStill} --- Rayleigh and the batch rectifier}

The default \code{model rayleigh;} is the classical open still:
constant boil-off \code{F\_vap}, the bubble point re-solved each step,
the pot trajectory following the Rayleigh integral
(\file{still01\_benzene\_toluene} reproduces the closed form to
$\sim$2\,\%; \file{still02\_ethanol\_water\_azeo} shows the pot moving
\emph{away} from the azeotrope --- a Rayleigh still cannot cross it from
below).  \code{dischargeTo receiver;} routes the condensed vapour into
a passive \code{batchAccumulator}, so pot + receiver conserve the
charge at all times.

\code{model rectifier;} adds the classical batch rectification:
\code{nStages} ideal stages above the pot and a total condenser, with
the reflux policy a first-class choice:

\begin{lstlisting}[language=dict]
{ name still;  type batchStill;  model rectifier;
  dischargeTo receiver;
  initial   { T 365 K;  P 1.013 bar;  totalMoles 1e-3;
              molarComposition { benzene 0.5;  toluene 0.5; } }
  operation { P 1.013 bar;  F_vap 1.5e-6;      // boilup V, kmol/s
              nStages 3;
              refluxPolicy constantReflux;  refluxRatio 3.0; } }
\end{lstlisting}

Under \code{constantReflux}, \code{F\_vap} is the internal boilup $V$;
the product is $D = V/(R+1)$ and the distillate composition
$x_D(t)$ is solved each instant from the quasi-steady stage cascade ---
it starts rich and drifts down, the constant-$R$ purity trade
(\file{still04\_rectifier\_benzene\_toluene}; with \code{nStages 0} and
$R = 0$ it reproduces plain Rayleigh to the last digit).

\code{refluxPolicy constantComposition;} holds a light-key purity
instead: \code{lightKey benzene;\ x\_D\_target 0.93;\ refluxMax 8;\
onRefluxLimit stopDistillation;} --- $R(t)$ is solved each step by a
bracketed search, climbing as the pot depletes; when the ceiling can no
longer reach the target the run \emph{announces} the limit and executes
the declared action (distillation stops, the event lands in the
trajectory and the KPIs).  Distinct refusals exist for a target below
the $R=0$ floor, a target unreachable at \code{refluxMax}, and an
azeotrope barrier --- computed from the VLE, not guessed
(\file{still05\_rectifier\_constant\_xD}).

\subsection{\code{batchCrystalliser}}

A supersaturated charge desupersaturates in time: RK4 on the packed
moments $(\mu_0,\mu_1,\mu_2,\mu_3,n_{\mathrm{solute}})$, against the
same kinetics library the steady MSMPR uses
(\file{constant/crystallisation}), so the crystal-size distribution
grows through the batch (\file{batch05\_crystalliser}).

\subsection{Recipes --- scripting a campaign}

A \code{recipe (...)} list in \file{flowsheetDict} schedules events by
time or by condition:

\begin{lstlisting}[language=dict]
recipe
(
    { time 400;  action transfer;      from reactor;  to still;  fraction 1.0; }
    { time 200;  action setParameter;  unit reactor;  key T_setpoint;  value 350.0; }
    { when { unit reactor;  quantity x_ethanol;  op lt;  value 0.30; }
      action transfer;  from reactor;  to still;  fraction 0.5; }
);
\end{lstlisting}

Actions: \code{transfer} (partial via \code{fraction}; the receiving
vessel may start empty), \code{setParameter}, and condition-triggered
\code{when} events (\code{quantity} is \code{T}, \code{total},
\code{n\_<comp>} or \code{x\_<comp>}; \code{op} is
\code{gt}/\code{ge}/\code{lt}/\code{le}).  Every fired action lands in
the result's timeline --- the machine-readable campaign sequence ---
and, underneath it, the MATERIAL LEDGER: one structured record per
material edge (interval, vessels, per-component amount, transported
enthalpy integrated package-by-package).  The run closes a campaign
MASS BALANCE over all-phase vessel inventories ($m_0 = m_{final} +
m_{external}$; per-ELEMENT for reacting campaigns) and reports it under
the \code{campaign} KPIs; a genuine leak marks the run non-converged.
A still without a \code{dischargeTo} receiver is recorded as an
unrouted external outlet, never a silent loss.
\file{recipe01\_react\_then\_distill} runs the classic react-then-still
sequence in one case.

The material ledger has an ENERGY twin: per vessel, one record per
segment of constant physics (a \code{setParameter} or a charge closes
the segment), with the isothermal reactor's jacket duty priced exactly
from committed mole numbers ($Q = \Delta H$ of the closed constant-$P$
segment; a declared \code{dH\_rxn} prices lumped species via reaction
extents, announced).  A recipe \code{transfer} conserves enthalpy by
construction: the mix temperature solves $H$-equality on the elements
datum (heat-capacity differences matter --- hot ethanol into cold water
lands $\sim$4\,K above the molar average), falling back to the molar
average only when a species cannot be priced, announced.  Stills carry
reboiler and condenser duty records (phase-enthalpy differences of the
actual packages at their own temperatures; the rectifier's condenser
handles the whole boilup $V$ at the top-stage $T$).  The campaign
ENERGY BALANCE only claims a verdict when every piece is ledgered and
priceable --- otherwise the run prints \code{energy balance UNAVAILABLE}
naming each missing piece (a temperature jump, a fallen-back charge, a
species without formation data).  KPIs land under \code{campaign} as
\code{energy\_*}; see \file{batch08\_isothermal\_duty},
\file{recipe04\_charge\_enthalpy} and
\file{recipe01\_react\_then\_distill} (the react-then-distil showcase
closes end to end at $10^{-16}$).  Valid duties are then allocated to
catalogue UTILITIES by temperature level (the still's reboiler draws LP
steam by its latent heat; a 293\,K crystalliser needs chilled water,
not cooling water), with campaign totals under \code{utility\_*} ---
energy, mass and cost per utility; a duty no utility can serve is
listed, never dropped.

% =====================================================================
\section{Dynamic simulation and control: \code{choupoCtrl}}\label{sec:ctrl}

\begin{whatyoulllearn}
How the control binary integrates a continuously fed plant while
controllers write manipulated variables --- and why the controller
sample time and the integrator step are two different things.
\end{whatyoulllearn}

\code{choupoCtrl} integrates open vessels with continuous feeds ---
$\mathrm{d}Y/\mathrm{d}t = f(Y, u, t)$ --- with a \code{controllers}
layer writing the manipulated variables $u$ at each sample.  In
\file{controlDict}, \code{deltaT} is the controller \emph{sample time};
with \code{timeStepping adaptive;} the plant is sub-stepped between
samples by the stiff integrator while the manipulated variable is held
--- two time steps, both correct
(\file{ctrl03\_adaptive\_disturbance}).

\subsection{\code{dynamicCSTR}}

\begin{lstlisting}[language=dict]
{ name reactor;  type dynamicCSTR;
  initial { T 320 K;  P 1.013 bar;  V 0.001;  totalMoles 0.0185;
            molarComposition { compA 1.0; } }
  feed    { F 5e-5 kmol/s;  T 320 K;  composition { compA 1.0; } }
  operation { reactions ( a_to_b );  jacketUA 30 W/K; } }
\end{lstlisting}

Constant-volume continuous CSTR with an optional jacket
($UA\,(T_j - T)$); RK4 on the packed $(n_i, T)$.  Its measured (CV)
and manipulated (MV) variables are string-keyed so controllers bind by
name.  Kinetics: \code{Arrhenius} or \code{LHHW} (the shared rate law
--- \file{ctrl07\_lhhw\_inhibition} closes a temperature loop on a
reaction that poisons itself, the exotherm fading as the product covers
the catalyst).  Add \code{start steadyState;} inside \code{initial}
and the engine relaxes the holdup ODE to its fixed point at the
declared feed and prints the seeded state --- a clean start at the
operating point.

\subsection{Controllers}

\begin{lstlisting}[language=dict]
controllers
(
    {
        name  TC1;
        type  PID;
        measurement { unit reactor;  cv T_reactor; }
        actuator    { unit reactor;  mv T_jacket; }
        setpoint    350 K;
        Kp  4.0;  Ki  0.04;  Kd  0.0;
        bias  320;  uLow  280;  uHigh  420;
    }
    {
        name  FeedDist;
        type  Signal;                                  // open-loop forcing
        actuator { unit reactor;  mv T_in; }
        signal   { type sinusoidal;  mean 320 K;  amplitude 15 K;  period 600 s; }
    }
);
\end{lstlisting}

Three controller types: \code{PID} (clamped anti-windup,
derivative-on-PV), \code{Schedule} (open-loop piecewise-constant steps
--- the disturbance injector of \file{ctrl02\_disturbance\_rejection}),
and \code{Signal} --- the full forcing-function vocabulary
(\code{step}, \code{staircase}, \code{ramp}, \code{pulse},
\code{sine}, \code{prbs}).  Inject a sinusoid on an inlet and watch the
loop's forced response, attenuated and phase-lagged; sweep the period to
build a Bode plot one point at a time (\file{ctrl06\_sine\_disturbance}).
For identification experiments, \code{type prbs;} generates the classic
broadband binary sequence from a declared-seed maximal-length LFSR ---
deterministic and exactly repeatable, with the period verified at load
and the leading bits printed (\file{ctrl08\_prbs\_ident}):
\code{signal \{ type prbs; mean 320 K; amplitude 10 K; bitPeriod 120;
registers 4; seed 1; \}}.

Disturbances can target the unit's INLET FACE by field name ---
\code{actuator \{ unit reactor; inletField T; \}} (or \code{F},
\code{moleFraction.<c>}, \code{componentMolarFlow.<c>}) --- the only
legal disturbance surface.  The inlet's authoritative state is the
per-component molar-flow vector: \code{componentMolarFlow.<c>} writes
one component's flow and re-derives $F$ and $z$ (the disturbance never
hides from the other components), while \code{moleFraction.<c>} holds
the total flow and renormalises the others proportionally --- two
explicit semantics, each announced at start-up, both validated hard
(\file{ctrl09\_stream\_disturbance} pins the derived pair to hand
arithmetic).  The trajectory carries \code{F\_in} and \code{z\_in\_*}
as first-class columns.
The trajectory CSV carries every CV, MV and setpoint --- plot it and
read the tuning off the page.

% =====================================================================
\section{The property workbench: \code{choupoProps}}\label{sec:props}

\begin{whatyoulllearn}
How to interrogate the thermodynamic foundation \emph{before} wiring a
flowsheet: single points, 1-D/2-D/ternary scans, data-consistency
tests, and parameter fits --- all from a \file{propsDict}.
\end{whatyoulllearn}

A props case replaces \file{flowsheetDict} with a
\file{system/propsDict} carrying \code{operations (...)}:

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{type} & \textbf{What it does} \\
\midrule
\code{propertyPoint}   & one state, properties to stdout \\
\code{propertyScan1D}  & sweep one variable, long-form CSV ($P^{\mathrm{sat}}(T)$, $\gamma(x)$,~\dots) \\
\code{propertyScan2D}  & a $(T,P)$ or composition grid ($Z$, residuals,~\dots) \\
\code{propertyScanTernary} & walk the ternary simplex at fixed $(T,P)$, classify the phase regime per node (the same VLLE flash kernel as the unit op), collect tie-lines \\
\code{fitParameters}   & Levenberg--Marquardt regression of any dotted package parameter against data, with identifiability statistics \\
\code{estimateComponent} & create a component from molecular groups (Joback + corresponding-states closures), shown against reference data \\
\code{vleConsistency}  & thermodynamic consistency tests on VLE data (Herington area + Gibbs--Duhem point test) \\
\bottomrule
\end{tabularx}
\end{center}

Two flavours to give the taste:

\begin{lstlisting}[language=dict]
{ name psat;  type propertyScan1D;
  vary { variable T;  from 250 K;  to 400 K;  n 51; }
  properties ( Psat_ethanol  Psat_water );
  output { file psat.csv; } }

{ name fit_nrtl;  type fitParameters;
  parameters
  (
      { path activityModel.pairs[0].a_ij;  initial -0.8;  min -15;  max 15; }
      { path activityModel.pairs[0].b_ij;  initial 246;   min -5000; max 5000; }
  );
  residual
  {
      kind  T_bubble;   P 1.01325 bar;
      data ( 0.05 361.95   0.10 359.95   0.20 358.05  ... );
  } }
\end{lstlisting}

The fit writes \file{fit\_history.csv} and \file{parity.csv} and
reports which parameters the data can actually identify --- the honest
part most fitting exercises skip.  A fitted pair is written as a dated
\emph{proposal} under the case's \file{constant/}; promoting it into
the frozen standard catalogue is a deliberate human act (the engine
refuses to write under \file{data/standards/}).

This is the tip of a much larger workbench --- component estimation
from groups, model-vs-data overlays, electrolyte activity models
(Pitzer, eNRTL), steam-table verification, speciation and scaling
indices.  The \emph{Properties Guide} (\file{docs/propsGuide.pdf}) and
the \emph{Explorer Guide} cover it in depth; start with
\file{tutorials/props/README.md} and the worked cases under
\file{tutorials/props/}.

% =====================================================================
\section{Tour of the tutorials}\label{sec:tutorials}

\begin{whatyoulllearn}
Choupo ships about 200 tutorials under \file{tutorials/}, in six
categories: \file{steady/}, \file{batch/}, \file{ctrl/}, \file{props/},
\file{plant/}, \file{electrochem/}.  \code{listCases} tabulates them
all with one-line descriptions.  The selection below is the on-ramp ---
pick the one closest to what you want to do and edit from there.
\end{whatyoulllearn}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Tutorial} & \textbf{What you learn} \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/flash, steady/thermoTest}} \\
\code{flash01\_benzene\_toluene}        & First run: Raoult VLE, Rachford--Rice, Newton trace, the \file{0/} state files. \\
\code{flash02\_ethanol\_water}          & NRTL + Wegstein outer loop. \\
\code{flash06\_sweep\_T} / \code{flash07\_gridsweep\_TP} & One knob $\to$ a curve; two knobs $\to$ the flash envelope. \\
\code{flash08\_co2\_water\_package}     & The Henry dilute-solution world, declared as a package manifest. \\
\code{flash09\_n2ch4\_stryjek}          & The $\varphi$--$\varphi$ world: one cubic, both phases, declared $k_{ij}$. \\
\code{vlle03\_audit\_artificial}        & Three-phase VLLE by Gibbs minimisation, on a constructed global minimum. \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/reactors, steady/gibbs}} \\
\code{cstr01\_first\_order} / \code{pfr01\_first\_order} & Newton on the extent; RK4 by hand vs the analytic $X = 1 - e^{-k\tau}$. \\
\code{cstr03} / \code{pfr03\_series\_selectivity} & Series selectivity: back-mixing eats the intermediate (53\,\% vs 79\,\%). \\
\code{cstr04\_adiabatic} / \code{cstr05\_multiplicity} & Non-isothermal CSTR; three steady states, all printed. \\
\code{cstr07\_lhhw\_methylAcetate}      & A published LHHW law on UNIQUAC activities (\S\ref{sec:lhhw}). \\
\code{conversion01\_hda}                & Specified per-pass conversion + recycle --- the honest gas-phase reactor. \\
\code{equil01\_reforming}               & Reforming + shift driven to simultaneous $K_p(T)$ equilibrium. \\
\code{gibbs01\_water\_gas\_shift} / \code{gibbs05\_adiabatic\_flame} & Gibbs minimisation; the flame temperature as a DesignSpec. \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/distillation, absorption}} \\
\code{column01\_benzene\_toluene}       & Wang--Henke bubble-point column. \\
\code{column03\_azeotrope\_mesh}        & The simultaneous MESH through the ethanol/water azeotrope. \\
\code{column05\_reactive\_methylacetate} & Reactive distillation, paper-validated (\S\ref{sec:advanced-distillation}). \\
\code{column09} / \code{column10} / \code{column11} & Tray hydraulics design; the flooded twin; Murphree trays. \\
\code{shortcut01\_benzene\_toluene}     & Fenske--Underwood--Gilliland sizing. \\
\code{absorber01\_NH3\_water} / \code{extract01\_ethanol\_water\_benzene} & Henry absorber; staged LL extraction. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Tutorial} & \textbf{What you learn} \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/heat, power, rotating, hydraulics}} \\
\code{hx01\_heater} / \code{heatExchanger02\_geometry\_U} & The duty device; $U$ computed from the bundle geometry. \\
\code{hxWorkflow1} / \code{hxWorkflow2} & Design from a duty target, then rate the designed bundle. \\
\code{condenser01\_film\_nusselt} / \code{reboiler\_water\_copper} & Duty emerging from film condensation / nucleate boiling (CHF guard). \\
\code{rankine02\_water} / \code{combined02\_brayton\_rankine\_shaft} & IF97 Rankine cycle; the shaft-split combined cycle. \\
\code{compressor01\_air}                & $W_{\mathrm{shaft}} + \eta$; $P_{\mathrm{out}}$ a result. \\
\code{pipe02\_airwater\_twophase}       & Lockhart--Martinelli two-phase pressure drop. \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/membranes, separation, solids, drying, crystallisation}} \\
\code{membrane01\_RO\_NaCl\_seawater}   & The canonical seawater-RO element. \\
\code{membrane07\_scaling\_si}          & Speciation $\times$ membrane: wall scaling three modules ahead of the bulk. \\
\code{psa01\_h2\_psa}                   & Equilibrium-theory pressure-swing adsorption. \\
\code{cyclone01\_dust\_removal}         & PSD-aware cyclone. \\
\code{crystalliser01\_sugar} / \code{crystalliser02\_msmpr} & Cooling equilibrium yield; the MSMPR population balance. \\
\code{sprayDryer01\_sugar}              & Atomisation + droplet drying to a real particle size. \\
\midrule
\multicolumn{2}{@{}l}{\emph{steady/flowsheets, optimisation, evaporation}} \\
\code{process01\_reactor\_flash} / \code{process03\_recycle} & First flowsheet; the tear + Newton-on-tears recycle. \\
\code{sensitivity01\_column\_reflux}    & The reflux sweep to CSV. \\
\code{designSpec01\_triple\_equal\_areas} & Two knobs, two specs on a triple-effect evaporator train. \\
\code{optim01\_column\_reflux} / \code{optim04\_membrane\_recovery\_scaling} & Nelder--Mead; constrained SQP with shadow prices. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{center}\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Tutorial} & \textbf{What you learn} \\
\midrule
\multicolumn{2}{@{}l}{\emph{batch/, ctrl/, electrochem/, plant/}} \\
\code{batch01\_first\_order} / \code{batch02\_adiabatic} & RK4 vs the analytic decay; thermal runaway. \\
\code{still01} / \code{still04} / \code{still05} & Rayleigh; constant-$R$ rectifier; constant-$x_D$ policy at the reflux ceiling. \\
\code{recipe01\_react\_then\_distill}   & A scripted transfer between vessels. \\
\code{ctrl01\_cstr\_temp\_control} / \code{ctrl02\_disturbance\_rejection} & PID tracking; step-disturbance rejection. \\
\code{ed01\_nacl\_desalination}         & The electrodialysis stack, below and at the limiting current. \\
\code{plant/lithiumBrinePlant}          & A fractal multi-sector plant across four thermodynamic worlds. \\
\bottomrule
\end{tabularx}
\end{center}

Every tutorial's header comment states what it demonstrates and, where
applicable, the published number it reproduces.  The
\emph{Tutorials Guide} (\file{docs/tutorialsGuide.pdf}) is the long-form
companion, one chapter per family.

% =====================================================================
\section{Output anatomy}\label{sec:output}

\begin{whatyoulllearn}
What every section of a run's log means, and why the inner iteration
trace is the part you should read.
\end{whatyoulllearn}

A typical steady run with \code{verbosity 3} prints, in order:

\begin{enumerate}[leftmargin=*]
  \item \textbf{Banner} --- version + licence.
  \item \textbf{Case header} --- case dir, database root, the property
        package that was found, which optional dicts are present, and
        the \code{[overlay]}/\code{[estimate]} announcements for any
        case-local component data.
  \item \textbf{State seeding} ---
        \code{[state] seeded N stream(s) from 0/} and the completeness
        check.
  \item \textbf{Topology} --- the source streams and the per-unit
        wiring.
  \item \textbf{Per-unit run} --- for each unit operation:
    \begin{itemize}
      \item the thermo provenance line
            (\code{thermo: inherited (global package)} or the local
            override);
      \item inputs echo ($F, T, P, \mathbf{z}$);
      \item the inner iteration trace (Newton steps with $g(x)$,
            $\mathrm{d}g/\mathrm{d}x$, $\Delta$);
      \item for $\gamma$-models, the outer composition loop;
      \item the result block (regime, $V/F$, $\mathbf{x}$, $\mathbf{y}$,
            $\mathbf{K}$, duty).
    \end{itemize}
  \item \textbf{Recycle loop} (if tears exist) --- per-iteration tear
        residuals, Newton or Wegstein steps, announced convergence aids.
  \item \textbf{Reports} --- one \code{[report]} line per file written.
  \item \textbf{State writing} --- \code{[state] wrote converged/}.
  \item \textbf{Structured result} --- a JSON block between
        \code{<<<Choupo:result-begin>>>} markers carrying every stream,
        KPI, convergence history and advisory (this is what the GUI and
        scripts read).
\end{enumerate}

An excerpt of the part that matters, from the first tutorial:

\begin{lstlisting}
  [outer 0] Rachford-Rice Newton (inner):
    it          V           g(V)         dg/dV         dV
    ----  -----------  -------------  -------------  -------------
    0     0.50000000   -3.75806e-02   -1.87920e-01   -1.99981e-01
    1     0.30001863   7.81978e-04    -1.97366e-01   3.96206e-03
    2     0.30398070   5.66359e-07    -1.97081e-01   2.87374e-06
    3     0.30398357   2.94959e-13    -1.97081e-01   0.00000e+00
\end{lstlisting}

\begin{keyidea}
The inner iteration trace is where the simulator stops being a black
box.  The residual column tells you whether Newton is converging
quadratically (the exponent doubling per line, as above) or linearly
(the problem is ill-conditioned) --- diagnosis you can do by eye.
\end{keyidea}

Batch and control runs additionally write \file{trajectory.csv} on the
\code{writeInterval} grid, and print an epilogue of headline events
(conversions reached, recipe actions fired, reflux limits hit,
ignition delay where relevant).

% =====================================================================
\section{The web GUI: a runner and visualiser}\label{sec:gui}

\begin{whatyoulllearn}
How to open a case in the browser, run it, and read the result --- and
why the GUI never edits your dicts.
\end{whatyoulllearn}

Choupo ships a browser front-end under \file{gui/}.  Its design stance
is deliberate and worth stating up front: \textbf{the GUI is a runner
and visualiser, not an editor.}  Cases are authored as plain-text dicts
on disk, edited with any text editor (increasingly with LLM help).  The
GUI loads those dicts, draws the flowsheet, runs the solver \emph{in
the browser} (compiled to WebAssembly), and renders the log + plots.
The files on disk remain the single source of truth.  The architectural
inspiration is ParaView --- a text-defined case plus a viewer --- not a
block-diagram editor.

\subsection{Starting the GUI}

\begin{lstlisting}
make wasm-gui          # build the WASM solver (choupoSolve + choupoProps)
bin/runGui             # boots the dev server and opens a browser window
bin/runGui --kill      # stops everything
\end{lstlisting}

\subsection{The layout and interactions}

The shell is a thin top bar over a full-screen flowsheet canvas --- no
left tree, no docked panels.  Auxiliary views (log, plots, stream
table, thermo foundation) open on demand as workspaces.

\begin{itemize}[leftmargin=*,nosep,topsep=2pt]
  \item \textbf{Single-click} a node or edge --- raises a floating,
        read-only selection card (labels, units, help tooltips).
  \item \textbf{Double-click} --- pops the unit out, or drills into a
        composite sector (the drilled sub-case is seeded from the
        parent's persisted \file{converged/} state).
  \item \textbf{Drag} nodes and edge bends to lay the flowsheet out;
        the layout persists per case in the browser's
        \code{localStorage}.
  \item \textbf{Stream-class toggles} --- show/hide whole classes of
        streams (e.g.\ energy wires) to declutter.
  \item \textbf{Run} --- streams the same log you would see from
        \code{runCase}; \textbf{Stop} cancels mid-solve.
\end{itemize}

Stream colour encodes phase and comes straight from the solver ($vf$
plus solid mass fraction); it never changes on hover or selection ---
selection is signalled by a halo, not a recolour.

\begin{pitfall}
\textbf{MOCK vs WASM.}  If the WASM solver has not been built
(\code{make wasm-gui}), the GUI falls back to a \emph{mock} adapter
that returns canned numbers so the UI still renders.  The Run badge
reads \code{MOCK} in that case, not \code{WASM} --- check it before
trusting any result.  After changing any \cpp{} source, rebuild the
WASM (it is a separate artefact from the native binary) or the browser
keeps running stale code.
\end{pitfall}

% =====================================================================
\section{Building your own case}\label{sec:build-own}

\begin{whatyoulllearn}
The fastest path from a blank idea to a converged result: scaffold or
clone, wire the topology, author the inlets, initialise, run.
\end{whatyoulllearn}

Two starting points:

\begin{lstlisting}
# (a) scaffold a fresh case in your workspace:
newCase myFlash "isothermal flash of my mixture"

# (b) clone the tutorial closest to your idea:
cp -r tutorials/steady/flash/flash01_benzene_toluene tutorials/steady/flash/myFlash
\end{lstlisting}

Then the loop is always the same:

\begin{enumerate}[leftmargin=*,nosep,topsep=2pt]
  \item edit \file{system/flowsheetDict} (topology) and
        \file{constant/propertyDict} (components + models);
  \item author the \emph{inlet} files in \file{0/} (and a seed for any
        tear);
  \item \code{bin/choupo-init0 <case>} to materialise the remaining
        stream files;
  \item \code{runCase <case>}, read the log, iterate.
\end{enumerate}

\paragraph{Adding a missing component.}  Two locations --- pick the
right tier for what you are adding:

\begin{itemize}[leftmargin=*]
  \item \file{data/standards/components/<name>.dat} --- a contribution
        to the \emph{shared catalogue} of well-curated chemicals
        (primary-cited, per-value provenance; the engine refuses to
        write here --- promotion is a human act).
  \item \file{<case>/constant/components/<name>.dat} ---
        \emph{sample-specific or one-off data}: a particular powder's
        sorption isotherm, a laboratory measurement, a NASA-7 $C_p$
        override.  Case-local files are overlaid block-by-block on the
        standard entry of the same name, so a slim file with just the
        bespoke blocks is enough --- and every overlay is announced in
        the log (\code{[overlay]~...}).
\end{itemize}

When a new file carries a \code{gibbsFormation} block, declare the
\code{phase} of the tabulated $\Delta h^\circ_f$ explicitly
(\code{gas} / \code{liquid} / \code{solid}) --- it tells the solver
which rung of the enthalpy ladder the datum pins (Theory Guide
\S\,``Which rung does the data file pin?'').  A component with no
catalogue entry at all can be \emph{estimated} from its molecular
groups on the props bench (\S\ref{sec:props}) and promoted as a dated,
glass-box proposal.

\paragraph{Using a reaction.}  Declare it in \file{constant/reactions}
and reference it by name (\code{reaction <name>;} or
\code{reactions (...)}) --- \S\ref{sec:reactors}.

\paragraph{Custom unit operations.}  A case may carry its own \cpp{}
unit under \file{<case>/code/} (\file{MyUnit.H/.cpp} + an explicit
registration file); \code{bin/buildCode} compiles it into a per-case
binary.  See
\file{tutorials/steady/userops/\allowbreak userOp01\_yield\_reactor}
and the Developer Guide.

\paragraph{Archiving.}  Seal the case
(\code{bin/choupo-import}, \S\ref{sec:self-contained}) so it runs
without the installation catalogue --- the right form for a thesis
appendix or a paper's supplementary material.

% =====================================================================
\section{Troubleshooting}\label{sec:troubleshoot}

\begin{whatyoulllearn}
The failure modes that account for nearly every support question, and
exactly what to do about each.
\end{whatyoulllearn}

\subsection*{``0/ is incomplete'' / missing or orphan stream file}

The completeness contract (\S\ref{sec:zero}): every stream in the graph
needs a file in \file{0/}, and every file must correspond to a graph
stream.  The error names the offending stream.  Author the inlets, then
\code{bin/choupo-init0 <case>} writes the rest; a leftover file from a
renamed stream must be deleted (the validator refuses to guess which
stream you meant).

\subsection*{``Component does not exist''}

\begin{lstlisting}
ERROR: Database: component file not found: .../components/foo.dat
\end{lstlisting}

The component name is misspelt, or the file is missing.  Names are
case-sensitive and resolved exactly ---
check \file{data/standards/components/} and your
\file{constant/components/} overlay.

\subsection*{``moleFractions do not close''}

Stream-state fractions must sum to 1 within $10^{-6}$
(\S\ref{sec:zero}); the error names the stream and the actual sum.
This is deliberate --- a state file is data, and data with a hole in it
should stop the run, not be silently patched.

\subsection*{Dimension mismatch on a parameter}

\begin{lstlisting}
parameter 'P' expected dimensions [1 -1 -2 0 0] (Pa) but the dict
declared [1 0 0 0 0] (kg).
\end{lstlisting}

A wrong unit suffix (\S\ref{sec:units-suffixes}).  Fix the suffix; do
not strip it (raw numbers pass unchecked).

\subsection*{Build fails after pulling new code}

\begin{lstlisting}
make distclean && make all
\end{lstlisting}

Necessary when headers change incompatibly.  If the GUI misbehaves
after a source change, rebuild the WASM too (\code{make wasm-gui}) ---
it is a separate artefact.

\subsection*{Simulator runs but a stream shows NaN}

Most common: an upstream unit was fed an empty stream ($F = 0$
emptying the composition vector), or Antoine was evaluated far outside
its range (e.g.\ $T$ so low the extrapolated $P^{\mathrm{sat}}$ turns
non-physical).  Re-read the input echo of the first unit whose output
is NaN; \code{bin/runTests} guards every tutorial against silent
NaN/inf for exactly this reason.

\subsection*{Newton fails to converge}

\begin{itemize}[leftmargin=*]
  \item \textbf{Outer composition loop} ($\gamma$-model refresh): try
        \code{outerAccelerator directSubstitution} in
        \file{solverDict} for strongly non-ideal systems, or raise
        \code{maxOuterIter}.
  \item \textbf{Inner Newton} (Rachford--Rice, bubble-$T$): raise
        \code{maxIter}; near a phase boundary, tighten
        \code{compositionTol}.
  \item \textbf{Recycle}: give the tear a better \file{0/} seed (run
        \code{choupo-init0 --force} after improving the inlets), or
        switch \code{recycleSolver} between \code{Newton} and
        \code{Wegstein} --- they have different failure modes.  Every
        aid the solver applies (auto-seed, bound, fallback) is
        announced in the log --- read what it did before overriding it.
\end{itemize}

\subsection*{Wang--Henke column oscillates or crosses an azeotrope}

The bubble-point method is not stable on strongly non-ideal systems.
Select \code{model simultaneous;} on the column --- the whole-block
Newton converges quadratically and refuses to invent a
``past-the-azeotrope'' answer (\S\ref{sec:columns}).

\subsection*{A sweep fails on some points}

Almost always the solver's iteration caps are too small for the edge
points (high reflux, high pressure).  Raise the limits in
\file{solverDict}; failed points are written as \code{nan} and the
sweep continues, so the CSV tells you exactly where the envelope ends.

\begin{keyidea}
If a case behaves unexpectedly, \emph{read the log}.  Every Newton
step, every Wegstein update, every $\gamma$ value, every announced
default is printed.  Black-box debugging is exactly the failure mode
Choupo refuses to participate in.
\end{keyidea}

% =====================================================================
\section{Where to look next}

Choupo's documentation is a family of guides, each with its own job:

\begin{itemize}[leftmargin=*]
  \item \file{docs/theoryGuide.pdf} --- every equation the engine
        solves, derived: flashes, MESH, kinetics, enthalpy ladders,
        population balances, the numerical methods.
  \item \file{docs/tutorialsGuide.pdf} --- the long-form walk through
        the tutorial corpus, family by family.
  \item \file{docs/propsGuide.pdf} --- the property workbench in depth:
        estimation, fitting, consistency, electrolytes.
  \item \file{docs/explorerGuide.pdf} --- the property-explorer view of
        the GUI.
  \item \file{docs/designGuide.pdf} --- sizing, costing, and design
        heuristics.
  \item \file{docs/developerGuide.pdf} --- adding unit operations,
        thermo models, components, outer drivers, post-processors.
  \item \file{tutorials/<category>/<family>/<name>/} --- every tutorial
        is a fully-worked, runnable example with an explanatory header.
  \item \file{src/} --- the ultimate reference.  Every model documented
        here is a readable \cpp{} class; start from
        \file{src/applications/\allowbreak choupoSolve/main.cpp} and
        follow the calls.
\end{itemize}

\vspace{2cm}
\noindent\rule{\linewidth}{0.3pt}
\noindent\small\textit{Colophon.}\quad This manual was typeset in Latin
Modern with the Choupo shared LaTeX preamble.  Sources live in
\file{docs/}; rebuild with \code{cd docs \&\& make user}.

\end{document}
