%-------------------------------------------------------------------------------
% paper/paper.tex — the FULL companion article for Choupo-2607.
% JOSS itself ONLY accepts paper.md (250–1000 words); this .tex is the
% long-form article (preprint / co-author circulation), sharing paper.bib.
% Build:  pdflatex paper && bibtex paper && pdflatex paper && pdflatex paper
%-------------------------------------------------------------------------------
\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage[margin=2.6cm]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\definecolor{linkink}{RGB}{20,60,110}
\usepackage[colorlinks=true, linkcolor=linkink,
            citecolor=linkink, urlcolor=linkink]{hyperref}
\newcommand{\code}[1]{\texttt{#1}}

\title{Choupo: a glass-box chemical process simulator for teaching
       transport phenomena and process engineering}
% TODO(Vítor): decidir a lista de autores (a autoria do paper é decisão tua,
% não deriva do copyright do código) e acrescentar o ORCID.
\author{Vítor Geraldes\\[2pt]
  \small Department of Chemical Engineering, Instituto Superior Técnico,\\
  \small Universidade de Lisboa, Portugal}
\date{20 July 2026 \,·\, describes release Choupo-2607}

\begin{document}
\maketitle

\begin{abstract}
Choupo is an open-source, educational chemical process simulator written in
C++17 with zero external dependencies, in which every numerical method is
readable, every model choice is announced at run time, and conservation
laws are first-class teaching artefacts.  Cases are plain-text
dictionaries in a file-first layout; each case seals its own property
records under a checksummed manifest, so it reruns identically with no
installation catalogue present.  Release Choupo-2607 ships four solver
binaries by problem class, 49 unit-operation models, a curated
thermodynamic catalogue (247 components; NRTL, UNIQUAC, Wilson, UNIFAC,
COSMO-SAC, SRK, Peng--Robinson, a non-associating PC-SAFT core, Pitzer and
electrolyte-NRTL electrolytes, IAPWS-IF97 steam), and 288 runnable
tutorial cases validated by a golden-master regression suite.  Every
converged steady run emits a plant-boundary element balance; batch
campaigns carry material and energy ledgers closing to machine precision
as exact state differences on the elements-of-formation datum; dynamic
runs integrate a conservation ledger on accepted integrator steps.  A
browser front-end runs the same solvers compiled to WebAssembly as a
runner and visualiser, never a black-box editor.  Choupo is an
independent, self-funded personal project, licensed GPL-3.0-or-later.
\end{abstract}

\section{Introduction}

Commercial process simulators dominate both industry and the classroom,
but they are pedagogically opaque: the student specifies a column and
receives converged profiles without ever seeing a residual, a Jacobian, or
the reference state of an enthalpy.  Open-source alternatives exist, but
they generally prioritise breadth and industrial workflows over the
readability of the computation itself.  Choupo takes the opposite stance
--- \emph{transparency over breadth}.  At the default verbosity every
Newton iteration, every $K$-value and every convergence aid is printed and
attributed; the engine refuses to substitute silent defaults for missing
data (a missing binary-interaction pair is announced before it is
ideal-defaulted; a missing formation enthalpy makes an energy balance
refuse rather than fabricate a zero); and the source itself --- small
C++17 classes with no external numerical libraries --- is meant to be
read by students, not only executed.

Choupo is an independent personal project, conceived, self-funded and
developed by the author in his free time, on personal equipment; it is not
an institutional product.  It is designed for teaching transport phenomena
and process engineering --- and for the LLM era of case authoring: the
documentation ships a machine-oriented reference so a student can author
case dictionaries with an assistant while the engine remains the sole
arbiter of the physics.  This article describes release Choupo-2607.

\section{Architecture and case grammar}

\subsection{Three layers, four binaries}

The engine is organised in three layers.  The \emph{simulator core} turns
one flowsheet into one solution: streams, unit results and key performance
indicators.  An optional \emph{outer driver} wraps that pure function to
run parameter sweeps, design specifications, optimisation
(Nelder--Mead \citep{nelder1965} or a line-search SQP) and regression of
thermodynamic parameters against experimental data
(Levenberg--Marquardt \citep{marquardt1963}).  An optional
\emph{post-processor} chain augments a converged solution with equipment
sizing and Guthrie-style costing.  Four binaries partition the problem
classes: \code{choupoSolve} (steady state, $F(x)=0$, with a
Newton-on-tears recycle solver and Wegstein acceleration
\citep{wegstein1958} as an alternative), \code{choupoBatch} (batch
campaigns, $\mathrm{d}Y/\mathrm{d}t=f$, driven by a recipe of
time-triggered events), \code{choupoCtrl} (dynamic continuous units with
feedback-control loops) and \code{choupoProps} (a property workbench:
evaluations, scans, consistency tests and parameter fitting).  Every
extensible base class uses an explicit factory registered in
\code{main()} --- no self-registration macros, no static-initialiser
magic --- so a student can answer ``where does this model come from?'' by
reading one function.

\subsection{File-first cases}

A case is a directory of plain-text dictionaries.  \code{system/} carries
the control, topology, solver, outer-driver and post-processing
dictionaries; \code{constant/} carries the thermophysical system and any
reaction, crystallisation or drying-kinetics libraries; \code{0/} carries
the complete initial state, one file per stream --- the flowsheet
dictionary declares \emph{topology only}, and a stream's role (feed,
internal, product) is inferred from that topology rather than declared.
The steady solution is persisted to \code{converged/}, one file per
stream, so a hierarchical (``fractal'') plant can be drilled into with the
child case materialised from the parent's persisted state.  Completeness
is contractual: a missing stream file is a fatal error, and the
\code{choupo-init0} tool materialises missing files by explicit
propagation from authored inlets --- it refuses to invent a recycle seed.

Two properties of this grammar matter pedagogically.  First,
\emph{self-containment}: an importer copies the transitive closure of
every property record a case consumes into the case's own
\code{constant/} tree under a manifest with per-record SHA-256 checksums;
a sealed case reruns byte-for-byte with the installation catalogue hidden,
which is exactly what a thesis archive needs.  Second, \emph{units on
every scalar}: dictionary values carry named units (\code{P 1 bar;}) that
the parser converts to SI and cross-checks dimensionally, with a clear
error on mismatch.

\section{Thermodynamics: declared, not defaulted}

A case carries a single thermophysical-system dictionary
(\code{constant/thermoPhysPropDict}) that names its components and its
equilibrium \emph{formulation} --- the $K$-value structure the whole
vapour--liquid equilibrium runs on: \code{gammaPhi}
($K_i=\gamma_iP^{\mathrm{sat}}_i/\varphi_iP$), \code{gammaGamma} (two or
more liquid phases, one excess-Gibbs surface each), \code{diluteSolution}
(solvent on the Raoult convention, each solute on the full
Krichevsky--Kasarnovsky Henry form \citep{krichevsky1935}), \code{phiPhi}
(one cubic equation of state serving both phases,
$K_i=\varphi^L_i/\varphi^V_i$) and \code{electrolyteGammaPhi} (aqueous
electrolyte over an ideal vapour).  The builder assembles the declared
formulation and refuses unknown shapes by name; the run log announces the
assembled world, each phase's model and reference convention, and every
parameter file engaged with its literature source.

The activity library spans NRTL \citep{renon1968}, UNIQUAC
\citep{abrams1975}, Wilson, predictive UNIFAC \citep{fredenslund1975} and
COSMO-SAC in the 2002 formulation \citep{lin2002} cross-checked against
the NIST benchmark implementation \citep{bell2019}, with each compound
carrying its own named $\sigma$-profile sets.  Cubic equations of state
(SRK \citep{soave1972}, Peng--Robinson \citep{peng1976}) deliver both
roots, and a validated non-associating PC-SAFT core \citep{gross2001} is
available on the $\varphi$--$\varphi$ route.  Electrolytes are served at
two levels: a single-salt Pitzer adapter \citep{pitzer1973} and
electrolyte-NRTL \citep{chen1986} for flowsheet VLE --- including the
generalized segment-based mixed-solvent form \citep{chen2004} --- and a
multi-ion Harvie--M{\o}ller--Weare speciation engine \citep{harvie1984}
(with a Davies option for dilute work) for saturation-index and scaling
studies.  Water-dominated power cycles can route through a hand-rolled
IAPWS-IF97 steam kernel \citep{wagner2000}.  Transport properties (Chung,
Eucken, Fuller, Vogel, Andrade, Sato--Riedel, Latini, Wilke--Chang,
Scheibel, Brock--Bird) are declared per phase.

Property estimation is deliberately a \emph{curation-time} act, not a
runtime fallback: group-contribution estimates (Joback and relatives) are
produced at the property workbench as reviewable data files flagged as
estimates, and the runtime builder loads and verifies --- it never
estimates.  Enthalpy uses one datum throughout: the elements at
298.15\,K, so energy balances may legally cross reactors; the heat of a
reaction is always $\sum_i\nu_ih_i(T)$ over species carrying formation
data, and a dissolving salt's solid formation enthalpy is derived from
its aqueous ions plus a cited heat of solution, never stored as a second,
drifting copy.

\section{Numerical methods, all readable}

Every solver is hand-written and small: Newton--Raphson in one and $n$
dimensions with visible residuals and Jacobians, Gauss elimination,
Wegstein acceleration \citep{wegstein1958}, a Newton-on-tears recycle
solver, Runge--Kutta~4 and an adaptive Rosenbrock method for stiff
dynamics, Nelder--Mead \citep{nelder1965} and a line-search SQP for
optimisation, Levenberg--Marquardt \citep{marquardt1963} for regression,
and Michelsen's tangent-plane stability test \citep{michelsen1982} for
phase-split detection.  Liquid--liquid equilibrium is solved by direct
Gibbs-energy minimisation with multi-start seeding rather than
successive substitution --- the symmetric-model saddle at $K=1$ is a
known trap the code refuses to fall into.  Convergence aids obey a
``no silent crutch'' rule: every initial guess, tear estimate and bound
is first-class in the dictionaries, owned by the student, and announced
when it binds.

\section{Unit operations}

Release 2607 registers 49 unit-operation models: the flash and
saturation family; seven reactor forms (stoichiometric conversion,
equilibrium by named reactions, Gibbs free-energy minimisation, CSTR and
PFR with power-law and Langmuir--Hinshelwood kinetics, plus batch and
dynamic reactors) sharing one reaction grammar; rigorous distillation by
bubble-point Wang--Henke \citep{wang1966} and by simultaneous MESH
Newton, with Murphree tray efficiencies and a tray-hydraulics
rating/design pass; shortcut distillation (Fenske--Underwood--Gilliland);
absorbers, strippers and liquid--liquid extractors; heat exchangers
($\varepsilon$-NTU rating and a geometry-based design mode), multi-stream
exchangers and a phase changer; evaporators; crystallisers (equilibrium,
MSMPR and batch population-balance forms); membrane separations (a
solution-diffusion spiral-wound module, a Donnan-steric pore model for
charged nanofiltration, electrodialysis and ion exchange); adsorption
(isotherm benches, linear-driving-force uptake, fixed-bed breakthrough
and twin-bed pressure/temperature-swing shortcuts); solid and spray
dryers with an atomiser library; cyclones and bag filters; pneumatic
conveying; rotating equipment specified by shaft work (outlet pressure is
a result, not an input); pipes with Darcy--Weisbach hydraulics; and batch
vessels including a passive accumulator that receives a still's
distillate continuously.  Heat and shaft work travel on explicit energy
wires between units; duties are allocated to a utility catalogue by
temperature level, and the browser front-end computes pinch composite
curves from a converged run.

\section{Conservation as the curriculum}

The release's defining feature is a three-level balance discipline ---
mass, per-element atoms, energy --- owned by the engine, with the
interface only drawing engine artefacts.

\textbf{Steady.}  Every converged \code{choupoSolve} run emits a
plant-boundary \emph{element balance} by default: one shared formula
parser builds the atom matrix $a_{ie}$ from each component's formula (or
from a declared, provenance-gated elemental composition for
formula-less substances such as petroleum cuts), and the report prints,
per element, the atoms in, atoms out, residual and closure.  Because a
$+$C error can cancel a $-$O error in a signed total, the conservation
claim requires \emph{every} element to close; the compact indicator is
the worst per-element deviation.  Claims are tri-state --- full, partial
(a declared composition with unaccounted mass), unavailable (a species
that cannot resolve, named, with the reason) --- never silent zeros.

\textbf{Batch.}  Campaigns carry material and energy ledgers: one record
per transfer with per-package enthalpy, one record per segment of
constant physics, every integral an exact state difference on the
elements datum (never a quadrature).  The reference tutorial --- react,
transfer, then distil into a receiver --- retains all initial mass in
its three vessels and closes mass and every element to $\sim\!10^{-15}$
relative, with the energy ledger closing to the integration honesty of
the mix solve.

\textbf{Dynamic.}  \code{choupoCtrl} integrates a conservation ledger on
\emph{accepted} integrator steps (mass inventory plus per-element
residuals over time).  The physical first-law closure for the dynamic
CSTR honestly \emph{refuses}: its energy equation is a heat-capacity/
convective model, not the exact derivative of a stored internal energy,
and the ledger says so rather than claim a closure the model cannot
support.

\section{The browser front-end}

The graphical interface (React, with the solvers compiled to WebAssembly)
is a \emph{runner and visualiser}: cases are authored as dictionaries on
disk, and the browser loads, runs, and draws --- flowsheet topology with
recipe transfers as labelled dashed edges, stream tables, global-balance
views (the element balance as total atoms in/out with per-element
drill-down), campaign Gantt charts in which continuous actions span their
true intervals, trajectories, profiles, pinch composite curves, and an
interactive property-exploration bench (binary envelopes,
McCabe--Thiele, flash constructions, LLE maps, psychrometric charts,
scaling scans, steam tables).  Every view deep-links to the section of
the PDF theory guide that derives it.  The same physics runs in the
browser and at the command line; the interface never computes physics of
its own.

\section{Validation and quality gates}

The regression suite runs all 288 tutorial cases on every change: an
exit-code check alone is banned (a solver can return NaN with exit code
0), so every case passes a NaN/infinity guard, and 255 cases additionally
pin key performance indicators against golden-master references.
Doctrine gates enforce structural honesty (retired grammar cannot
reappear in any document a student reads; sealed copies are audited
against the live catalogue; the element-balance and control-ledger
contracts are exercised by dedicated adversarial cases).  Physical
validation anchors the models to primary data: Henry's-law solubilities
against Wiebe/Larson measurements ($\sim$2\%), the N$_2$--CH$_4$
$\varphi$--$\varphi$ isotherm against Stryjek's data ($\sim$2\%),
mixed-solvent electrolyte-NRTL against Esteso's NaCl/ethanol--water
activity coefficients \citep{esteso1989} with zero refitting (AAD
$\sim$4.3\% over 47 points, with the systematic deviation at high
ethanol declared as the ion-pairing physics a complete-dissociation
model cannot carry), the PC-SAFT core against literature ($\sim$1\%),
COSMO-SAC bit-for-bit against an independent implementation of the same
equations and profiles, an IF97 Rankine cycle closing its first law to
$10^{-12}$, and Pitzer--HMW seawater saturation indices inside the
published oceanographic band.

\section{Data governance and licensing}

The public repository redistributes no third-party databank values.  The
curated catalogue (247 components, 205 Henry pairs, 55 Pitzer and 3
eNRTL parameter sets, 4 membranes, 3 adsorbents, 9 plant utilities)
cites the primary source per value; beyond it, Choupo ships its own
open group-contribution \emph{estimates}, clearly flagged as such, and a
private, git-ignored local tier receives whatever data a user imports or
licenses --- consumed values are announced as unverified.  The code is
GPL-3.0-or-later; the documentation manuals are CC~BY-SA~4.0; bundled
data files may carry their own open licences as an aggregation.  The
\emph{Choupo} name is a trademark of TalentGround Lda.

\section{Limitations}

Choupo does not aim at breadth parity with industrial tools, and says so:
the component catalogue is curated rather than encyclopaedic;
bubble-point distillation is slow near azeotropes (the simultaneous MESH
method is the documented escape); the dynamic CSTR's physical energy
closure is refused pending a reformulation of its energy equation around
a stored state function; and PC-SAFT ships without the association term.
Each limitation is stated in the documentation next to the feature it
bounds --- the project treats an honest refusal as a feature.

\section*{Availability}

Source, tutorials, documentation and the browser application:
\url{https://github.com/choupo-admin/choupo} and
\url{https://www.choupo.org} (release tag \code{Choupo-2607}).

\section*{Acknowledgements}

Substantial parts of the initial implementation, documentation and
tutorial corpus were produced with assistance from Anthropic Claude; the
published project is human-curated, reviewed and maintained by the
authors.

\bibliographystyle{unsrtnat}
\bibliography{paper}

\end{document}
